### 2020-05-05 Writeup/manuscript.tex
Added writeup

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\title{Template for preparing your research report submission to PNAS using Overleaf}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Peter J. Kohler}
\author[b]{Oda \r{A}sgard} 
\author[b]{Alasdair D. F. Clarke}

\affil[a]{Stanford University, Department of Psychology}
\affil[b]{University of Essex, Department of Psychology, Colchester, UK, CO4 3SQ}

% Please give the surname of the lead author for the running footer
\leadauthor{Kohler}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any conflict of interest here.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: author.two\@email.com}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
A century of vision research has demonstrated that symmetry contributes to numerous domains of visual perception \cite{RN1311, RN1166,RN1682,RN1824}. In a 2-D image, the four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in 17 distinct ways. These 17 “wallpaper” groups \cite{RN1562,RN1563,RN1425} obey a hierarchy, determined by mathematical group theory, where simpler groups are subgroups of more complex ones\cite{RN1711}. Here we measure symmetry-specific Steady-State Visual Evoked Potentials (SSVEPs), and find that the partial ordering of SSVEPs from visual cortex in human participants ($n=25$), preserve the hierarchical relationships between the wallpaper groups in $88.3\%$ (53 out of 60) of cases. This remarkable consistency between the structure of symmetry representations and mathematical group theory, is likely generated over visual development, through implicit learning of regularities in the environment. 
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{S}ymmetries are present at many scales in images of natural scenes, due to a complex interplay of physical forces that govern pattern formation in nature. The importance of symmetry for visual perception has been known at least since the gestalt movement of the early 20th century. Since then, symmetry has been shown to contribute to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection \cite{RN1337}. Most of this work has focused on mirror symmetry,

In the two spatial dimensions relevant for images, the four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in 17 distinct ways, the “wallpaper” groups \cite{RN1562,RN1563,RN1425}. The 17 wallpaper groups obey a hierarchy of complexity, determined by mathematical group theory, where simpler groups are sub-groups of more complex ones \cite{RN1711}. Here we use Steady-State Visual Evoked Potentials (SSVEPs) to measure responses to 16 out of 17 wallpaper groups and show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups. Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases. Visual cortex thus encodes all the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory. Given that most participants had no knowledge of group theory, the ordered structure of visual responses to wallpaper groups is likely learned implicitly from regularities in the visual environment.

\section*{Results}

The visual stimuli for our experiment were multiple exemplar images belonging to each of the 17 wallpaper groups, generated from random-noise textures, as described in detail elsewhere \cite{RN1725}. To isolate brain activity specific to the symmetry structure in the images from activity associated with modulation of low-level features, we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group. Because all wallpapers are periodic images due to their lattice tiling structure, the phase-scrambled images are also a wallpaper group (P1). P1 contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \cite{RN1425}. 
Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of P1 exemplars, that were matched in terms of their Fourier power spectrum. Because the P1 group serves a control stimulus in this approach, the experiment was restricted to the remaining 16 groups. This design allows us to isolate responses to structural features beyond the shared power spectrum, including any symmetries other than translation, in the odd harmonics of the image update frequency \cite{RN1725,RN1540} [NORCIA 2002 MISSING]. Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1. A wallpaper group is a topologically discrete group of isometries of the Euclidean plane, i.e. transformations that preserve distance \cite{RN1425}. Wallpaper groups differ in the number and kind of these transformations. In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \cite{RN1425}. Subgroup relationships between wallpaper groups can be distinguished by their indices. The index of a subgroup relationship is the number of cosets, i.e. the number of times the subgroup is found in the outer group \cite{RN1425}. As an example, let us consider groups P6 and P2. If we ignore the translations in two directions that both groups share, group P6 consists of the set of rotations {$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$}, in which P2 {$0^{\circ}$, $180^{\circ}$} is contained. P2 is thus a subgroup of P6, and the full P6 set can be generated by every combination of P2 and rotations {$0^{\circ}$, $120^{\circ}$, $240^{\circ}$}. Because P2 is repeated three times in P6, P2 is a subgroup of P6 with index 3 \cite{RN1425}.

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{../ac_models_eeg_data/plots/subgroup_comprms_Occ_Odd.png}
\caption{Posterior distributions for the difference in mean RMS EEG response. Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero. We can see that 55/64 subgroup relationships have $p(\Delta|data)>0.9$.}
\label{fig:durations_rotations}
\end{figure}

\section*{Discussion}

Here we show that beyond merely responding to the elementary symmetry operations of reflection \cite{RN1170} and rotation \cite{RN1725}, the visual system explicitly represents hierarchical structure of the 17 wallpaper groups, and thus the compositions of all four of the fundamental symmetry transformations (rotation, reflection, translation, glide reflection) which comprise regular textures. The RMS measure of SSVEP amplitude, preserves the complex hierarchy of subgroup relationships among the wallpaper groups \cite{RN1711}. Out of a total of 60 relationships, 53 were preserved in a significant number of participants, and 49 were significant even at a stricter threshold (p < 0.002). The ordering was highly stable in individual participants, with an average preservation rate of ~21 of 25 participants across all 60 relationships (see Figure 3). This remarkable consistency was specific to the odd harmonics of the stimulus frequency, that capture the symmetry-specific response \cite{RN1725} and to electrodes in an ROI over occipital cortex. When the same analysis was done on the even harmonics of the occipital cortex ROI, the ordering of responses was much less apparent (see Figure S2) and preservation rates much lower (see Figure S4). The odd harmonics from electrodes in an ROI over parietal cortex, showed even weaker evidence of preserving the hierarchy among sub-groups (see Figure S5). Importantly, no relationships were preserved in either of these control analyses that were not also preserved in the main analysis of the odd harmonics in the occipital cortex ROI. The current data provide a complete description of the visual system’s response to symmetries in the 2-D plane. Our design does not allow us to independently measure the response to P1, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 2), we can conclude that the relationships between P1 and all other groups, where P1 is the subgroup, are also preserved by the visual system. The subgroup relationships are not obvious perceptually, and most participants had no knowledge of group theory. Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships. Furthermore, behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, they are generally error-prone when asked to assign exemplar images to the appropriate group \cite{RN172}. The correspondence between responses in the visual system and group theory that we demonstrate here, may reflect a form of implicit learning that depends on the structure of the natural world. The environment is itself constrained by physical forces underlying pattern formation and these forces are subject to multiple symmetry constraints \cite{RN1634}. The ordered structure of responses to wallpaper groups could be driven by a central tenet of neural coding, that of efficiency. If coding is to be efficient, neural resources should be distributed in such a way that the structure of the environment is captured with minimum redundancy considering the visual geometric optics, the capabilities of the subsequent neural coding stages and the behavioral goals of the organism \cite{RN1758, RN1760, RN1757, RN1756}. Early work within the efficient coding framework suggested that natural images had a $1/f$ spectrum and that the corresponding redundancy between pixels in natural images could be coded efficiently with a sparse set of oriented filter responses, such as those present in the early visual pathway \cite{RN1740, RN1446}. Our results suggest that the principle of efficient coding extends to a much higher level of structural redundancy – that of symmetries in visual images. The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective and occlusion. Near-regular textures, however abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \cite{RN1519}. The correspondence between brain data and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space. This framework resembles norm-based encoding strategies that have been proposed for other stimulus classes, most notably faces \cite{RN435}, and leads to the prediction that adaptation to wallpaper patterns should distort perception of near-regular textures, similar to the aftereffects found for faces \cite{RN1768}. Field biologist have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \cite{RN1775}. In the norm-based encoding framework, wallpaper groups can be considered super-textures, exaggerated examples of the near-regular textures that surround us. Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \cite{RN1764}. Wallpaper groups are visually compelling and have been widely used in human artistic expression going back to the Neolithic age \cite{RN1949}. If wallpapers are super-textures, their prevalence may be a direct consequence of the strategy the human visual system uses for encoding visual textures.

% \section*{Guide to using this template on Overleaf}

% Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout. For more detailed information please see the \href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information for Authors}.

% If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}. You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.

% \subsection*{Author Affiliations}

% Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. Authors with an ORCID ID may supply this information at submission.

% \subsection*{Submitting Manuscripts}

% All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

% \subsection*{Format}

% Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References. Other orders and headings are permitted.

% \subsection*{Manuscript Length}

% PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a Direct Submission research article is six pages and a Direct Submission Plus research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.  When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for Direct Submission Plus.

% \subsection*{Data Archival}

% PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.

% % \begin{SCfigure*}[\sidecaptionrelwidth][t]
% % \centering
% % % \includegraphics[width=11.4cm,height=11.4cm]{frog}
% % \caption{This caption would be placed at the side of the figure, rather than below it.}\label{fig:side}
% % \end{SCfigure*}

% \subsection*{Digital Figures}

% Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size. Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.  Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

% Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

% Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.

% \subsection*{Tables}
% In addition to including your tables within this manuscript file, PNAS requires that each table be uploaded to the submission separately as a “Table” file.  Please ensure that each table .tex file contains a preamble, the \verb|\begin{document}| command, and the \verb|\end{document}| command. This is necessary so that the submission system can convert each file to PDF.

% \subsection*{Single column equations}

% Authors may use 1- or 2-column equations in their article, according to their preference.

% To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.

% Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used.

% \begin{figure*}[bt!]
% \begin{align*}
% (x+y)^3&=(x+y)(x+y)^2\\
%        &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
%        &=x^3+3x^2y+3xy^3+x^3. 
% \end{align*}
% \end{figure*}

% \begin{table}%[tbhp]
% \centering
% \caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
% \begin{tabular}{lrrr}
% Species & CBS & CV & G3 \\
% \midrule
% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
% \bottomrule
% \end{tabular}

% \addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
% \end{table}

% \subsection*{Supporting Information (SI)}

% Authors should submit SI as a single separate PDF file, combining all text, figures, tables, movie legends, and SI references.  PNAS will publish SI uncomposed, as the authors have provided it.  Additional details can be found here: \href{http://www.pnas.org/page/authors/journal-policies}{policy on SI}.  For SI formatting instructions click \href{https://www.pnascentral.org/cgi-bin/main.plex?form_type=display_auth_si_instructions}{here}.  The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.  Refer to the SI Appendix in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc.

% Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.

% \subsubsection*{SI Datasets}

% Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.

% \matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.

\subsection*{Participants}
Twenty-five participants (11 females, mean age 28.7±13.3) took part in the EEG experiment. Their informed consent was obtained before the experiment under a protocol that was approved by the Institutional Review Board of Stanford University. 11 participants (8 females mean age 20.73±1.21) took part in the psychophysics experiment. All participants had normal or corrected-to-normal vision. Their informed consent was obtained before the experiment under a protocol that was approved by the University of Essex's Ethics Committee.

\subsection*{Stimulus Generation}
Exemplars from the different wallpaper groups were generated using a modified version of the methodology developed by Clarke and colleagues\cite{RN172} that we have described in detail elsewhere\cite{RN1725}. Briefly, exemplar patterns for each group were generated from random-noise textures, which were then repeated and transformed to cover the plane, according to the symmetry axes and geometric lattice specific to each group. The use of noise textures as the starting point for stimulus generation allowed the creation of an almost infinite number of distinct exemplars of each wallpaper group. For each exemplar image, phase-randomized control exemplars were generated that had the same power spectrum as the exemplar images for each group. The phase scrambling eliminates rotation, reflection and glide-reflection symmetries within each exemplar, but the phase-scrambled images inherent the spectral periodicity arising from the periodic tiling. This means that all control exemplars, regardless of which wallpaper group they are derived from, degenerate to another symmetry group, namely P1. P1 is the simplest of the wallpaper groups, and contains only translations of a region whose shape derives from the lattice. Because the different wallpaper groups have different lattices, P1 controls matched to different groups have different power spectra. Our experimental design takes these differences into account by comparing the neural responses evoked by each wallpaper group to responses evoked by the matched control exemplars.

\subsection*{Stimulus Presentation}
Stimulus Presentation. For the EEG experiment, the stimuli were shown on a 24.5" Sony Trimaster EL PVM-2541 organic light emitting diode (OLED) display at a screen resolution of $1920\times1080$ pixels, 8-bit color depth and a refresh rate of 60 Hz, viewed at a distance of 70 cm. The mean luminance was 69.93 cd/m2 and contrast was 95\%. The diameter of the circular aperture in which the wallpaper pattern appeared was $13.8^\circ$ of visual angle presented against a mean luminance gray background. Stimulus presentation was controlled using in-house software.

For the psychophysics experiment, the stimuli were shown on a $48 \times 27$cm VIEWPixx/3D LCD Display monitor, model VPX-VPX-2005C, resolution $1920 \times 1080$ pixels, with a viewing distance of approximately 40cm and linear gamma. Stimulus presentation was controlled using MatLab and Psychtoolbox-3 \cite{kleiner2007,brainard1997spatial}. The diameter of the circular aperture for the stimuli was $21.5^\circ$.

\subsection*{EEG Procedure}
Visual Evoked Potentials were measured using a steady-state design, in which P1 control images alternated with test images from each of the 16 other wallpaper groups[2]. Exemplar images were always preceded by their matched P1 control image. A single 0.83 Hz stimulus cycle consisted of a control P1 image followed by an exemplar image, each shown for 600 ms. A trial consisted of 10 such cycles (12 sec) over which 10 different exemplar images and matched controls from the same rotation group were presented. For each group type, the individual exemplar images were always shown in the same order within the trials. Participants initiated each trial with a button-press, which allowed them to take breaks between trials. Trials from a single wallpaper group were presented in blocks of four repetitions, which were themselves repeated twice per session, and shown in random order within each session. To control fixation, the participants were instructed to fixate a small white cross in the center of display. To control vigilance, a contrast dimming task was employed. Two times per trial, an image pair was shown at reduced contrast, and the participants were instructed to press a button on a response pad. We adjusted the contrast reduction such that average accuracy for each participant was kept at ~85\% correct, so that the difficulty of the vigilance task was kept constant.

\subsection*{Psychophysics Procedure}
The experiment consisted of 16 blocks, one for each of the wallpaper groups (excluding $P1$). In each trial, participants were presented with two stimuli (one of which was the wallpaper group for the current block of trials, the other being $P1$), one after the other (inter stimuli interval of 700ms). After each stimuli had been presented, it was masked with white noise for 300ms. After both stimuli had been presented, participants made a response on the keyboard to indicate whether they thought the first or second contained the most symmetry. Each block started with 10 practise trials, (stimulus display duration of 500ms) to allow participants to familiarise themselves with the current block's wallpaper pattern. If they achieved an accuracy of 9/10 in these trials they progressed to the rest of the block, otherwise they carried out another set of 10 practise trials. This process was repeated until the required accuracy of 9/10 was obtained. The rest of the block consisted of four interleaved staircases (using the QUEST algorithm \cite{watson1983quest}, full details given in the SI) of 30 trials each. on average, a block of trials took around 10 minutes to complete.

\subsection*{EEG Acquisition and Preprocessing}
Electroencephalogram Acquisition and Preprocessing. The time-locked Steady-State Visual Evoked Potentials were collected with 128-sensor HydroCell Sensor Nets (Electrical Geodesics, Eugene, OR) and were band-pass filtered from 0.3 to 50 Hz. Raw data were evaluated off line according to a sample-by-sample thresholding procedure to remove noisy sensors that were replaced by the average of the six nearest spatial neighbors. On average, less than 5\% of the electrodes were substituted; these electrodes were mainly located near the forehead or the ears. The substitutions can be expected to have a negligible impact on our results, as the majority of our signal can be expected to come from electrodes over occipital, temporal and parietal cortices. After this operation, the waveforms were re-referenced to the common average of all the sensors. The data from each 12s trial were segmented into five 2.4 s long epochs (i.e., each of these epochs was exactly 2 cycles of image modulation). Epochs for which a large percentage of data samples exceeding a noise threshold (depending on the participant and ranging between 25 and 50 $\mu$V) were excluded from the analysis on a sensor-by-sensor basis. This was typically the case for epochs containing artifacts, such as blinks or eye movements. The use of steady-state stimulation drives cortical responses at specific frequencies directly tied to the stimulus frequency. It is thus appropriate to quantify these responses in terms of both phase and amplitude. Therefore, a Fourier analysis was applied on every remaining epoch using a discrete Fourier transform with a rectangular window. The use of epochs two-cycles (i.e., 2.4 s) long, was motivated by the need to have a relatively high resolution in the frequency domain, $\delta$f = 0.42 Hz. For each frequency bin, the complex-valued Fourier coefficients were then averaged across all epochs within each trial. Each participant did two sessions of 8 trials per condition, which resulted in a total of 16 trials per condition.

\subsection*{EEG Analysis}
Response waveforms were generated for each group by selective filtering in the frequency domain. For each participant, the average Fourier coefficients from the two sessions were averaged over trials and sessions. The Steady-State Visual Evoked Potentials paradigm we used allowed us to separate symmetry-related responses from non-specific contrast transient responses. Previous work has demonstrated that symmetry-related responses are predominantly found in the odd harmonics of the stimulus frequency, whereas the even harmonics consist mainly of responses unrelated to symmetry, that arise from the contrast change associated with the appearance of the second image[2-4]. This functional distinction of the harmonics allowed us to generate a single-cycle waveform containing the response specific to symmetry, by filtering out the even harmonics in the spectral domain, and then back-transforming the remaining signal, consisting only of odd harmonics, into the time-domain. For our main analysis, we averaged the odd harmonic single-cycle waveforms within a six-electrode region of interest (ROI) over occipital cortex (electrodes 70, 74, 75, 81, 82, 83). These waveforms, averaged over participants, are shown in Figure 2 in the main paper. The same analysis was done for the even harmonics (see Figure S1) and for the odd harmonics within a six electrode ROI over parietal cortex (electrodes 53, 54, 61, 78, 79, 86; see Figure S2). The root-mean square values of these waveforms, for each individual participant, were used to determine whether each of the wallpaper subgroup relationships were preserved in the brain data.

\subsection*{Bayesian Analysis of EEG and Psychophysical data}
Bayesian analysis was carried out using R (v3.6.1) \cite{R} with the \texttt{brms} package (v2.9.0 \cite{burkner2017} and rStan (v2.19.2 \cite{rStan}). The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16 level factor, and random effects for participant. The EEG data and display thresholds were modelled using log-normal distributions with weakly informative ($N(0, 2)$) priors. After fitting the model to the data, samples were drawn from the posterior distribution for each mean of the EEG response (display duration) for each wallpaper group. These samples were then recombined to calculate the distribution of differences for each pair of subgroup and super-group. These distributions were then summarised by computing the conditional probability that of obtaining a positive (negative) difference, $p(\Delta | data)$.

For further technical details, please see the supplementary materials where the full R code and model specification can be found.

\showmatmethods{} % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliographystyle{pnas-new} 
\bibliography{literature}

\end{document}

### 2020-05-05 experiment/experiment.m
Added quest detection study

@@ -1,0 +1,282 @@
+function experiment
+
+
+
+PsychImaging(‘PrepareConfiguration’);
+PsychImaging(‘AddTask’, ‘FinalFormatting’, ‘DisplayColorCorrection’, ‘SimpleGamma’);
+// >>Open Window code<<
+PsychColorCorrection(‘SetEncodingGamma’, win, 1/gamma);
+
+
+
+groupsToTest = {...
+    'P2', 'PM' ,'PG', 'CM', 'PMM', 'PMG', 'PGG', 'CMM', 'P4', 'P4M', 'P4G', 'P3', 'P3M1', 'P31M', 'P6', 'P6M'};
+
+obsID = input('please enter participant id number: ');
+resultsFolder = ['results/person', int2str(obsID)];
+mkdir(resultsFolder);
+
+% init random seed!
+rng(obsID, 'twister');
+
+
+%% set parameters for experiment
+params.staircaseLength = 30; % number of trials in each staircase
+params.nRepetitions = 4; % number of interleaved staircases
+params.nLearning = 10; % total number of practise trials
+
+params.initDisplayDur = [0.256, 0.128, 0.064, 0.032, 0.064]; % initial display durations for each staircase
+params.practiceDur = 0.500; % how long should the practise duration
+
+params.minDuration = 0.005;
+params.maxDuration = 1.000;
+
+params.fixCrossDur = 1.000;
+params.isiDuration = 0.700; % inter-stimuli blank
+params.noiseMaskDuration = 0.300;
+params.stimuliContrast = 0.5;
+params.rotate = 1;
+
+params.imagesPerCondition = 50;
+
+addpath('scripts/');
+
+% check which groups have been done already!
+files = dir([resultsFolder '/*.txt']);
+files = files([files.isdir] == 0);
+
+for ff = 1:length(files)
+    grp = regexp(files(ff).name, '[PMGC12346]*(?=.txt)', 'match');
+    groupsToTest = groupsToTest(~strcmp(groupsToTest, grp));
+end
+groupsLeft = length(groupsToTest);
+% randomly select one of the remaining groups to test this block
+group = groupsToTest{randi(groupsLeft)};
+
+% set up screen and basic textures
+SetUpScreen;
+
+% get list of stimuli for current group
+stimuli = stimuliToTest(group, params.imagesPerCondition);
+
+% output file
+fout = fopen([resultsFolder '/' group '.txt'], 'w');
[…]

### 2020-05-05 experiment/generate_stimuli.m

@@ -1,0 +1,3 @@
+addpath('../../../MATLAB/WPGeneratorDev/');
+
+generateWPTImages_multi({'P2', 'PM' ,'PG', 'CM', 'PMM', 'PMG', 'PGG', 'CMM', 'P4', 'P4M', 'P4G', 'P3', 'P3M1', 'P31M', 'P6', 'P6M'  }, 50, 100, '/Users/alasdairclarke/Documents/ongoing projects/SymmetryTimecourse/experiment/stimuli')

### 2020-05-05 experiment/process_results/1_init_proc.R

@@ -1,0 +1,76 @@
+library(tidyverse)
+library(forcats)
+library(rethinking)
+library(ggthemes)
+# library(lme4)
+# library(car)
+# library(scales)
+# --------------------------------------------------
+# read in data
+# remember that person 1,2 & 3 used a normal monitor
+# --------------------------------------------------
+
+people <- dir('../results/')
+
+wallpaper_groups <- c('P2', 'PM' ,'PG', 'CM', 'PMM', 'PMG', 'PGG', 'CMM', 'P3', 'P3M1', 'P31M', 'P4', 'P4M', 'P4G', 'P6', 'P6M')
+
+dat <- tibble(
+	person = factor(levels = people), 
+	group  = factor(levels = wallpaper_groups),
+	rep    = as.numeric(),
+	n      = as.numeric(),
+	t 	   = as.numeric(),
+	t1     = as.numeric(),
+	t2     = as.numeric(),
+	correct = as.numeric())
+
+thresholds <- tibble(
+	person = factor(levels = people), 
+	group  = factor(levels = wallpaper_groups),
+	staircase = as.numeric(),
+	threshold = as.numeric())
+
+for (person in people) {
+	
+	files <- dir(paste("../results/", person, sep="" ) , '*.txt')
+
+	for (group in files) {
+
+		# read in individual trial data
+		d <- read_csv(paste("../results/", person, "/", group, sep=""))
+		grp = unique(d$group)
+		dat <- rbind(dat, d)
+	}
+	
+	files <- dir(paste("../results/", person, "/thresholds/", sep="" ) , '*.txt')
+	
+	for (group in files) {
+
+		grp = strsplit(group, split = ".", fixed=T )[[1]][1]
+		# read in quest output
+		d <- read_csv(
+			paste("../results/", person, "/thresholds/", group, sep=""), 
+			col_names = FALSE)
+		thresholds <- bind_rows(thresholds, 
+			tibble(person = person, group = grp, staircase = 1:5, threshold = d$X1))
+
+		rm(d, grp)
+	}
+}
+
+rm(people, person, group)
+
+thresholds$person[thresholds$person=="person26"] <- "person0"
+
+# relabel staircase 5 to 0, - it is the 'joint staircase' containing data from all others
+thresholds$staircase[thresholds$staircase == 5] = 0
+
+# relevel so groups are in order of increasing complexity
+dat$group = as.factor(dat$group)
[…]

### 2020-05-05 experiment/scripts/LoadSymImages.m

@@ -1,0 +1,43 @@
+function [tex_stim1, tex_stim2, order] = LoadSymImages(stimuliScrn, stimuli, params)
+
+% 1 should be symmetrical,
+% 2 is control
+order = randperm(2);
+
+% select a random example to use
+ii = randi(params.imagesPerCondition);
+
+fname =  char(strcat('stimuli/', stimuli.sym(ii)));
+stimulusSym = ScaleContrastRotate(imread(fname), params);
+
+fname =  char(strcat('stimuli/', stimuli.ctl(ii)));
+stimulusCtl = ScaleContrastRotate(imread(fname), params);
+
+if order(1)==1
+    tex_stim1 = Screen('MakeTexture', stimuliScrn, stimulusSym);
+    tex_stim2 = Screen('MakeTexture', stimuliScrn, stimulusCtl);
+else
+    tex_stim1 = Screen('MakeTexture', stimuliScrn, stimulusCtl);
+    tex_stim2 = Screen('MakeTexture', stimuliScrn, stimulusSym);
+end
+
+
+end
+
+function im = ScaleContrastRotate(im, params)
+im  = im2double(im);
+
+% set mean to 0!
+im = params.stimuliContrast * (im - mean(im(:)));
+
+if params.rotate
+    im = imrotate(im, randi(360), 'bilinear');
+end
+
+% set to mean 0.5
+im = 255 * (im + 0.5);
+
+
+
+end
+

### 2020-05-05 experiment/scripts/RunPracticeTrials.m

@@ -1,0 +1,80 @@
+
+X = meshgrid(1:32)-round(32/2);
+Y = X';
+
+theta = targThetas(1);
+
+U = X*cos(theta) + Y*sin(theta);
+V = -X*sin(theta) + Y*cos(theta);
+
+lineSeg = zeros(32);
+lineSeg(abs(U)<5/2) = 1;
+lineSeg(abs(V)>25.6/2) = 0;
+
+
+lineSeg(lineSeg(:)==0) = 0.75;
+lineSeg(lineSeg(:)==1) = 0.2;
+
+lineSeg = 255*lineSeg;
+
+targTex = Screen('MakeTexture', stimuliScrn, lineSeg);
+Screen('DrawTexture', stimuliScrn, targTex);
+Screen('Close', targTex);
+DrawFormattedText(stimuliScrn, ...
+    ['Hello. First of all, we will carry out some practise trials \n' ...
+    'You will be loooking for the target shown:\n' ...
+    'Press f for target absent, j for target present. \n' ...
+    'Press any key when you are ready to start'], ...
+    display.midX-300, display.midY-200);
+Screen('flip', stimuliScrn);
+WaitSecs(1);
+KbWait;
+
+for trl = 1:2
+    trialInfo.target = 'present';
+    trialInfo.distracters = 'easy';
+    trialInfo.targTheta = targThetas(1);
+    % run a trial
+    [resp, rt, targLoc, targVar] = RunATrial(obs, 0, trl, ...
+        trialInfo, stimuliScrn, iLink, textures);
+end
+
+trl = 3;
+trialInfo.target = 'absent';
+trialInfo.distracters = 'easy';
+trialInfo.targTheta = targThetas(1);
+% run a trial
+[resp, rt, targLoc, targVar] = RunATrial(obs, 0, trl, ...
+    trialInfo, stimuliScrn, iLink, textures);
+trl = 4;
+trialInfo.target = 'absent';
+trialInfo.distracters = 'hard';
+trialInfo.targTheta = targThetas(1);
+% run a trial
+[resp, rt, targLoc, targVar] = RunATrial(obs, 0, trl, ...
+    trialInfo, stimuliScrn, iLink, textures);
+
+for trl = 5:6
+    trialInfo.target = 'present';
+    trialInfo.distracters = 'hard';
+    trialInfo.targTheta = targThetas(1);
+    % run a trial
[…]

### 2020-05-05 experiment/scripts/SetUpScreen.m

@@ -1,0 +1,47 @@
+%% set-up screen
+bkgrndGreyLevel = 127;
+N = 512;
+% do we want to remove this line eventually?
+Screen('Preference', 'SkipSyncTests', 1);
+
+stimuliScrn = Screen('OpenWindow',0, bkgrndGreyLevel );% , 
+[display.width, display.height]=Screen('WindowSize', stimuliScrn);
+display.midX = round(display.width/2);
+display.midY = round(display.height/2);
+
+% blank screen
+blank = bkgrndGreyLevel*ones(N);
+textures.blank = Screen('MakeTexture', stimuliScrn, blank);
+clear blank
+
+% fixation cross
+fixCross = makeFixCross(N, [bkgrndGreyLevel bkgrndGreyLevel bkgrndGreyLevel], N/2, N/2, 32);
+textures.fixCross = Screen('MakeTexture', stimuliScrn, fixCross);
+clear fixCross
+
+% set font size
+Screen('TextSize',stimuliScrn, 20);
+
+KbName('UnifyKeyNames')
+
+% load GetSecs and WaitSecs from memory
+GetSecs;
+WaitSecs(.01);
+
+noiseMask = 127 + 255 * (rand(150, 150)-0.5);
+noiseMask = imresize(noiseMask, 4, 'nearest');
+noiseMask(noiseMask<0) = 0;
+noiseMask(noiseMask>255) = 255;
+% create a cicrular mask
+x = repmat(1:600, [600, 1])-300; 
+d = x.^2 + x'.^2;
+d = d<=300^2;
+noiseMask = noiseMask .* d;
+noiseMask(noiseMask(:) == 0) = 127;
+
+
+textures.noiseMask = Screen('MakeTexture', stimuliScrn, noiseMask);
+
+clear noiseMask
+
+% HideCursor;

### 2020-05-05 experiment/scripts/StartNewBlock.m

@@ -1,0 +1,43 @@
+if blk > 1
+    
+    DrawFormattedText(stimuliScrn, ...
+        'End of block. Please take a break', ...
+        display.midX-300, display.midY-200);
+    Screen('flip', stimuliScrn);
+    WaitSecs(1);
+    KbWait;
+    EyelinkDoTrackerSetup(iLink.el);
+    
+end
+
+X = meshgrid(1:32)-round(32/2);
+Y = X';
+
+theta = trialList(blk, 1).targTheta;
+
+U = X*cos(theta) + Y*sin(theta);
+V = -X*sin(theta) + Y*cos(theta);
+
+lineSeg = zeros(32);
+lineSeg(abs(U)<5/2) = 1;
+lineSeg(abs(V)>25.6/2) = 0;
+
+
+lineSeg(lineSeg(:)==0) = 0.75;
+lineSeg(lineSeg(:)==1) = 0.2;
+
+lineSeg = 255*lineSeg;
+
+targTex = Screen('MakeTexture', stimuliScrn, lineSeg);
+Screen('DrawTexture', stimuliScrn, targTex);
+Screen('Close', targTex);
+DrawFormattedText(stimuliScrn, ...
+    ['Welcome to block ', int2str(blk), ' \n' ...
+    'You will be looking for the target shown:\n' ...
+    'Press f for target absent, j for target present. \n' ...
+    'Press any key when you are ready to start'], ...
+    display.midX-300, display.midY-200);
+Screen('flip', stimuliScrn);
+WaitSecs(1);
+KbWait;
+

### 2020-05-05 experiment/scripts/getObserverInput.m

@@ -1,0 +1,28 @@
+
+function [resp, responseKeyHit] = getObserverInput(c1, c2)
+responseKeyHit=0;
+while responseKeyHit==0
+    % check if a key has been pressed
+    [keyIsDown, ~, keyCode] = KbCheck;
+    % wait for keypress
+    while ~keyIsDown
+        [keyIsDown, ~, keyCode] = KbCheck;
+        WaitSecs(0.01);
+    end
+    % check what key was pressed
+    if find(keyCode) == KbName(c1);
+        responseKeyHit = 1;
+        resp = 1;
+    elseif find(keyCode) == KbName(c2);
+        responseKeyHit = 1;
+        resp = 2;
+    elseif  find(keyCode) == KbName('q');
+        responseKeyHit = 1;
+        resp = 3;
+    else
+        resp = 3;
+        responseKeyHit = 0;
+    end
+    WaitSecs(0.01);
+end
+end

### 2020-05-05 experiment/scripts/makeFixCross.m

@@ -1,0 +1,16 @@
+
+function fixCross = makeFixCross(N, colbk, x, y, w)
+if length(N)==2
+fixCross = ones(N(2), N(1),3);
+else 
+    fixCross = ones(N, N,3);
+end
+fixCross(:,:,1) = colbk(1);
+fixCross(:,:,2) = colbk(2);
+fixCross(:,:,3) = colbk(3);
+for i = 1:length(x)
+    fixCross(round(y(i)), (round(x(i))-w):(round(x(i))+w),:) = 0;
+    fixCross((round(y(i))-w):(round(y(i))+w), round(x(i)),:) = 0;
+end
+
+end

### 2020-05-05 experiment/scripts/stimuliToTest.m

@@ -1,0 +1,14 @@
+function stimuli = stimuliToTest(grp, stimuliPerCondition)
+
+keySet = {'P1', 'P2', 'PM' ,'PG', 'CM', 'PMM', 'PMG', 'PGG', 'CMM', 'P4', 'P4M', 'P4G', 'P3', 'P3M1', 'P31M', 'P6', 'P6M'};
+valueSet = 101:1:117;
+
+code = valueSet(strcmp(grp, keySet));
+stimuli.group = grp;
+for jj = 1:stimuliPerCondition
+    stimuli.sym{jj} =  sprintf('%d%03d.PNG', code, jj);
+    stimuli.ctl{jj} =  sprintf('%d%03d.PNG', code+17, jj);
+end
+
+end
+

### 2020-05-05 analysis/analysis.Rmd
Adding EEG data and analysis script

---
title: "Bayesian Modelling"
author: "Alasdair DF Clarke"
date: "21 October 2019"
output: 
  html_document:
    number_sections: true
---

I have set `echo=FALSE` so that most of the code chunks will not display. Please refer to the .Rmd file for source code.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)

# specify some colours to use in the plots
v_cols <- ggthemes::ptol_pal()(6)
```

# Loading packages and importing data

##  Packages

Details of packages, etc, are given below.

```{r packages, results='hold'}
library(tidyverse)
library(brms)
library(tidybayes)
library(ggridges)
library(latex2exp)
library(see)
library(ggthemes)
```

```{r}
sessionInfo()
```

Use multiple cores and set number of iterations. 
```{r}
options(mc.cores = parallel::detectCores())
n_iter <- 10000

# make empty lists to store things in
my_models <- list()
my_samples <- list()
comp_summary <- list()
subgroup_comp <- list()
```

## Import Data

Import data and quickly plot.

### EEG Data

Importing the eeg occ odd data.

```{r eeg_import, cache=TRUE}
fl <- "rms_Occ_Odd.csv"

# set column types
my_cols <- cols(
  .default = col_double(),
  Row = col_character()
)

#import
read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
  rename(wallpaper_group = "Row") %>%
  gather("subject", "rms", -wallpaper_group) %>%
  glimpse() -> d_eeg 
```

It's clearly skewed, and negative display duration are impossible, so will fit a `glm` with `family = 'lognormal`.

```{r}
d_eeg %>% ggplot(aes(x = rms)) + 
    geom_histogram(binwidth = 0.2, fill = "skyblue", colour = "black") +
    theme_lucid()
```

### Threshold Data

Here we import that data and select the columns that we're interested in. Threshold gives the required display duration (in seconds) for the two stimuli to allow for accurate discrimination.

```{r threshold_import, cache=TRUE}
fl <- "thresholds"

my_cols_threshold <- cols(
  person = col_character(),
  group = col_character(),
  staircase = col_double(),
  threshold = col_double(),
  log_threshold = col_double()
)

read_csv("../Experiment/analysis/thresholds.csv", col_types = my_cols_threshold) %>%
  select(subject = "person", wallpaper_group = "group", threshold) %>%
  glimpse() -> d_dispthresh

rm(my_cols_threshold)
```

As above, eyeball the data…

```{r}
d_dispthresh %>% ggplot(aes(x = threshold)) + 
    geom_histogram(binwidth = 0.2, fill = "skyblue", colour = "black") +
    theme_lucid()
```
Again, we have a skewed distribution, so will fit with `family = 'lognormal'`.

### Control Data

We will also include some control EEG data. These are the …. < details from Peter >

```{r import_control}

fl <- "rms_Occ_Even.csv"

read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
  rename(wallpaper_group = "Row") %>%
  gather("subject", "rms", -wallpaper_group) -> d_eeg_occ_even

fl <- "rms_Par_Odd.csv"

read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
  rename(wallpaper_group = "Row") %>%
  gather("subject", "rms", -wallpaper_group) -> d_eeg_par_odd

rm(my_cols, fl)
```

# Bayesian Analysis

Here are the details of the Bayesian multi-level modelling. Our general approach is:

*  Use weakly-informative priors.
*  Fit a multi-level model using the lognormal distribution.
  + independent variable (wallpaper group) and one dependent variable (rms/threshold).
  +maximal random effect structure.
* Check model fit (r̂ statistics and trace plots).
* Extract mcmc samples from the posterior distributions.
* Use these samples to estimate the difference between sub- and super-grounps.

## Define Priors

First we have to define some priors: are we happy with these?

We will keep the default priors for the group-level (‘random’) effects. From the brms documentation:

> [...] restricted to be non-negative and, by default, have a half student-t prior with 3 degrees of freedom and a scale parameter that depends on the standard deviation of the response after applying the link function. Minimally, the scale parameter is 10. This prior is used (a) to be only very weakly informative in order to influence results as few as possible, while (b) providing at least some regularization to considerably improve convergence and sampling efficiency.

Now we can fit a Bayesian multi-level model. First we will run with sample_prior = 'only' to check that our priors appear reasonable.

```{r}
## define some priors
prior <- c(
  set_prior("normal(0,2)", class = "b"),    
  set_prior("cauchy(0,2)", class = "sigma"))
```

Now we can fit a Bayesian multi-level model. First we will run with `sample_prior = 'only'` to check that our priors appear reasonable.

```{r eeg_prior_pred_check, cache=TRUE}
m_prior <- brm(data = d_eeg, 
  rms ~ wallpaper_group-1 + (1|subject),
  family = "lognormal", 
  prior = prior, 
  iter = n_iter ,
  sample_prior = 'only')
```

Now we will define some handy functions so we can reuse the same code to visualise the results of the model

```{r}
get_model_samples <- function(my_model) {

  my_samples <- posterior_samples(my_model)

  names(my_samples)  %>% 
    gsub('[b_]*wallpaper_group', "", .) -> names(my_samples)

  return(my_samples)
}

get_model_predictions <- function(my_samples) {
  my_samples %>%
    as_tibble() %>% 
    mutate_all(exp) %>%
    select(P2, PM, PG, CM, PMM, PMG, PGG, CMM, P4, P4M, P4G, P3, P3M1, P31M, P6, P6M) %>%
    gather() -> my_model_predictions

  return(my_model_predictions)
}

plot_model_output <- function(mdat, edat, my_title, my_colour, x_label) {
  x_var <- names(edat)[3]
  # plot model fit (mdat) against empiricaly data (edat)
  ggplot(mdat, aes(x = value, y = key)) +
      stat_pointintervalh(.width = c(.66, .95), colour = my_colour) +
      geom_density_ridges(data = edat,aes_string(x = x_var, "wallpaper_group"),
        fill = "grey",
        alpha = 0.25,
        scale = 2,
        bandwidth = 0.05) +
      scale_x_log10(x_label) +
    ggtitle(my_title) +
      theme_lucid()
}
```

Now we can use these functions to visualise the predictions made by the model:

```{r}

prior_samples <- get_model_samples(m_prior) %>% glimpse()

# Tidy up a little
prior_predictions <- get_model_predictions(prior_samples)%>% glimpse()

# # And now plot
plot_model_output(prior_predictions, d_eeg,
                   "Prior Predictive Check: EEG RMS", "deepskyblue3", "rms")

```

The density plots show the distribution of the data, while the point interval shows the median with $66\%$ and $95\%$ quantiles for our prior.

It looks like our priors are in the right range for the data.

### Prior Predictive Check for Display Duration Data

Now we can repeat this process (with the same priors) for the display duration threshold data.

```{r dispthresh_prior_pred_check, cache=TRUE, echo=FALSE}
m_prior <- brm(data = d_dispthresh, 
               threshold ~ wallpaper_group-1 + (1|subject),
               family = "lognormal", 
               prior = prior, iter = n_iter ,
               sample_prior = 'only')

# Extract samples from model
prior_samples <- get_model_samples(m_prior)

# Tidy up a little
prior_predictions <- get_model_predictions(prior_samples)

plot_model_output(prior_predictions, d_dispthresh, 
                  "Prior Predictive Check: Display Duration Thresholds", "deepskyblue3", "display threshold (ms)")
```

The density plots show the distribution of the data, while the point interval shows the median with $66\%$ and $95\%$ quantiles for our prior.

It looks like our priors are in the right range for the data.

## Compute Posterior

### Fit Model to EEG Data

We will now fit the model to the data.

```{r eeg_fit_model, cache=TRUE}
m_eeg <- brm(data = d_eeg, 
             rms ~ wallpaper_group-1 + (1|subject),
             family = "lognormal", 
             prior = prior, iter = n_iter,
             control = list(adapt_delta = 0.99))

summary(m_eeg)
```

Hopefully all the $\hat{r} = 1$ and there aren't any warning messages, etc.

We will now have at what the model predicts for the average participant (i.e, ignoring the random intercepts). If anybody else would like to suggest over sensible ways to summarise this model, do let me know! I've been trying to use some of the functions from the `tidybayes` package, and they don't (yet?) have much on random effects.

```{r}
 # Extract samples from model

my_samples$eeg <- get_model_samples(m_eeg)

 # Tidy up a little
 my_models$eeg <- get_model_predictions(my_samples$eeg)

 # Plot
 plot_model_output(my_models$eeg, d_eeg, 
                   "Model Fits: EEG RMS", "orangered3", "rms")
```

### Fit Model to Display Duration Threshold Data

We will now fit the model to the data.

```{r threshold_fit_model, cache=TRUE, echo=FALSE}
m_threshold <- brm(data = d_dispthresh, 
                   threshold ~ wallpaper_group-1 + (1|subject),
                   family = "lognormal", 
                   prior = prior, iter = n_iter,
                   control = list(adapt_delta = 0.99))

summary(m_threshold)
```

Hopefully all the $\hat{r} = 1$ and there aren't any warning messages, etc.

We will now have at what the model predicts for the average participant (i.e, ignoring the random intercepts). If anybody else would like to suggest over sensible ways to summarise this model, do let me know! I've been trying to use some of the functions from the `tidybayes` package, and they don't (yet?) have much on random effects.

```{r}
# Extract samples from model
my_samples$threshold <- get_model_samples(m_threshold)

# Tidy up a little
my_models$threshold <- get_model_predictions(my_samples$threshold)

# Plot
plot_model_output(my_models$threshold, d_dispthresh, 
                  "Model Fits: Display Duration Threshold", "orangered3", "threshold (ms)")
```

### EEG Control Data

We will also fit models to the blah and blah data as control conditions.

First for the occ_even data...

```{r occ_even_model, cache=TRUE}

m_eeg_occ_even <- brm(data = d_eeg_occ_even, 
                      rms ~ wallpaper_group-1 + (1|subject),
                      family = "lognormal", 
                      prior = prior, iter = n_iter,
                      control = list(adapt_delta = 0.99))

# Extract samples from model
my_samples$occ_even <- get_model_samples(m_eeg_occ_even)
```

And now for th par_odd data...

```{r par_odd_model, cache=TRUE}

m_eeg_par_odd <- brm(data = d_eeg_par_odd, 
                     rms ~ wallpaper_group-1 + (1|subject),
                     family = "lognormal", 
                     prior = prior, iter = n_iter,
                     control = list(adapt_delta = 0.99))

my_samples$par_odd <- get_model_samples(m_eeg_par_odd)
```

# Subgroup Comparisons

Helper function to do make comparisons and create plot:

```{r}
source("get_subgroup_comparisons.R")
```

```{r subgroup_comparisons, cache=TRUE}

subgroup_comp$eeg <- get_subgroup_comparisons(my_samples$eeg, 2*n_iter) %>% glimpse()
subgroup_comp$threshold <- get_subgroup_comparisons(my_samples$threshold, 2*n_iter)

# Also compute for control datasets
subgroup_comp$occ_even <- get_subgroup_comparisons(my_samples$occ_even, 2*n_iter) 
subgroup_comp$par_odd  <- get_subgroup_comparisons(my_samples$par_odd,  2*n_iter)

```

## EEG RMS

Finally, I calculate the probability that the difference between subgroup and supergroup is larger than zero (small than zero for the display duration) given the data. This information is then binned so we can colour in the posterior density plots.

First we will define a handy function that we can reuse a few times.

```{r define_comparison_summary_function, echo=TRUE}
comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
  # compute probabilties
  subgroup_comp_dat %>%
    group_by(key, index, normal) %>%
    summarise(
      mean_value = mean(value),
      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)))  %>%
    mutate(
      measurement = measurement_lab,
      index_labs = as.numeric(as.factor(index)),
      lab_cols = v_cols[index_labs],
      prob_cat = cut(prob_one_tailed, c(0, 0.5, 0.75, 0.9, 0.99, 1))) -> comp_summary_dat

  return(comp_summary_dat)
}
```

Now we will use it!

```{r, plot_eeg_comp, fig.height=8}
# compute probabilties
comp_summary$eeg <- comparison_summary(subgroup_comp$eeg, TRUE, 'rms') %>% glimpse()

subgroup_comp$eeg <- full_join(
  subgroup_comp$eeg, 
  comp_summary$eeg, 
  by = c("key", "index", "normal"))

# Finally, plot everything:
p1 <- plot_comparisons(
  filter(subgroup_comp$eeg, index == 2),
  filter(comp_summary$eeg, index == 2), 
  1)

p2 <-plot_comparisons(
  filter(subgroup_comp$eeg, index != 2),
  filter(comp_summary$eeg, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts
```

## Display Duration Threshold

We can do the same for the display duration thresholds.

```{r plot_threshold_comp, fig.height=8}

# compute probabilties
comp_summary$threshold <- comparison_summary(subgroup_comp$threshold, FALSE, 'threshold') %>% glimpse()

subgroup_comp$threshold <- full_join(
  subgroup_comp$threshold, 
  comp_summary$threshold, 
  by = c("key", "index", "normal"))

# Finally, plot everything:
p1 <- plot_comparisons(
  filter(subgroup_comp$threshold, index == 2),
  filter(comp_summary$threshold, index == 2), 
  1, TRUE)
p2 <- plot_comparisons(
  filter(subgroup_comp$threshold, index != 2),
  filter(comp_summary$threshold, index != 2), 
  2, TRUE)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts
```

## Control Data

We will now do exactly the same with the control data...

```{r comp_control, fig.height=8, echo = FALSE}

# compute probabilties
comp_summary$occ_even <- comparison_summary(subgroup_comp$occ_even, TRUE, 'rms')

subgroup_comp$occ_even <- full_join(
  subgroup_comp$occ_even, 
  comp_summary$occ_even, 
  by = c("key", "index", "normal"))

p1 <- plot_comparisons(
  filter(subgroup_comp$occ_even, index == 2),
  filter(comp_summary$occ_even, index == 2), 
  1)
p2 <- plot_comparisons(
  filter(subgroup_comp$occ_even, index != 2),
  filter(comp_summary$occ_even, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
both_plts

# compute probabilties
comp_summary$par_odd <- comparison_summary(subgroup_comp$par_odd, TRUE, 'rms')

subgroup_comp$par_odd <- full_join(
  subgroup_comp$par_odd, 
  comp_summary$par_odd, 
  by = c("key", "index", "normal"))

p1 <- plot_comparisons(
  filter(subgroup_comp$par_odd, index == 2),
  filter(comp_summary$par_odd, index == 2), 
  1)
p2 <- plot_comparisons(
  filter(subgroup_comp$par_odd, index != 2),
  filter(comp_summary$par_odd, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts

```

## Summary

Some summary stats for how many subgroups show a difference. We have 64 subgroupds in total.

```{r}

count_comparisons <- function(comparisons, p) {
  x <- sum(comparisons$prob_one_tailed >= p)
  return(x)
}

df <- tibble()

for (p in seq(0.5, 0.99, 0.01)) {

  df <- bind_rows(df, map_dbl(comp_summary,  count_comparisons, p))
}

df$p <- seq(0.5, 0.99, 0.01)

df %>% gather("measurement", "number", -p) %>%
  mutate(proportion = number/64) %>%
  ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
  theme_minimal() + ggthemes::scale_color_ptol() +
  xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")

filter(df, p == 0.95)

```

# Additional Analysis

## Index and Normality

## Correlation Between EEG RMS and Display Duration Threshold

```{r compare_comparisons}

bind_rows(
  select(comp_summary$eeg, measurement, key, mean_value),
  select(comp_summary$threshold, measurement, key, mean_value)) %>%
  spread(measurement, mean_value) %>%
  glimpse()-> d

m <- brm(data = d, threshold ~ rms)

summary(m)

bayes_R2(m)

d %>% 
  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
  add_fitted_draws(m, n = 100) %>%
  ggplot(aes(x = rms, y = threshold, color = index)) +
  geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
  geom_point(data = d) +
  theme_minimal() + 
  ggthemes::scale_colour_ptol()

```

### 2020-05-05 analysis/scripts/get_subgroup_comparisons.R

@@ -1,0 +1,228 @@
+get_subgroup_comparisons <- function(m_samples, n_iter)
+{
+    m_samples %>%
+      as_tibble() %>% 
+      mutate_all(exp) %>%
+      mutate(
+        'P4M$\\rightarrow$ P4G' = P4M - P4G,
+        'P4M$\\rightarrow$ P4'  = P4M - P4,
+        'P4M$\\rightarrow$ CMM' = P4M - CMM,
+        'P4M$\\rightarrow$ PMM' = P4M - PMM, 
+        'P6M$\\rightarrow$ P6' = P6M - P6,
+        'P6M$\\rightarrow$ P31M' = P6M - P31M,
+        'P6M$\\rightarrow$ P3M1' = P6M - P3M1,
+        'PMM$\\rightarrow$ PMG' = PMM - PMG,
+        'PMM$\\rightarrow$ PM' = PMM - PM,
+        'PMM$\\rightarrow$ P2' = PMM - P2,
+        'P3M1$\\rightarrow$ P3'= P3M1- P3,
+        'CMM$\\rightarrow$ PMG' = CMM - PMG,
+        'CMM$\\rightarrow$ PGG' = CMM - PGG,
+        'CMM$\\rightarrow$ CM' = CMM - CM,
+        'CMM$\\rightarrow$ P2' = CMM - P2,
+        'P4G$\\rightarrow$ P4' = P4G - P4,
+        'P4G$\\rightarrow$ CMM' = P4G - CMM,
+        'P4G$\\rightarrow$ PGG' = P4G - PGG,
+        'PMG$\\rightarrow$ PGG' = PMG - PGG, 
+        'PMG$\\rightarrow$ PM' = PMG - PM,
+        'PMG$\\rightarrow$ PG' = PMG - PG,
+        'PMG$\\rightarrow$ P2' = PMG - P2,
+        'P31M$\\rightarrow$ P3'= P31M- P3,
+        'P6$\\rightarrow$ P3'  = P6  - P3,
+        'CM$\\rightarrow$ PG'  = CM  - PG,
+        'P4$\\rightarrow$ P2'  = P4  - P2,
+        'PM$\\rightarrow$ PG'  = PM  - PG, 
+        'PGG$\\rightarrow$ PG' = PGG - PG, 
+        'PGG$\\rightarrow$ P2' = PGG - P2, 
+        'P2$\\rightarrow$ PG'  = P2 - PG,
[…]

### 2020-05-05 analysis/analysis.Rmd
Sup mat new version up to PPC

@@
 ---
-title: "Bayesian Modelling"
-author: "Alasdair DF Clarke"
-date: "21 October 2019"
-output: 
-  html_document:
-    number_sections: true
+title: "Supplementary Materials"
+author: "Peter J. Kohler, Oda Åsgard, and Alasdair D. F. Clarke"
+date:
+output: bookdown::html_document2
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE)
+
+library(bookdown)
+library(tidyverse)
+library(brms)
+library(tidybayes)
+library(ggridges)
+library(latex2exp)
+library(see)
+library(ggthemes)
+library(patchwork)
+
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, fig.align='center')
 
@@
 v_cols <- ggthemes::ptol_pal()(6)
+
+# read in some helper functions that will be used later
+source("scripts/helper_functions.R")
 ```
@@
 Details of packages, etc, are given below.
-
-```{r packages, results='hold'}
-library(tidyverse)
-library(brms)
-library(tidybayes)
-library(ggridges)
-library(latex2exp)
-library(see)
-library(ggthemes)
-```
 
@@
 
-Import data and quickly plot.
+Here we import our data and make some sumamry plots.
 
@@
 
-Importing the eeg occ odd data.
+Importing the odd-harmonic EEG data from six electrodes region-of-interest over occipital cortex.
 
-```{r eeg_import, cache=TRUE}
+```{r eeg_import}
 fl <- "rms_Occ_Odd.csv"
@@
 
-```{r}
+```{r cache=TRUE, fig.cap = "Data are the root-mean-squaured (rms) over the odd-harmonic filtered waveforms."}
 d_eeg %>% ggplot(aes(x = rms)) + 
@@
 
-```{r threshold_import, cache=TRUE}
+```{r threshold_import}
 fl <- "thresholds"
@@
 
-read_csv("../Experiment/analysis/thresholds.csv", col_types = my_cols_threshold) %>%
+read_csv("../Experiment/process_results/thresholds.csv", col_types = my_cols_threshold) %>%
   select(subject = "person", wallpaper_group = "group", threshold) %>%
@@
 
-As above, eyeball the data…
+As above, a summary of the data.
 
-```{r}
+```{r, fig.cap="Histogram of the display duration thresholds."}
 d_dispthresh %>% ggplot(aes(x = threshold)) + 
@@
 ```
+
 Again, we have a skewed distribution, so will fit with `family = 'lognormal'`.
@@
 
-We will also include some control EEG data. These are the …. < details from Peter >
+In addition to the primary EEG data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data 
+from six parietal electrodes (see main paper).
 
@@
 *  Fit a multi-level model using the lognormal distribution.
-  + independent variable (wallpaper group) and one dependent variable (rms/threshold).
-  +maximal random effect structure.
+  + independent variable (wallpaper group) and one dependent variable (rms or threshold).
+  + maximal random effect structure.
 * Check model fit (r̂ statistics and trace plots).
 * Extract mcmc samples from the posterior distributions.
-* Use these samples to estimate the difference between sub- and super-grounps.
+* Use these samples to estimate the difference between sub- and super-groups.
 
@@
 
-First we have to define some priors: are we happy with these?
+In this section we will specify some priors. We then then use a prior-predictive check to assess whether the prior is reasonable or not (i.e., on the same order of magnitude as our measurements).
 
-We will keep the default priors for the group-level (‘random’) effects. From the brms documentation:
+### Fixed Effects
+
+Our independent variable is a categorical factor with 16 levels.
+We will drop the intercept from our model and instead fit a coefficent for each factor level ($y \sim x - 0$).
+As our dependant variable will be log-transformed, we can use the priors below:
+
+```{r, echo = TRUE}
+prior <- c(
+  set_prior("normal(0,2)", class = "b"),    
+  set_prior("cauchy(0,2)", class = "sigma"))
+```
+
+### Group-level Effects
+
+We will keep the default weakly informative priors for the group-level (‘random’) effects. From the brms documentation:
 
@@
 
-Now we can fit a Bayesian multi-level model. First we will run with sample_prior = 'only' to check that our priors appear reasonable.
+### Prior Predictive Check
 
-```{r}
-## define some priors
-prior <- c(
-  set_prior("normal(0,2)", class = "b"),    
-  set_prior("cauchy(0,2)", class = "sigma"))
-```
+Now we can specify our Bayesian multi-level model and priors. Note that as we are using  `sample_prior = 'only'`, the model will not learn anything from our data.
 
-Now we can fit a Bayesian multi-level model. First we will run with `sample_prior = 'only'` to check that our priors appear reasonable.
-
-```{r eeg_prior_pred_check, cache=TRUE}
+```{r eeg_prior_pred_check, cache=TRUE, echo=TRUE}
 m_prior <- brm(data = d_eeg, 
@@
 
-Now we will define some handy functions so we can reuse the same code to visualise the results of the model
+We can use this model to generate data.
 
-```{r}
-get_model_samples <- function(my_model) {
-
-  my_samples <- posterior_samples(my_model)
-
-  names(my_samples)  %>% 
-    gsub('[b_]*wallpaper_group', "", .) -> names(my_samples)
-
-  return(my_samples)
-}
-
-get_model_predictions <- function(my_samples) {
-  my_samples %>%
-    as_tibble() %>% 
-    mutate_all(exp) %>%
-    select(P2, PM, PG, CM, PMM, PMG, PGG, CMM, P4, P4M, P4G, P3, P3M1, P31M, P6, P6M) %>%
-    gather() -> my_model_predictions
-
-  return(my_model_predictions)
-}
-
-plot_model_output <- function(mdat, edat, my_title, my_colour, x_label) {
-  x_var <- names(edat)[3]
-  # plot model fit (mdat) against empiricaly data (edat)
-  ggplot(mdat, aes(x = value, y = key)) +
-      stat_pointintervalh(.width = c(.66, .95), colour = my_colour) +
-      geom_density_ridges(data = edat,aes_string(x = x_var, "wallpaper_group"),
-        fill = "grey",
-        alpha = 0.25,
-        scale = 2,
-        bandwidth = 0.05) +
-      scale_x_log10(x_label) +
-    ggtitle(my_title) +
-      theme_lucid()
-}
-```
-
-Now we can use these functions to visualise the predictions made by the model:
-
-```{r}
-
-prior_samples <- get_model_samples(m_prior) %>% glimpse()
+```{r, fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
+prior_samples <- get_model_samples(m_prior)
 
@@
 # # And now plot
-plot_model_output(prior_predictions, d_eeg,
-                   "Prior Predictive Check: EEG RMS", "deepskyblue3", "rms")
+p1 <- plot_model_output(prior_predictions, d_eeg,
+                   "odd-harmonic data from \noccipital electrodes", "cornflowerblue", "log(rms)")
 
+p2 <- plot_model_output(prior_predictions, d_dispthresh,
+                   "display duration threshold data", "cornflowerblue", "log(ms)")
+
+# draw pot
+plt <- p1 + p2
+plt + plot_annotation(title = 'Prior Predictive Checks')
 ```
 
-The density plots show the distribution of the data, while the point interval shows the median with $66\%$ and $95\%$ quantiles for our prior.
-
-It looks like our priors are in the right range for the data.
-
-### Prior Predictive Check for Display Duration Data
-
-Now we can repeat this process (with the same priors) for the display duration threshold data.
-
-```{r dispthresh_prior_pred_check, cache=TRUE, echo=FALSE}
-m_prior <- brm(data = d_dispthresh, 
-               threshold ~ wallpaper_group-1 + (1|subject),
-               family = "lognormal", 
-               prior = prior, iter = n_iter ,
-               sample_prior = 'only')
-
-# Extract samples from model
-prior_samples <- get_model_samples(m_prior)
-
-# Tidy up a little
-prior_predictions <- get_model_predictions(prior_samples)
-
-plot_model_output(prior_predictions, d_dispthresh, 
-                  "Prior Predictive Check: Display Duration Thresholds", "deepskyblue3", "display threshold (ms)")
-```
-
-The density plots show the distribution of the data, while the point interval shows the median with $66\%$ and $95\%$ quantiles for our prior.
-
-It looks like our priors are in the right range for the data.
-
-## Compute Posterior
-
-### Fit Model to EEG Data
-
-We will now fit the model to the data.
-
-```{r eeg_fit_model, cache=TRUE}
-m_eeg <- brm(data = d_eeg, 
-             rms ~ wallpaper_group-1 + (1|subject),
-             family = "lognormal", 
-             prior = prior, iter = n_iter,
-             control = list(adapt_delta = 0.99))
-
-summary(m_eeg)
-```
-
-Hopefully all the $\hat{r} = 1$ and there aren't any warning messages, etc.
-
-We will now have at what the model predicts for the average participant (i.e, ignoring the random intercepts).
-If anybody else would like to suggest over sensible ways to summarise this model, do let me know!
-I've been trying to use some of the functions from the `tidybayes` package, and they don't (yet?) have much on random effects.
-
-```{r}
- # Extract samples from model
-
-my_samples$eeg <- get_model_samples(m_eeg)
-
- # Tidy up a little
- my_models$eeg <- get_model_predictions(my_samples$eeg)
-
- # Plot
- plot_model_output(my_models$eeg, d_eeg, 
-                   "Model Fits: EEG RMS", "orangered3", "rms")
-```
-
-### Fit Model to Display Duration Threshold Data
-
-We will now fit the model to the data.
-
-```{r threshold_fit_model, cache=TRUE, echo=FALSE}
-m_threshold <- brm(data = d_dispthresh, 
-                   threshold ~ wallpaper_group-1 + (1|subject),
-                   family = "lognormal", 
-                   prior = prior, iter = n_iter,
-                   control = list(adapt_delta = 0.99))
-
-summary(m_threshold)
-```
-
-Hopefully all the $\hat{r} = 1$ and there aren't any warning messages, etc.
-
-We will now have at what the model predicts for the average participant (i.e, ignoring the random intercepts).
-If anybody else would like to suggest over sensible ways to summarise this model, do let me know!
-I've been trying to use some of the functions from the `tidybayes` package, and they don't (yet?) have much on random effects.
-
-```{r}
-# Extract samples from model
-my_samples$threshold <- get_model_samples(m_threshold)
-
-# Tidy up a little
-my_models$threshold <- get_model_predictions(my_samples$threshold)
-
-# Plot
-plot_model_output(my_models$threshold, d_dispthresh, 
-                  "Model Fits: Display Duration Threshold", "orangered3", "threshold (ms)")
-```
-
-### EEG Control Data
-
-We will also fit models to the blah and blah data as control conditions.
-
-First for the occ_even data...
-
-```{r occ_even_model, cache=TRUE}
-
-m_eeg_occ_even <- brm(data = d_eeg_occ_even, 
-                      rms ~ wallpaper_group-1 + (1|subject),
-                      family = "lognormal", 
-                      prior = prior, iter = n_iter,
-                      control = list(adapt_delta = 0.99))
-
-# Extract samples from model
-my_samples$occ_even <- get_model_samples(m_eeg_occ_even)
-```
-
-And now for th par_odd data...
-
-```{r par_odd_model, cache=TRUE}
-
-m_eeg_par_odd <- brm(data = d_eeg_par_odd, 
-                     rms ~ wallpaper_group-1 + (1|subject),
-                     family = "lognormal", 
-                     prior = prior, iter = n_iter,
-                     control = list(adapt_delta = 0.99))
-
-my_samples$par_odd <- get_model_samples(m_eeg_par_odd)
-```
-
-# Subgroup Comparisons
-
-Helper function to do make comparisons and create plot:
-
-```{r}
-source("get_subgroup_comparisons.R")
-```
-
-```{r subgroup_comparisons, cache=TRUE}
-
-subgroup_comp$eeg <- get_subgroup_comparisons(my_samples$eeg, 2*n_iter) %>% glimpse()
-subgroup_comp$threshold <- get_subgroup_comparisons(my_samples$threshold, 2*n_iter)
-
-# Also compute for control datasets
-subgroup_comp$occ_even <- get_subgroup_comparisons(my_samples$occ_even, 2*n_iter) 
-subgroup_comp$par_odd  <- get_subgroup_comparisons(my_samples$par_odd,  2*n_iter)
-
-```
-
-## EEG RMS
-
-Finally, I calculate the probability that the difference between subgroup and supergroup is larger than zero (small than zero for the display duration) given the data.
-This information is then binned so we can colour in the posterior density plots.
-
-First we will define a handy function that we can reuse a few times.
-
-```{r define_comparison_summary_function, echo=TRUE}
-comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
-  # compute probabilties
-  subgroup_comp_dat %>%
-    group_by(key, index, normal) %>%
-    summarise(
-      mean_value = mean(value),
-      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)))  %>%
-    mutate(
-      measurement = measurement_lab,
-      index_labs = as.numeric(as.factor(index)),
-      lab_cols = v_cols[index_labs],
-      prob_cat = cut(prob_one_tailed, c(0, 0.5, 0.75, 0.9, 0.99, 1))) -> comp_summary_dat
-
-  return(comp_summary_dat)
-}
-```
-
-Now we will use it!
-
-```{r, plot_eeg_comp, fig.height=8}
-# compute probabilties
-comp_summary$eeg <- comparison_summary(subgroup_comp$eeg, TRUE, 'rms') %>% glimpse()
-
-subgroup_comp$eeg <- full_join(
-  subgroup_comp$eeg, 
-  comp_summary$eeg, 
-  by = c("key", "index", "normal"))
-
-# Finally, plot everything:
-p1 <- plot_comparisons(
-  filter(subgroup_comp$eeg, index == 2),
-  filter(comp_summary$eeg, index == 2), 
-  1)
-
-p2 <-plot_comparisons(
-  filter(subgroup_comp$eeg, index != 2),
-  filter(comp_summary$eeg, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts
-```
-
-## Display Duration Threshold
-
-We can do the same for the display duration thresholds.
-
-```{r plot_threshold_comp, fig.height=8}
-
-# compute probabilties
-comp_summary$threshold <- comparison_summary(subgroup_comp$threshold, FALSE, 'threshold') %>% glimpse()
-
-subgroup_comp$threshold <- full_join(
-  subgroup_comp$threshold, 
-  comp_summary$threshold, 
-  by = c("key", "index", "normal"))
-
-# Finally, plot everything:
-p1 <- plot_comparisons(
-  filter(subgroup_comp$threshold, index == 2),
-  filter(comp_summary$threshold, index == 2), 
-  1, TRUE)
-p2 <- plot_comparisons(
-  filter(subgroup_comp$threshold, index != 2),
-  filter(comp_summary$threshold, index != 2), 
-  2, TRUE)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts
-```
-
-## Control Data
-
-We will now do exactly the same with the control data...
-
-```{r comp_control, fig.height=8, echo = FALSE}
-
-# compute probabilties
-comp_summary$occ_even <- comparison_summary(subgroup_comp$occ_even, TRUE, 'rms')
-
-subgroup_comp$occ_even <- full_join(
-  subgroup_comp$occ_even, 
-  comp_summary$occ_even, 
-  by = c("key", "index", "normal"))
-
-p1 <- plot_comparisons(
-  filter(subgroup_comp$occ_even, index == 2),
-  filter(comp_summary$occ_even, index == 2), 
-  1)
-p2 <- plot_comparisons(
-  filter(subgroup_comp$occ_even, index != 2),
-  filter(comp_summary$occ_even, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
-both_plts
-
-# compute probabilties
-comp_summary$par_odd <- comparison_summary(subgroup_comp$par_odd, TRUE, 'rms')
-
-subgroup_comp$par_odd <- full_join(
-  subgroup_comp$par_odd, 
-  comp_summary$par_odd, 
-  by = c("key", "index", "normal"))
-
-p1 <- plot_comparisons(
-  filter(subgroup_comp$par_odd, index == 2),
-  filter(comp_summary$par_odd, index == 2), 
-  1)
-p2 <- plot_comparisons(
-  filter(subgroup_comp$par_odd, index != 2),
-  filter(comp_summary$par_odd, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts
-
-```
-
-## Summary
-
-Some summary stats for how many subgroups show a difference. We have 64 subgroupds in total.
-
-```{r}
-
-count_comparisons <- function(comparisons, p) {
-  x <- sum(comparisons$prob_one_tailed >= p)
-  return(x)
-}
-
-df <- tibble()
-
-for (p in seq(0.5, 0.99, 0.01)) {
-
-  df <- bind_rows(df, map_dbl(comp_summary,  count_comparisons, p))
-}
-
-df$p <- seq(0.5, 0.99, 0.01)
-
-df %>% gather("measurement", "number", -p) %>%
-  mutate(proportion = number/64) %>%
-  ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
-  theme_minimal() + ggthemes::scale_color_ptol() +
-  xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")
-
-filter(df, p == 0.95)
-
-```
-
-# Additional Analysis
-
-## Index and Normality
-
-## Correlation Between EEG RMS and Display Duration Threshold
-
-```{r compare_comparisons}
-
-bind_rows(
-  select(comp_summary$eeg, measurement, key, mean_value),
-  select(comp_summary$threshold, measurement, key, mean_value)) %>%
-  spread(measurement, mean_value) %>%
-  glimpse()-> d
-
-m <- brm(data = d, threshold ~ rms)
-
-summary(m)
-
-bayes_R2(m)
-
-d %>% 
-  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
-  add_fitted_draws(m, n = 100) %>%
-  ggplot(aes(x = rms, y = threshold, color = index)) +
-  geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
-  geom_point(data = d) +
-  theme_minimal() + 
-  ggthemes::scale_colour_ptol()
-
-```
+We can see that i) our priors are relatively weak as the predictions span several orders of magnitide, and ii) our empirical data falls within this range.

### 2020-05-05 analysis/scripts/helper_functions.R

@@ -1,0 +1,37 @@
+get_model_samples <- function(my_model) {
+  
+  my_samples <- posterior_samples(my_model)
+  
+  names(my_samples)  %>% 
+    gsub('[b_]*wallpaper_group', "", .) -> names(my_samples)
+  
+  return(my_samples)
+}
+
+get_model_predictions <- function(my_samples) {
+  my_samples %>%
+    as_tibble() %>% 
+    mutate_all(exp) %>%
+    select(P2, PM, PG, CM, PMM, PMG, PGG, CMM, P4, P4M, P4G, P3, P3M1, P31M, P6, P6M) %>%
+    gather() -> my_model_predictions
+  
+  return(my_model_predictions)
+}
+
+plot_model_output <- function(mdat, edat, my_title, my_colour, x_label) {
+  x_var <- names(edat)[3]
+  # plot model fit (mdat) against empiricaly data (edat)
+  plt <- ggplot(mdat, aes(x = value, y = key)) +
+    stat_pointintervalh(.width = c(.66, .95), colour = my_colour) +
+    geom_density_ridges(data = edat,aes_string(x = x_var, "wallpaper_group"),
+                        fill = "grey",
+                        alpha = 0.25,
+                        scale = 2,
+                        bandwidth = 0.05) +
+    scale_x_log10(x_label) +
+    scale_y_discrete("wallpaper group") +
+    ggtitle(my_title) +
+    theme_lucid()
+  
+  return(plt)
+}

### 2020-05-05 analysis/analysis.Rmd
Running models

@@
 
+\tiny
 ```{r eeg_import}
@@
 read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
-  rename(wallpaper_group = "Row") %>%
-  gather("subject", "rms", -wallpaper_group) %>%
+  rename(wg = "Row") %>%
+  gather("subject", "rms", -wg) %>%
   glimpse() -> d_eeg 
@@
 read_csv("../Experiment/process_results/thresholds.csv", col_types = my_cols_threshold) %>%
-  select(subject = "person", wallpaper_group = "group", threshold) %>%
+  select(subject = "person", wg = "group", threshold) %>%
   glimpse() -> d_dispthresh
@@
 read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
-  rename(wallpaper_group = "Row") %>%
-  gather("subject", "rms", -wallpaper_group) -> d_eeg_occ_even
+  rename(wg = "Row") %>%
+  gather("subject", "rms", -wg) -> d_eeg_occ_even
 
@@
 read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
-  rename(wallpaper_group = "Row") %>%
-  gather("subject", "rms", -wallpaper_group) -> d_eeg_par_odd
+  rename(wg = "Row") %>%
+  gather("subject", "rms", -wg) -> d_eeg_par_odd
 
@@
 m_prior <- brm(data = d_eeg, 
-  rms ~ wallpaper_group-1 + (1|subject),
+  rms ~ wg-1 + (1|subject),
   family = "lognormal", 
@@
 We can see that i) our priors are relatively weak as the predictions span several orders of magnitide, and ii) our empirical data falls within this range.
+
+## Compute Posterior
+
+### Fit Model to EEG Data
+
+We will now fit the model to the data.
+
+```{r eeg_fit_model, cache=TRUE}
+m_eeg <- brm(data = d_eeg, 
+             rms ~ wg-1 + (1|subject),
+             family = "lognormal", 
+             prior = prior, iter = n_iter,
+             control = list(adapt_delta = 0.99))
+
+summary(m_eeg)
+```
+
+We will now have at what the model predicts for the average participant (i.e, ignoring the random intercepts).
+If anybody else would like to suggest over sensible ways to summarise this model, do let me know!
+I've been trying to use some of the functions from the `tidybayes` package, and they don't (yet?) have much on random effects.
+
+```{r, fig.cap ="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
+
+ # Extract samples from model
+my_samples$eeg <- get_model_samples(m_eeg)
+
+ # Tidy up a little
+ my_models$eeg <- get_model_predictions(my_samples$eeg)
+
+ # Plot
+plot_model_output(
+  my_models$eeg, d_eeg,
+  "Model Fits for the odd-harmonic data from occipital electrodes",
+  "orangered3", 
+  "log(rms)")
+```
+
+### Fit Model to Display Duration Threshold Data
+
+We will now fit the model to the data.
+
+```{r threshold_fit_model, cache=TRUE, echo=FALSE}
+m_threshold <- brm(data = d_dispthresh, 
+                   threshold ~ wg-1 + (1|subject),
+                   family = "lognormal", 
+                   prior = prior, iter = n_iter,
+                   control = list(adapt_delta = 0.99))
+
+summary(m_threshold)
+```
+
+```{r, fig.cap ="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
+# Extract samples from model
+my_samples$threshold <- get_model_samples(m_threshold)
+
+# Tidy up a little
+my_models$threshold <- get_model_predictions(my_samples$threshold)
+
+# Plot
+plot_model_output(my_models$threshold, d_dispthresh, 
+                  "Model Fits: Display Duration Threshold", "orangered3", "threshold (ms)")
+```

### 2020-05-05 analysis/scripts/helper_functions.R

@@ -3,7 +3,7 @@ get_model_samples <- function(my_model) {
   my_samples <- posterior_samples(my_model)
   
   names(my_samples)  %>% 
-    gsub('[b_]*wallpaper_group', "", .) -> names(my_samples)
+    gsub('[b_]*wg', "", .) -> names(my_samples)
   
   return(my_samples)
 }
@@ -23,7 +23,7 @@ plot_model_output <- function(mdat, edat, my_title, my_colour, x_label) {
   # plot model fit (mdat) against empiricaly data (edat)
   plt <- ggplot(mdat, aes(x = value, y = key)) +
     stat_pointintervalh(.width = c(.66, .95), colour = my_colour) +
-    geom_density_ridges(data = edat,aes_string(x = x_var, "wallpaper_group"),
+    geom_density_ridges(data = edat,aes_string(x = x_var, "wg"),
                         fill = "grey",
                         alpha = 0.25,
                         scale = 2,

### 2020-05-07 analysis/analysis.Rmd
Subgroup comparisons

@@
 ```
+
+### EEG Control Data
+
+We will also fit models to the control data. As we can see from Figure 2.4, the group differences are much smaller.
+
+```{r cache=TRUE, fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
+
+m_eeg_par_odd <- brm(data = d_eeg_par_odd, 
+                     rms ~ wg-1 + (1|subject),
+                     family = "lognormal", 
+                     prior = prior, iter = n_iter,
+                     control = list(adapt_delta = 0.99))
+
+my_samples$par_odd <- get_model_samples(m_eeg_par_odd)
+my_models$par_odd <- get_model_predictions(my_samples$par_odd)
+plt1 <- plot_model_output(my_models$par_odd, d_eeg_par_odd, 
+                  "odd-harmonic data from \nparietal electrodes", "orangered3", "threshold (ms)")
+
+m_eeg_occ_even <- brm(data = d_eeg_occ_even, 
+                      rms ~ wg-1 + (1|subject),
+                      family = "lognormal", 
+                      prior = prior, iter = n_iter,
+                      control = list(adapt_delta = 0.99))
+
+# Extract samples from model
+my_samples$occ_even <- get_model_samples(m_eeg_occ_even)
+my_models$occ_even <- get_model_predictions(my_samples$occ_even)
+plt2 <- plot_model_output(my_models$occ_even, d_eeg_occ_even, 
+                  "even-harmonic data from \noccipital electrodes", "orangered3", "threshold (ms)")
+
+plt1 + plt2 + plot_annotation(title = 'Model Fits for the ')
+```
+
+# Subgroup Comparisons
+
+We will now compute the difference between sub- and super-groups.
+
+```{r subgroup_comparisons, cache=TRUE}
+source("scripts/get_subgroup_comparisons.R")
+
+subgroup_comp$eeg <- get_subgroup_comparisons(my_samples$eeg, 2*n_iter)
+subgroup_comp$threshold <- get_subgroup_comparisons(my_samples$threshold, 2*n_iter)
+subgroup_comp$occ_even <- get_subgroup_comparisons(my_samples$occ_even, 2*n_iter)
+subgroup_comp$par_odd  <- get_subgroup_comparisons(my_samples$par_odd,  2*n_iter)
+```
+
+## EEG RMS
+
+Finally, I calculate the probability that the difference between subgroup and supergroup is larger than zero (small than zero for the display duration) given the data.
+This information is then binned so we can colour in the posterior density plots.
+
+```{r define_comparison_summary_function}
+comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
+  # compute probabilties
+  subgroup_comp_dat %>%
+    group_by(key, index, normal) %>%
+    summarise(
+      mean_value = mean(value),
+      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)))  %>%
+    mutate(
+      measurement = measurement_lab,
+      index_labs = as.numeric(as.factor(index)),
+      lab_cols = v_cols[index_labs],
+      prob_cat = cut(prob_one_tailed, c(0, 0.5, 0.75, 0.9, 0.99, 1))) -> comp_summary_dat
+
+  return(comp_summary_dat)
+}
+```
+
+Now we will use it!
+
+```{r, fig.height=8, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
+The index of each relationship is indicated by the colour of the y-axis label.
+The fill of the density plots indicated the probability of the difference being greater than zero."}
+# compute probabilties
+comp_summary$eeg <- comparison_summary(subgroup_comp$eeg, TRUE, 'rms')
+
+subgroup_comp$eeg <- full_join(
+  subgroup_comp$eeg, 
+  comp_summary$eeg, 
+  by = c("key", "index", "normal"))
+
+# Finally, plot everything:
+p1 <- plot_comparisons(
+  filter(subgroup_comp$eeg, index == 2),
+  filter(comp_summary$eeg, index == 2), 
+  1)
+
+p2 <-plot_comparisons(
+  filter(subgroup_comp$eeg, index != 2),
+  filter(comp_summary$eeg, index != 2), 
+  2)
+
+both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
+both_plts
+```
+
+## Display Duration Threshold
+
+We can do the same for the display duration thresholds.
+
+```{r ,fig.height=8, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups.
+The index of each relationship is indicated by the colour of the y-axis label.
+The fill of the density plots indicated the probability of the difference being less than zero."}
+
+# compute probabilties
+comp_summary$threshold <- comparison_summary(subgroup_comp$threshold, FALSE, 'threshold')
+
+subgroup_comp$threshold <- full_join(
+  subgroup_comp$threshold, 
+  comp_summary$threshold, 
+  by = c("key", "index", "normal"))
+
+# Finally, plot everything:
+p1 <- plot_comparisons(
+  filter(subgroup_comp$threshold, index == 2),
+  filter(comp_summary$threshold, index == 2), 
+  1, TRUE)
+p2 <- plot_comparisons(
+  filter(subgroup_comp$threshold, index != 2),
+  filter(comp_summary$threshold, index != 2), 
+  2, TRUE)
+
+both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
+both_plts
+```
+
+## Control Data
+
+We will now do exactly the same with the control data...
+
+```{r fig.height=8, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
+The index of each relationship is indicated by the colour of the y-axis label.
+The fill of the density plots indicated the probability of the difference being greater than zero."}
+
+# compute probabilties
+comp_summary$occ_even <- comparison_summary(subgroup_comp$occ_even, TRUE, 'rms')
+
+subgroup_comp$occ_even <- full_join(
+  subgroup_comp$occ_even, 
+  comp_summary$occ_even, 
+  by = c("key", "index", "normal"))
+
+p1 <- plot_comparisons(
+  filter(subgroup_comp$occ_even, index == 2),
+  filter(comp_summary$occ_even, index == 2), 
+  1)
+p2 <- plot_comparisons(
+  filter(subgroup_comp$occ_even, index != 2),
+  filter(comp_summary$occ_even, index != 2), 
+  2)
+
+both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
+both_plts
+
+# compute probabilties
+comp_summary$par_odd <- comparison_summary(subgroup_comp$par_odd, TRUE, 'rms')
+
+subgroup_comp$par_odd <- full_join(
+  subgroup_comp$par_odd, 
+  comp_summary$par_odd, 
+  by = c("key", "index", "normal"))
+
+p1 <- plot_comparisons(
+  filter(subgroup_comp$par_odd, index == 2),
+  filter(comp_summary$par_odd, index == 2), 
+  1)
+p2 <- plot_comparisons(
+  filter(subgroup_comp$par_odd, index != 2),
+  filter(comp_summary$par_odd, index != 2), 
+  2)
+
+both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
+both_plts
+```
+
+## Summary
+
+We can summarise the subgroup comparison plots above by plotting ROC curves for each of our four measurements (Figure 3.5).
+
+```{r, fig.cap="This figure shows how many of our 64 comparisons are classed as having a greater-than-zero difference (less-than-zero for the display durations) for difference thresholds.
+between 0.5 an 1.0."}
+
+count_comparisons <- function(comparisons, p) {
+  x <- sum(comparisons$prob_one_tailed >= p)
+  return(x)
+}
+
+df <- tibble()
+
+for (p in seq(0.5, 0.99, 0.01)) {
+
+  df <- bind_rows(df, map_dbl(comp_summary,  count_comparisons, p))
+}
+
+df$p <- seq(0.5, 0.99, 0.01)
+
+df %>% gather("measurement", "number", -p) %>%
+  mutate(proportion = number/64) %>%
+  ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
+  theme_minimal() + ggthemes::scale_color_ptol() +
+  xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")
+
+filter(df, p == 0.95)
+
+```

### 2020-05-07 analysis/analysis.Rmd
Added correlation back in

@@
 filter(df, p == 0.95)
+```
+
+# Additional Analysis
+
+## Index and Normality
+
+## Correlation Between EEG RMS and Display Duration Threshold
+
+Finally, we will investigate whether there is a correlation between the EEG rms measures and our display duration thresholds.
+As our two different measures come from different samples of participants, we are unable to do a direct comparison.
+However, we can use the results of the models discussed in Section 3 and compare p(eeg difference >0 | data) correlates with p(display duration difference < 0 | data).
+
+```{r compare_comparisons}
+
+bind_rows(
+  select(comp_summary$eeg, measurement, key, mean_value),
+  select(comp_summary$threshold, measurement, key, mean_value)) %>%
+  spread(measurement, mean_value) -> d
+
+m <- brm(data = d, threshold ~ rms)
+
+bayes_R2(m)
+```
+
+We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, $95\%$ credible interval is 0.026 - 0.302$).
+
+```{r fig.cap="Scatter plot showing the correlation between our two measures. Each line is a sample from the posterior of a Bayesian linear regression."}
+d %>% 
+  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
+  add_fitted_draws(m, n = 100) %>%
+  ggplot(aes(x = rms, y = threshold, color = index)) +
+  geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
+  geom_point(data = d) +
+  theme_minimal() + 
+  ggthemes::scale_colour_ptol()
 

### 2020-05-07 analysis/analysis.Rmd
index x normality analysis

@@
 
+Subgroup relations can be classified by their index, and wheter they are normal or not.
+Here we investigate the extent to which these two variables can account for the variation between the subgrop relationships.
+
+```{r}
+
+my_models <- list()
+my_models$full <- brm(data = comp_summary$eeg, mean_value ~ as.numeric(as.character(index))  + normal)
+
+summary(my_models$full)
+
+comp_summary$eeg %>% 
+  modelr::data_grid(index, normal) %>%
+  add_fitted_draws(my_models$full, n = 100) %>%
+  ggplot(aes(x =  as.numeric(as.character(index)), y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
+  geom_line(alpha = .2) +
+  geom_jitter(data = comp_summary$eeg, aes(x =  as.numeric(as.character(index)), y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
+  theme_minimal() + 
+  ggthemes::scale_colour_ptol() + scale_x_continuous(breaks = c(2,3,4,6,8,12)) +
+  theme(panel.grid.minor = element_blank())
+
+```
+
+```{r}
+
+my_models <- list()
+my_models$full <- brm(data = comp_summary$threshold, mean_value ~ as.numeric(as.character(index))  + normal)
+
+summary(my_models$full)
+
+comp_summary$threshold %>% 
+  modelr::data_grid(index, normal) %>%
+  add_fitted_draws(my_models$full, n = 100) %>%
+  ggplot(aes(x =  as.numeric(as.character(index)), y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
+  geom_line(alpha = .2) +
+  geom_jitter(data = comp_summary$threshold, aes(x =  as.numeric(as.character(index)), y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
+  theme_minimal() + 
+  ggthemes::scale_colour_ptol() + scale_x_continuous(breaks = c(2,3,4,6,8,12))+
+  theme(panel.grid.minor = element_blank())
+
+```
+
 ## Correlation Between EEG RMS and Display Duration Threshold
@@
 As our two different measures come from different samples of participants, we are unable to do a direct comparison.
-However, we can use the results of the models discussed in Section 3 and compare p(eeg difference >0 | data) correlates with p(display duration difference < 0 | data).
+However, we can use the results of the models discussed in Section 3 and check for a correlation between the predicted .
 
@@
 
-We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, $95\%$ credible interval is 0.026 - 0.302$).
+We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, 95\% credible interval is 0.026 - 0.302).
 

### 2020-05-07 analysis/analysis.Rmd
minor edits

--- 2020-05-07 analysis/analysis.Rmd
@@
 filter(df, p == 0.95)
+```
 
+# Additional Analysis
+
+## Index and Normality
+
+Subgroup relations can be classified by their index, and wheter they are normal or not.
+Here we investigate the extent to which these two variables can account for the variation between the subgrop relationships.
+
+```{r}
+
+my_models <- list()
+my_models$full <- brm(data = comp_summary$eeg, mean_value ~ as.numeric(as.character(index))  + normal)
+
+summary(my_models$full)
+
+comp_summary$eeg %>% 
+  modelr::data_grid(index, normal) %>%
+  add_fitted_draws(my_models$full, n = 100) %>%
+  ggplot(aes(x =  as.numeric(as.character(index)), y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
+  geom_line(alpha = .2) +
+  geom_jitter(data = comp_summary$eeg, aes(x =  as.numeric(as.character(index)), y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
+  theme_minimal() + 
+  ggthemes::scale_colour_ptol() + scale_x_continuous(breaks = c(2,3,4,6,8,12)) +
+  theme(panel.grid.minor = element_blank())
 ```
+
+```{r}
+my_models <- list()
+my_models$full <- brm(data = comp_summary$threshold, mean_value ~ as.numeric(as.character(index))  + normal)
+
+summary(my_models$full)
+
+comp_summary$threshold %>% 
+  modelr::data_grid(index, normal) %>%
+  add_fitted_draws(my_models$full, n = 100) %>%
+  ggplot(aes(x =  as.numeric(as.character(index)), y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
+  geom_line(alpha = .2) +
+  geom_jitter(data = comp_summary$threshold, aes(x =  as.numeric(as.character(index)), y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
+  theme_minimal() + 
+  ggthemes::scale_colour_ptol() + scale_x_continuous(breaks = c(2,3,4,6,8,12))+
+  theme(panel.grid.minor = element_blank())
+```
+
+## Correlation Between EEG RMS and Display Duration Threshold
+
+Finally, we will investigate whether there is a correlation between the EEG rms measures and our display duration thresholds.
+As our two different measures come from different samples of participants, we are unable to do a direct comparison.
+However, we can use the results of the models discussed in Section 3 and check for a correlation between the predicted .
+
+```{r compare_comparisons}
+bind_rows(
+  select(comp_summary$eeg, measurement, key, mean_value),
+  select(comp_summary$threshold, measurement, key, mean_value)) %>%
+  spread(measurement, mean_value) -> d
+
+m <- brm(data = d, threshold ~ rms)
+
+bayes_R2(m)
+```
+
+We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, 95\% credible interval is 0.026 - 0.302).
+
+```{r fig.cap="Scatter plot showing the correlation between our two measures. Each line is a sample from the posterior of a Bayesian linear regression."}
+d %>% 
+  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
+  add_fitted_draws(m, n = 100) %>%
+  ggplot(aes(x = rms, y = threshold, color = index)) +
+  geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
+  geom_point(data = d) +
+  theme_minimal() + 
+  ggthemes::scale_colour_ptol()
+```

### 2020-05-12 analysis/analysis.Rmd
Added figure
language edits to the sup mat

@@
 title: "Supplementary Materials"
-author: "Peter J. Kohler, Oda Åsgard, and Alasdair D. F. Clarke"
+author: "Peter J. Kohler and Alasdair D. F. Clarke"
 date:
@@
 
-Here we import our data and make some sumamry plots.
+Here we import our data and make some summary plots.
 
@@
 
-Importing the odd-harmonic EEG data from six electrodes region-of-interest over occipital cortex.
+Importing the primary EEG data set. This is odd-harmonic filtered data from region-of-interest consisting of six electrodes over occipital cortex.
 
-\tiny
+![RMS data from two of the wallpaper groups, P2 and P4M.
+Odd harmonics are shown in A and B, while even harmonic data are shown in C and D, and occipital and parietal regions of interest are shown in dark nad light gray, respectively.
+The two groups elicit very different response amplitudes for odd harmonics over occipital cortex, but for even harmonics those differences are much less pronounced.](../figures/occTopo_merged.png)
+
 ```{r eeg_import}
@@
 In addition to the primary EEG data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data 
-from six parietal electrodes (see main paper).
+from six parietal electrodes (see Figure 1.1 and the main paper).
 
@@
 
-### Fit Model to Display Duration Threshold Data
+### Fit Model to Psychophysical Data
 
@@
 plot_model_output(my_models$threshold, d_dispthresh, 
-                  "Model Fits: Display Duration Threshold", "orangered3", "threshold (ms)")
+                  "Model Fits: Psychophysical Data", "orangered3", "threshold (ms)")
 ```
@@
 
-## EEG RMS
+## Primary EEG Data
 
-Finally, I calculate the probability that the difference between subgroup and supergroup is larger than zero (small than zero for the display duration) given the data.
+Finally, we calculate the probability that the RMS difference between subgroup and supergroup is larger than zero given the data.
 This information is then binned so we can colour in the posterior density plots.
@@
 
-## Display Duration Threshold
+## Psychophysical Data
 
-We can do the same for the display duration thresholds.
+We can do the same for the display duration thresholds from our psychophysics experiment.
+Here we are looking for the opposite effect, namely that display larger are larger for subgroups than for supergroups (see main paper), so we calculate the probability that differences in duration are smaller than zero.
 
@@
 
-## Control Data
+## Control EEG Data
 
-We will now do exactly the same with the control data...
+We will now do exactly the same with the control data (odd harmonic data from parietal electrodes and even harmonic data from occipital electrodes)
 
@@
 
-Subgroup relations can be classified by their index, and wheter they are normal or not.
-Here we investigate the extent to which these two variables can account for the variation between the subgrop relationships.
+Subgroup relations can be classified by their index, and by whether they are normal or not.
+Here we investigate the extent to which these two variables can account for the variation between the subgroup relationships.
 
@@
 
-## Correlation Between EEG RMS and Display Duration Threshold
+## Correlation Between Primary EEG data and Psychophysical Thresholds
 
-Finally, we will investigate whether there is a correlation between the EEG rms measures and our display duration thresholds.
+Finally, we will investigate whether there is a correlation between the our primary EEG measure (rms amplitude of odd harmonics over occipital cortex) and our display duration thresholds.
 As our two different measures come from different samples of participants, we are unable to do a direct comparison.
-However, we can use the results of the models discussed in Section 3 and check for a correlation between the predicted .
+However, we can use the results of the models discussed in Section 3 and check for a correlation between the predicted values of the two measures.
 

### 2020-05-13 analysis/disp_dur_plot_for_paper.R
knitting Peter's changes

@@ -1,0 +1,33 @@
+library(tidyverse)
+
+
+fl <- "thresholds"
+
+my_cols_threshold <- cols(
+  person = col_character(),
+  group = col_character(),
+  staircase = col_double(),
+  threshold = col_double(),
+  log_threshold = col_double()
+)
+
+read_csv("../Experiment/process_results/thresholds.csv", col_types = my_cols_threshold) %>%
+  select(subject = "person", wg = "group", threshold) %>%
+  mutate(
+  	wg = as_factor(wg),
+  	wg = fct_relevel(wg, 
+  		"P2", "PG", "P3", "PGG", "PM", "P4", "CM", "P6", "P31M", "PMG", "P4G", "CMM", "P3M1", "PMM", "P6M", "P4M"),
+  	wg = fct_rev(wg)) %>%
+  glimpse() -> d_dispthresh 
+
+rm(my_cols_threshold)
+
+d_dispthresh %>% ggplot(aes(y = threshold * 1000, x = wg)) +
+	geom_boxplot(width = 0.25, fill = "grey") + coord_flip() +
+	theme_minimal() +
+	scale_y_log10("display duration threshold (ms)") +
+	theme(
+		panel.grid.minor = element_blank(),  
+		axis.title.y = element_blank(),
+		panel.grid.major = element_line(colour = "#E9E9E9"))
+	ggsave("disp_dur_boxplots.pdf", width = 2, height = 12)

### 2020-05-13 analysis/analysis.Rmd
Finally fixed first figure to have a proper caption!

@@
 library(patchwork)
+library(magick)
 
@@
 
-Use multiple cores and set number of iterations. 
 ```{r}
+# Use multiple cores and set number of iterations. 
 options(mc.cores = parallel::detectCores())
@@
 
-![RMS data from two of the wallpaper groups, P2 and P4M.
+```{r, fig.cap = "RMS data from two of the wallpaper groups, P2 and P4M.
 Odd harmonics are shown in A and B, while even harmonic data are shown in C and D, and occipital and parietal regions of interest are shown in dark nad light gray, respectively.
-The two groups elicit very different response amplitudes for odd harmonics over occipital cortex, but for even harmonics those differences are much less pronounced.](../figures/occTopo_merged.png)
+The two groups elicit very different response amplitudes for odd harmonics over occipital cortex, but for even harmonics those differences are much less pronounced."}
+plt <- cowplot::ggdraw() +
+  cowplot::draw_image("../figures/occTopo_merged.png")
+print(plt)
+
+```
+at
 

### 2020-05-13 analysis/analysis.Rmd
Edits to supmat

@@
 ```
-at
 
@@
 
-In addition to the primary EEG data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data 
-from six parietal electrodes (see Figure 1.1 and the main paper).
+In addition to the primary EEG data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data from six parietal electrodes (see Figure 1.1 and the main paper).
 
@@
 summary(m_eeg)
+
+# Also fit to display duration data so I can display both models together
+
+m_threshold <- brm(data = d_dispthresh, 
+                   threshold ~ wg-1 + (1|subject),
+                   family = "lognormal", 
+                   prior = prior, iter = n_iter,
+                   control = list(adapt_delta = 0.99))
 ```
 
-We will now have at what the model predicts for the average participant (i.e, ignoring the random intercepts).
-If anybody else would like to suggest over sensible ways to summarise this model, do let me know!
-I've been trying to use some of the functions from the `tidybayes` package, and they don't (yet?) have much on random effects.
+We will now look at the model's predicts for the average participant (i.e, ignoring the random intercepts).
 
@@
 my_samples$eeg <- get_model_samples(m_eeg)
+my_samples$threshold <- get_model_samples(m_threshold)
 
@@
  my_models$eeg <- get_model_predictions(my_samples$eeg)
+ my_models$threshold <- get_model_predictions(my_samples$threshold)
 
  # Plot
-plot_model_output(
+plt1 <- plot_model_output(
   my_models$eeg, d_eeg,
-  "Model Fits for the odd-harmonic data from occipital electrodes",
+  "odd-harmonic data from \noccipital electrodes",
   "orangered3", 
   "log(rms)")
+
+plt2 <- plot_model_output(
+  my_models$threshold, d_dispthresh, 
+  "Psychophysical Data", 
+  "orangered3", 
+  "threshold (ms)")
+
+plt1 + plt2 + plot_annotation(title = 'Model Fits:')
+
 ```
@@
 ```{r threshold_fit_model, cache=TRUE, echo=FALSE}
-m_threshold <- brm(data = d_dispthresh, 
-                   threshold ~ wg-1 + (1|subject),
-                   family = "lognormal", 
-                   prior = prior, iter = n_iter,
-                   control = list(adapt_delta = 0.99))
-
 summary(m_threshold)
-```
-
-```{r, fig.cap ="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
-# Extract samples from model
-my_samples$threshold <- get_model_samples(m_threshold)
-
-# Tidy up a little
-my_models$threshold <- get_model_predictions(my_samples$threshold)
-
-# Plot
-plot_model_output(my_models$threshold, d_dispthresh, 
-                  "Model Fits: Psychophysical Data", "orangered3", "threshold (ms)")
 ```
@@
 
-```{r cache=TRUE, fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
+```{r fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
 
@@
   1, TRUE)
+
 p2 <- plot_comparisons(
@@
 both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
-both_plts
+both_plts + plot_annotation(title = 'Odd harmonic data from parietal electrodes')
 
@@
 both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts
+both_plts + plot_annotation(title = 'Even harmonic data from occipital electrodes')
 ```
@@
   ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
-  theme_minimal() + ggthemes::scale_color_ptol() +
+  theme_minimal() + ggthemes::scale_color_ptol(labels = c("primary EEG", "sensor control EEG", "harmonic control EEG", "psychophysical threshold")) +
   xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")
+```
 
-filter(df, p == 0.95)
-```
+If we take p=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/64 = 87.5% and 49/64 = 76.6% of the comparisons for the primary EEG and display durations repectively.
+This compares to the $32/64= 50.0% and 22/64 = 34.4% for the control EEG conditions.
 
@@
 ```{r}
+my_models <- list()
+comp_summary$eeg$index <- as.numeric(as.character(comp_summary$eeg$index))
+comp_summary$threshold$index <- as.numeric(as.character(comp_summary$threshold$index))
 
-my_models <- list()
-my_models$full <- brm(data = comp_summary$eeg, mean_value ~ as.numeric(as.character(index))  + normal)
+my_models$eeg <- brm(data = comp_summary$eeg, mean_value ~ index * normal)
+my_models$disp <- brm(data = comp_summary$threshold, mean_value ~ index * normal)
+summary(my_models$eeg)
+summary(my_models$disp)
+```
 
-summary(my_models$full)
+We can see that the index of the subgroup relationship has an effect on both the difference in log(rms) and the difference in log(display duration): relationships with a higher index lead to larger differences.
 
+```{r, fig.cap = "The effect of index and normality on log(rms) and log(ms)."}
 comp_summary$eeg %>% 
   modelr::data_grid(index, normal) %>%
-  add_fitted_draws(my_models$full, n = 100) %>%
-  ggplot(aes(x =  as.numeric(as.character(index)), y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
-  geom_line(alpha = .2) +
-  geom_jitter(data = comp_summary$eeg, aes(x =  as.numeric(as.character(index)), y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
-  theme_minimal() + 
-  ggthemes::scale_colour_ptol() + scale_x_continuous(breaks = c(2,3,4,6,8,12)) +
-  theme(panel.grid.minor = element_blank())
-```
-
-```{r}
-my_models <- list()
-my_models$full <- brm(data = comp_summary$threshold, mean_value ~ as.numeric(as.character(index))  + normal)
-
-summary(my_models$full)
+  add_fitted_draws(my_models$eeg, n = 100) %>%
+  mutate(model = "primary eeg") -> d_plt_eeg
 
@@
   modelr::data_grid(index, normal) %>%
-  add_fitted_draws(my_models$full, n = 100) %>%
-  ggplot(aes(x =  as.numeric(as.character(index)), y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
-  geom_line(alpha = .2) +
-  geom_jitter(data = comp_summary$threshold, aes(x =  as.numeric(as.character(index)), y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
+  add_fitted_draws(my_models$disp, n = 100) %>%
+  mutate(model = "psychophysical threshold") -> d_plt_disp
+
+d_summary <- bind_rows(comp_summary$threshold %>% mutate(model = "psychophysical threshold"),
+                       comp_summary$eeg %>% mutate(model = "primary eeg"))
+
+bind_rows(d_plt_eeg, d_plt_disp) %>%
+  ggplot(aes(x =  index, y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
+  geom_line(alpha = .15) +
+  facet_wrap(~ model, scales="free_y") + 
+  geom_jitter(
+    data = d_summary, 
+    aes(x =  index, y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
   theme_minimal() + 
-  ggthemes::scale_colour_ptol() + scale_x_continuous(breaks = c(2,3,4,6,8,12))+
-  theme(panel.grid.minor = element_blank())
+  ggthemes::scale_colour_ptol(labels = c("non-normal", "normal")) + 
+  scale_x_continuous("index", breaks = c(2,3,4,6,8,12)) +
+  scale_y_continuous("value") +
+  theme(panel.grid.minor = element_blank(), legend.title = element_blank()) 
 ```
@@
   add_fitted_draws(m, n = 100) %>%
-  ggplot(aes(x = rms, y = threshold, color = index)) +
+  ggplot(aes(x = rms, y = threshold, color = as.factor(index))) +
   geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
@@
   theme_minimal() + 
-  ggthemes::scale_colour_ptol()
+  ggthemes::scale_colour_ptol("index")
 ```

### 2020-05-15 analysis/scripts/get_subgroup_comparisons.R
Script to read in the subgroup relations from csv files.

@@ -33,7 +33,7 @@ get_subgroup_comparisons <- function(m_samples, n_iter)
         'PM$\\rightarrow$ PG'  = PM  - PG, 
         'PGG$\\rightarrow$ PG' = PGG - PG, 
         'PGG$\\rightarrow$ P2' = PGG - P2, 
-        'P2$\\rightarrow$ PG'  = P2 - PG,
+        # 'P2$\\rightarrow$ PG'  = P2 - PG,
         'P6M$\\rightarrow$ CMM' = P6M - CMM,
         'P3M1$\\rightarrow$ CM' = P3M1 - CM,
         'P31M$\\rightarrow$ CM' = P31M - CM,
@@ -99,7 +99,7 @@ get_subgroup_comparisons <- function(m_samples, n_iter)
         'PM$\\rightarrow$ PG',  
         'PGG$\\rightarrow$ PG', 
         'PGG$\\rightarrow$ P2', 
-        'P2$\\rightarrow$ PG' ,
+        # 'P2$\\rightarrow$ PG' ,
         'P6M$\\rightarrow$ CMM', 
         'P3M1$\\rightarrow$ CM' ,
         'P31M$\\rightarrow$ CM' ,
@@ -138,14 +138,14 @@ get_subgroup_comparisons <- function(m_samples, n_iter)
       gather() %>%
       mutate(
         index = as.factor(c(
-          rep(2, n_iter * 30), 
+          rep(2, n_iter * 29), 
           rep(3, n_iter * 4), 
           rep(4, n_iter * 17), 
           rep(6, n_iter * 9),
           rep(8, n_iter * 2),
           rep(12,n_iter * 2))),
         normal = c(
-          rep(1, n_iter * 30), 
+          rep(1, n_iter * 29), 
           rep(c(0, 0, 0, 1), each = n_iter), 
           rep(c(0, 1, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 1, 0), each = n_iter),
           rep(c(0, 0, 0, 0, 1, 0, 0, 0, 0), each = n_iter),
@@ -156,7 +156,6 @@ get_subgroup_comparisons <- function(m_samples, n_iter)
     return(subgroup_comp)
 
 }
-
 
 plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE) 
 {

### 2020-05-15 analysis/scripts/import_subgroup_comparisons.R

@@ -1,0 +1,39 @@
+library(tidyverse)
+
+
+index_dat  <- read_csv("../../subgroups/subgroup_relations.csv")
+normal_dat <- read_csv("../../subgroups/subgroup_normal.csv")
+
+# sort out group names
+group_names  <- index_dat$group
+index_dat  <- index_dat[, -1]
+normal_dat <- normal_dat[, -1]
+
+# remove p1
+group_names  <- group_names[-1]
+index_dat  <- select(index_dat[-1,], -p1)
+normal_dat <- select(normal_dat[-1,], -p1)
+
+subgroups <- tibble(
+    group = as.character(), 
+    subgroup = as.character(), 
+    index = as.numeric(),
+    normal = as.logical())
+
+for (ii in 1:nrow(index_dat)) {
+    for (jj in 1:ncol(index_dat)) {
+
+
+        subgroups <- add_row(subgroups, 
+            group = group_names[ii], 
+            subgroup = group_names[jj], 
+            index = as.numeric(index_dat[ii,jj]),
+            normal = normal_dat[ii,jj] == 1)
+    }
+} 
+
+# filter by index to remove all non-valid relations
+# also remove all indentity relationships
+subgroups %>% filter(
+    is.finite(index),
+    group != subgroup) -> subgroups

### 2020-05-15 analysis/analysis.Rmd
Script to read in the subgroup relations from csv files.
Subgroup comparison tables.

@@
 
-It's clearly skewed, and negative display duration are impossible, so will fit a `glm` with `family = 'lognormal`.
+It's clearly skewed, and negative display duration are impossible, so will fit a `glm` with `family = 'lognormal'`.
 
@@
 
+## Subgroup Information
+
+Import subgroup information and display a table of the relationships that we will be investigating.
+
+```{r}
+source("scripts/import_subgroup_comparisons.R")
+
+glimpse(subgroups)
+opts <- options(knitr.kable.NA = "")
+knitr::kable(index_table, caption = "subgroup index")
+```
+
+```{r}
+knitr::kable(normal_table, caption = "normal and non-normal subgroups")
+```
+
 # Bayesian Analysis
@@
 
-```{r, fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
+```{r, fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals.", cache=TRUE}
 prior_samples <- get_model_samples(m_prior)
@@
 
-```{r fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
+```{r fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals.", cache=TRUE}
 
@@
 
-```{r}
+```{r, cache=TRUE}
 my_models <- list()
@@
 
-```{r compare_comparisons}
+```{r compare_comparisons, cache=TRUE}
 bind_rows(

### 2020-05-15 analysis/scripts/import_subgroup_comparisons.R

@@ -1,11 +1,11 @@
 library(tidyverse)
 
 
-index_dat  <- read_csv("../../subgroups/subgroup_relations.csv")
-normal_dat <- read_csv("../../subgroups/subgroup_normal.csv")
+index_dat  <- read_csv("../subgroups/subgroup_relations.csv")
+normal_dat <- read_csv("../subgroups/subgroup_normal.csv")
 
 # sort out group names
-group_names  <- index_dat$group
+group_names  <- str_to_upper(index_dat$group)
 index_dat  <- index_dat[, -1]
 normal_dat <- normal_dat[, -1]
 
@@ -23,17 +23,25 @@ subgroups <- tibble(
 for (ii in 1:nrow(index_dat)) {
     for (jj in 1:ncol(index_dat)) {
 
-
         subgroups <- add_row(subgroups, 
-            group = group_names[ii], 
-            subgroup = group_names[jj], 
+            group = group_names[jj], 
+            subgroup = group_names[ii], 
             index = as.numeric(index_dat[ii,jj]),
             normal = normal_dat[ii,jj] == 1)
     }
 } 
 
+subgroups %>% 
+    select(group, subgroup, index) %>% 
+    pivot_wider(names_from = group, values_from = index) -> index_table
+
+subgroups %>% 
+    select(group, subgroup, normal) %>% 
+    pivot_wider(names_from = group, values_from = normal) -> normal_table
+
 # filter by index to remove all non-valid relations
 # also remove all indentity relationships
 subgroups %>% filter(
     is.finite(index),
-    group != subgroup) -> subgroups
+    group != subgroup) %>%
+    mutate(label = paste(group, "$\\rightarrow$", subgroup)) -> subgroups

### 2020-05-15 analysis/analysis.Rmd
p2, p3, p4, p6 comparison

@@
 
+## Replicatation of Kohler et al (2016)
+
+We can lok at the groups that only contain orientation, and see if we obtain the parametric response as documented in Kohler et al.
+(2016).
+We one might predict based on those results, we get a corresponding pattern of display duration thresholds, with the time taken to detect the symmetry dreasing as we increase the amount of rotational symmetry.
+
+Curiosuly, we do not have a large difference betwee P4 and P6 in the eeg data.
+
+```{r, fig.cap = "The effect of increasing the amount of rotational symmetry.", fig.height=3}
+my_samples$eeg %>% select(P2, P3, P4, P6) %>%
+  pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>% 
+  mutate(exp = "eeg") %>%
+  bind_rows(
+    my_samples$threshold %>% select(P2, P3, P4, P6) %>% 
+      pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>%
+      mutate(exp = "display thresholds")) %>% 
+  ggplot() + aes(x = value, fill = `wallpaper group`) + 
+    geom_density(alpha = 0.35) + 
+    facet_wrap(~ exp, scale = "free_y") + 
+  theme_minimal() + 
+  scale_fill_viridis_d()
+```
+
 ## Index and Normality

### 2020-05-26 analysis/analysis.Rmd
trying to fix subgroup comp functions

@@
 
-```{r subgroup_comparisons, cache=TRUE}
+```{r subgroup_comparisons}
 source("scripts/get_subgroup_comparisons.R")
 
-subgroup_comp$eeg <- get_subgroup_comparisons(my_samples$eeg, 2*n_iter)
-subgroup_comp$threshold <- get_subgroup_comparisons(my_samples$threshold, 2*n_iter)
-subgroup_comp$occ_even <- get_subgroup_comparisons(my_samples$occ_even, 2*n_iter)
-subgroup_comp$par_odd  <- get_subgroup_comparisons(my_samples$par_odd,  2*n_iter)
+subgroup_comp$eeg <- get_subgroup_comparisons2(my_samples$eeg)
+subgroup_comp$threshold <- get_subgroup_comparisons2(my_samples$threshold,)
+subgroup_comp$occ_even <- get_subgroup_comparisons2(my_samples$occ_even)
+subgroup_comp$par_odd  <- get_subgroup_comparisons2(my_samples$par_odd)
 ```

### 2020-05-26 analysis/scripts/get_subgroup_comparisons.R

@@ -1,3 +1,20 @@
+get_subgroup_comparisons2 <- function(m_samples, n_iter) {
+  
+  subgroup_comp <- tibble(key = as.character(), value = as.numeric())
+  
+  for (ii in 1:nrow(subgroups)) {
+   subgroup_comp <- bind_rows(
+     subgroup_comp,
+     tibble(
+       key = subgroups$label[ii],
+       index = subgroups$index[ii],
+       normal = subgroups$normal[ii],
+       value = (m_samples[subgroups$subgroup[ii]] - m_samples[subgroups$group[ii]])[[1]]))
+  }
+  
+  return(subgroup_comp)
+}
+
 get_subgroup_comparisons <- function(m_samples, n_iter)
 {
     m_samples %>%

### 2020-05-27 analysis/analysis.Rmd
Fixed subgroup comparisons?

@@
 
+filter(df, p == 0.95)
+
 df %>% gather("measurement", "number", -p) %>%
-  mutate(proportion = number/64) %>%
+  mutate(proportion = number/nrow(subgroups)) %>%
   ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
   theme_minimal() + ggthemes::scale_color_ptol(labels = c("primary EEG", "sensor control EEG", "harmonic control EEG", "psychophysical threshold")) +
+  scale_y_continuous(limits = c(0, 1)) +
   xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")

### 2020-05-27 analysis/scripts/get_subgroup_comparisons.R

@@ -12,6 +12,8 @@ get_subgroup_comparisons2 <- function(m_samples, n_iter) {
        value = (m_samples[subgroups$subgroup[ii]] - m_samples[subgroups$group[ii]])[[1]]))
   }
   
+  subgroup_comp$index <- as.factor(subgroup_comp$index)
+  
   return(subgroup_comp)
 }
 

### 2020-05-29 analysis/analysis.Rmd
Fixed fill issue with subgroup comparison plots

@@
 ```{r setup, include=FALSE}
-
 library(bookdown)
@@
 source("scripts/helper_functions.R")
+
+options(mc.cores = parallel::detectCores())
 ```
@@
     mutate(
+      prob_one_tailed = if_else(prob_one_tailed==0, 0.01, prob_one_tailed),
       measurement = measurement_lab,

### 2020-05-29 analysis/scripts/get_subgroup_comparisons.R

@@ -206,10 +206,10 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
             plot.margin=unit(c(1,1,1.5,1.2),"cm"))  
        
      if (is_thresholds) {
-        plt <- plt + coord_cartesian((xlim = c(-1.5, 0.4))) + 
+        plt <- plt + coord_cartesian((xlim = c(-2.2, 0.7))) + 
             scale_fill_viridis_d("p(difference < 0 | data)", option = "plasma", drop = FALSE)  
     } else {
-         plt <- plt + coord_cartesian((xlim = c(-0.4, 1))) + 
+         plt <- plt + coord_cartesian((xlim = c(-0.5, 1.5))) + 
             scale_fill_viridis_d("p(difference > 0 | data)", option = "plasma", drop = FALSE) 
     }
 

### 2020-05-29 analysis/analysis.Rmd
Coloured coded table

@@
 glimpse(subgroups)
+
 opts <- options(knitr.kable.NA = "")
-knitr::kable(index_table, caption = "subgroup index")
-```
 
-```{r}
-knitr::kable(normal_table, caption = "normal and non-normal subgroups")
+knitr::kable(index_table, caption = "subgroup index", escape = "F") %>% kableExtra::kable_styling(full_width = F)
 ```

### 2020-05-29 analysis/test.Rmd
Coloured coded table

--- 2020-05-29 analysis/analysis.Rmd
@@
 ```
-
-# Bayesian Analysis
-
-Here are the details of the Bayesian multi-level modelling. Our general approach is:
-
-*  Use weakly-informative priors.
-*  Fit a multi-level model using the lognormal distribution.
-  + independent variable (wallpaper group) and one dependent variable (rms or threshold).
-  + maximal random effect structure.
-* Check model fit (r̂ statistics and trace plots).
-* Extract mcmc samples from the posterior distributions.
-* Use these samples to estimate the difference between sub- and super-groups.
-
-## Define Priors
-
-In this section we will specify some priors. We then then use a prior-predictive check to assess whether the prior is reasonable or not (i.e., on the same order of magnitude as our measurements).
-
-### Fixed Effects
-
-Our independent variable is a categorical factor with 16 levels.
-We will drop the intercept from our model and instead fit a coefficent for each factor level ($y \sim x - 0$).
-As our dependant variable will be log-transformed, we can use the priors below:
-
-```{r, echo = TRUE}
-prior <- c(
-  set_prior("normal(0,2)", class = "b"),    
-  set_prior("cauchy(0,2)", class = "sigma"))
-```
-
-### Group-level Effects
-
-We will keep the default weakly informative priors for the group-level (‘random’) effects. From the brms documentation:
-
-> [...] restricted to be non-negative and, by default, have a half student-t prior with 3 degrees of freedom and a scale parameter that depends on the standard deviation of the response after applying the link function.
-Minimally, the scale parameter is 10.
-This prior is used (a) to be only very weakly informative in order to influence results as few as possible, while (b) providing at least some regularization to considerably improve convergence and sampling efficiency.
-
-### Prior Predictive Check
-
-Now we can specify our Bayesian multi-level model and priors. Note that as we are using  `sample_prior = 'only'`, the model will not learn anything from our data.
-
-```{r eeg_prior_pred_check, cache=TRUE, echo=TRUE}
-m_prior <- brm(data = d_eeg, 
-  rms ~ wg-1 + (1|subject),
-  family = "lognormal", 
-  prior = prior, 
-  iter = n_iter ,
-  sample_prior = 'only')
-```
-
-We can use this model to generate data.
-
-```{r, fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals.", cache=TRUE}
-prior_samples <- get_model_samples(m_prior)
-
-# Tidy up a little
-prior_predictions <- get_model_predictions(prior_samples)%>% glimpse()
-
-# # And now plot
-p1 <- plot_model_output(prior_predictions, d_eeg,
-                   "odd-harmonic data from \noccipital electrodes", "cornflowerblue", "log(rms)")
-
-p2 <- plot_model_output(prior_predictions, d_dispthresh,
-                   "display duration threshold data", "cornflowerblue", "log(ms)")
-
-# draw pot
-plt <- p1 + p2
-plt + plot_annotation(title = 'Prior Predictive Checks')
-```
-
-We can see that i) our priors are relatively weak as the predictions span several orders of magnitide, and ii) our empirical data falls within this range.
-
-## Compute Posterior
-
-### Fit Model to EEG Data
-
-We will now fit the model to the data.
-
-```{r eeg_fit_model, cache=TRUE}
-m_eeg <- brm(data = d_eeg, 
-             rms ~ wg-1 + (1|subject),
-             family = "lognormal", 
-             prior = prior, iter = n_iter,
-             control = list(adapt_delta = 0.99))
-
-summary(m_eeg)
-
-# Also fit to display duration data so I can display both models together
-
-m_threshold <- brm(data = d_dispthresh, 
-                   threshold ~ wg-1 + (1|subject),
-                   family = "lognormal", 
-                   prior = prior, iter = n_iter,
-                   control = list(adapt_delta = 0.99))
-```
-
-We will now look at the model's predicts for the average participant (i.e, ignoring the random intercepts).
-
-```{r, fig.cap ="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
-
- # Extract samples from model
-my_samples$eeg <- get_model_samples(m_eeg)
-my_samples$threshold <- get_model_samples(m_threshold)
-
- # Tidy up a little
- my_models$eeg <- get_model_predictions(my_samples$eeg)
- my_models$threshold <- get_model_predictions(my_samples$threshold)
-
- # Plot
-plt1 <- plot_model_output(
-  my_models$eeg, d_eeg,
-  "odd-harmonic data from \noccipital electrodes",
-  "orangered3", 
-  "log(rms)")
-
-plt2 <- plot_model_output(
-  my_models$threshold, d_dispthresh, 
-  "Psychophysical Data", 
-  "orangered3", 
-  "threshold (ms)")
-
-plt1 + plt2 + plot_annotation(title = 'Model Fits:')
-
-```
-
-### Fit Model to Psychophysical Data
-
-We will now fit the model to the data.
-
-```{r threshold_fit_model, cache=TRUE, echo=FALSE}
-summary(m_threshold)
-```
-
-### EEG Control Data
-
-We will also fit models to the control data. As we can see from Figure 2.4, the group differences are much smaller.
-
-```{r fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals.", cache=TRUE}
-
-m_eeg_par_odd <- brm(data = d_eeg_par_odd, 
-                     rms ~ wg-1 + (1|subject),
-                     family = "lognormal", 
-                     prior = prior, iter = n_iter,
-                     control = list(adapt_delta = 0.99))
-
-my_samples$par_odd <- get_model_samples(m_eeg_par_odd)
-my_models$par_odd <- get_model_predictions(my_samples$par_odd)
-plt1 <- plot_model_output(my_models$par_odd, d_eeg_par_odd, 
-                  "odd-harmonic data from \nparietal electrodes", "orangered3", "threshold (ms)")
-
-m_eeg_occ_even <- brm(data = d_eeg_occ_even, 
-                      rms ~ wg-1 + (1|subject),
-                      family = "lognormal", 
-                      prior = prior, iter = n_iter,
-                      control = list(adapt_delta = 0.99))
-
-# Extract samples from model
-my_samples$occ_even <- get_model_samples(m_eeg_occ_even)
-my_models$occ_even <- get_model_predictions(my_samples$occ_even)
-plt2 <- plot_model_output(my_models$occ_even, d_eeg_occ_even, 
-                  "even-harmonic data from \noccipital electrodes", "orangered3", "threshold (ms)")
-
-plt1 + plt2 + plot_annotation(title = 'Model Fits for the ')
-```
-
-# Subgroup Comparisons
-
-We will now compute the difference between sub- and super-groups.
-
-```{r subgroup_comparisons}
-source("scripts/get_subgroup_comparisons.R")
-
-subgroup_comp$eeg <- get_subgroup_comparisons2(my_samples$eeg)
-subgroup_comp$threshold <- get_subgroup_comparisons2(my_samples$threshold,)
-subgroup_comp$occ_even <- get_subgroup_comparisons2(my_samples$occ_even)
-subgroup_comp$par_odd  <- get_subgroup_comparisons2(my_samples$par_odd)
-```
-
-## Primary EEG Data
-
-Finally, we calculate the probability that the RMS difference between subgroup and supergroup is larger than zero given the data.
-This information is then binned so we can colour in the posterior density plots.
-
-```{r define_comparison_summary_function}
-comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
-  # compute probabilties
-  subgroup_comp_dat %>%
-    group_by(key, index, normal) %>%
-    summarise(
-      mean_value = mean(value),
-      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)))  %>%
-    mutate(
-      prob_one_tailed = if_else(prob_one_tailed==0, 0.01, prob_one_tailed),
-      measurement = measurement_lab,
-      index_labs = as.numeric(as.factor(index)),
-      lab_cols = v_cols[index_labs],
-      prob_cat = cut(prob_one_tailed, c(0, 0.5, 0.75, 0.9, 0.99, 1))) -> comp_summary_dat
-
-  return(comp_summary_dat)
-}
-```
-
-Now we will use it!
-
-```{r, fig.height=8, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
-The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being greater than zero."}
-# compute probabilties
-comp_summary$eeg <- comparison_summary(subgroup_comp$eeg, TRUE, 'rms')
-
-subgroup_comp$eeg <- full_join(
-  subgroup_comp$eeg, 
-  comp_summary$eeg, 
-  by = c("key", "index", "normal"))
-
-# Finally, plot everything:
-p1 <- plot_comparisons(
-  filter(subgroup_comp$eeg, index == 2),
-  filter(comp_summary$eeg, index == 2), 
-  1)
-
-p2 <-plot_comparisons(
-  filter(subgroup_comp$eeg, index != 2),
-  filter(comp_summary$eeg, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts
-```
-
-## Psychophysical Data
-
-We can do the same for the display duration thresholds from our psychophysics experiment.
-Here we are looking for the opposite effect, namely that display larger are larger for subgroups than for supergroups (see main paper), so we calculate the probability that differences in duration are smaller than zero.
-
-```{r ,fig.height=8, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups.
-The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being less than zero."}
-
-# compute probabilties
-comp_summary$threshold <- comparison_summary(subgroup_comp$threshold, FALSE, 'threshold')
-
-subgroup_comp$threshold <- full_join(
-  subgroup_comp$threshold, 
-  comp_summary$threshold, 
-  by = c("key", "index", "normal"))
-
-# Finally, plot everything:
-p1 <- plot_comparisons(
-  filter(subgroup_comp$threshold, index == 2),
-  filter(comp_summary$threshold, index == 2), 
-  1, TRUE)
-
-p2 <- plot_comparisons(
-  filter(subgroup_comp$threshold, index != 2),
-  filter(comp_summary$threshold, index != 2), 
-  2, TRUE)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts
-```
-
-## Control EEG Data
-
-We will now do exactly the same with the control data (odd harmonic data from parietal electrodes and even harmonic data from occipital electrodes)
-
-```{r fig.height=8, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
-The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being greater than zero."}
-
-# compute probabilties
-comp_summary$occ_even <- comparison_summary(subgroup_comp$occ_even, TRUE, 'rms')
-
-subgroup_comp$occ_even <- full_join(
-  subgroup_comp$occ_even, 
-  comp_summary$occ_even, 
-  by = c("key", "index", "normal"))
-
-p1 <- plot_comparisons(
-  filter(subgroup_comp$occ_even, index == 2),
-  filter(comp_summary$occ_even, index == 2), 
-  1)
-p2 <- plot_comparisons(
-  filter(subgroup_comp$occ_even, index != 2),
-  filter(comp_summary$occ_even, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
-both_plts + plot_annotation(title = 'Odd harmonic data from parietal electrodes')
-
-# compute probabilties
-comp_summary$par_odd <- comparison_summary(subgroup_comp$par_odd, TRUE, 'rms')
-
-subgroup_comp$par_odd <- full_join(
-  subgroup_comp$par_odd, 
-  comp_summary$par_odd, 
-  by = c("key", "index", "normal"))
-
-p1 <- plot_comparisons(
-  filter(subgroup_comp$par_odd, index == 2),
-  filter(comp_summary$par_odd, index == 2), 
-  1)
-p2 <- plot_comparisons(
-  filter(subgroup_comp$par_odd, index != 2),
-  filter(comp_summary$par_odd, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts + plot_annotation(title = 'Even harmonic data from occipital electrodes')
-```
-
-## Summary
-
-We can summarise the subgroup comparison plots above by plotting ROC curves for each of our four measurements (Figure 3.5).
-
-```{r, fig.cap="This figure shows how many of our 64 comparisons are classed as having a greater-than-zero difference (less-than-zero for the display durations) for difference thresholds.
-between 0.5 an 1.0."}
-
-count_comparisons <- function(comparisons, p) {
-  x <- sum(comparisons$prob_one_tailed >= p)
-  return(x)
-}
-
-df <- tibble()
-
-for (p in seq(0.5, 0.99, 0.01)) {
-
-  df <- bind_rows(df, map_dbl(comp_summary,  count_comparisons, p))
-}
-
-df$p <- seq(0.5, 0.99, 0.01)
-
-filter(df, p == 0.95)
-
-df %>% gather("measurement", "number", -p) %>%
-  mutate(proportion = number/nrow(subgroups)) %>%
-  ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
-  theme_minimal() + ggthemes::scale_color_ptol(labels = c("primary EEG", "sensor control EEG", "harmonic control EEG", "psychophysical threshold")) +
-  scale_y_continuous(limits = c(0, 1)) +
-  xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")
-```
-
-If we take p=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/64 = 87.5% and 49/64 = 76.6% of the comparisons for the primary EEG and display durations repectively.
-This compares to the $32/64= 50.0% and 22/64 = 34.4% for the control EEG conditions.
-
-# Additional Analysis
-
-## Replicatation of Kohler et al (2016)
-
-We can lok at the groups that only contain orientation, and see if we obtain the parametric response as documented in Kohler et al.
-(2016).
-We one might predict based on those results, we get a corresponding pattern of display duration thresholds, with the time taken to detect the symmetry dreasing as we increase the amount of rotational symmetry.
-
-Curiosuly, we do not have a large difference betwee P4 and P6 in the eeg data.
-
-```{r, fig.cap = "The effect of increasing the amount of rotational symmetry.", fig.height=3}
-my_samples$eeg %>% select(P2, P3, P4, P6) %>%
-  pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>% 
-  mutate(exp = "eeg") %>%
-  bind_rows(
-    my_samples$threshold %>% select(P2, P3, P4, P6) %>% 
-      pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>%
-      mutate(exp = "display thresholds")) %>% 
-  ggplot() + aes(x = value, fill = `wallpaper group`) + 
-    geom_density(alpha = 0.35) + 
-    facet_wrap(~ exp, scale = "free_y") + 
-  theme_minimal() + 
-  scale_fill_viridis_d()
-```
-
-## Index and Normality
-
-Subgroup relations can be classified by their index, and by whether they are normal or not.
-Here we investigate the extent to which these two variables can account for the variation between the subgroup relationships.
-
-```{r, cache=TRUE}
-my_models <- list()
-comp_summary$eeg$index <- as.numeric(as.character(comp_summary$eeg$index))
-comp_summary$threshold$index <- as.numeric(as.character(comp_summary$threshold$index))
-
-my_models$eeg <- brm(data = comp_summary$eeg, mean_value ~ index * normal)
-my_models$disp <- brm(data = comp_summary$threshold, mean_value ~ index * normal)
-summary(my_models$eeg)
-summary(my_models$disp)
-```
-
-We can see that the index of the subgroup relationship has an effect on both the difference in log(rms) and the difference in log(display duration): relationships with a higher index lead to larger differences.
-
-```{r, fig.cap = "The effect of index and normality on log(rms) and log(ms)."}
-comp_summary$eeg %>% 
-  modelr::data_grid(index, normal) %>%
-  add_fitted_draws(my_models$eeg, n = 100) %>%
-  mutate(model = "primary eeg") -> d_plt_eeg
-
-comp_summary$threshold %>% 
-  modelr::data_grid(index, normal) %>%
-  add_fitted_draws(my_models$disp, n = 100) %>%
-  mutate(model = "psychophysical threshold") -> d_plt_disp
-
-d_summary <- bind_rows(comp_summary$threshold %>% mutate(model = "psychophysical threshold"),
-                       comp_summary$eeg %>% mutate(model = "primary eeg"))
-
-bind_rows(d_plt_eeg, d_plt_disp) %>%
-  ggplot(aes(x =  index, y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
-  geom_line(alpha = .15) +
-  facet_wrap(~ model, scales="free_y") + 
-  geom_jitter(
-    data = d_summary, 
-    aes(x =  index, y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
-  theme_minimal() + 
-  ggthemes::scale_colour_ptol(labels = c("non-normal", "normal")) + 
-  scale_x_continuous("index", breaks = c(2,3,4,6,8,12)) +
-  scale_y_continuous("value") +
-  theme(panel.grid.minor = element_blank(), legend.title = element_blank()) 
-```
-
-## Correlation Between Primary EEG data and Psychophysical Thresholds
-
-Finally, we will investigate whether there is a correlation between the our primary EEG measure (rms amplitude of odd harmonics over occipital cortex) and our display duration thresholds.
-As our two different measures come from different samples of participants, we are unable to do a direct comparison.
-However, we can use the results of the models discussed in Section 3 and check for a correlation between the predicted values of the two measures.
-
-```{r compare_comparisons, cache=TRUE}
-bind_rows(
-  select(comp_summary$eeg, measurement, key, mean_value),
-  select(comp_summary$threshold, measurement, key, mean_value)) %>%
-  spread(measurement, mean_value) -> d
-
-m <- brm(data = d, threshold ~ rms)
-
-bayes_R2(m)
-```
-
-We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, 95\% credible interval is 0.026 - 0.302).
-
-```{r fig.cap="Scatter plot showing the correlation between our two measures. Each line is a sample from the posterior of a Bayesian linear regression."}
-d %>% 
-  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
-  add_fitted_draws(m, n = 100) %>%
-  ggplot(aes(x = rms, y = threshold, color = as.factor(index))) +
-  geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
-  geom_point(data = d) +
-  theme_minimal() + 
-  ggthemes::scale_colour_ptol("index")
-```

### 2020-05-29 analysis/scripts/import_subgroup_comparisons.R

@@ -32,6 +32,14 @@ for (ii in 1:nrow(index_dat)) {
 } 
 
 subgroups %>% 
+    mutate(
+        normal_colour = as.numeric(normal),
+        normal_colour = if_else(is.na(normal_colour), 0, normal_colour),
+        normal_colour = if_else(normal_colour== 1, 1, 0),
+        index = kableExtra::cell_spec(
+            index, 
+            color = kableExtra::spec_color(as.numeric(normal), end = 0.7, na_color = "#FFFFFF",  option = "magma")),
+        ) %>%
     select(group, subgroup, index) %>% 
     pivot_wider(names_from = group, values_from = index) -> index_table
 

### 2020-05-29 analysis/test.Rmd
Removing undeeded comparisons and more table colouring

@@
 
-# read in some helper functions that will be used later
-source("scripts/helper_functions.R")
-
-options(mc.cores = parallel::detectCores())
-```
-
-# Loading packages and importing data
-
-##  Packages
-
-Details of packages, etc, are given below.
-
-```{r}
-sessionInfo()
-```
-
-```{r}
-# Use multiple cores and set number of iterations. 
-options(mc.cores = parallel::detectCores())
-n_iter <- 10000
-
-# make empty lists to store things in
-my_models <- list()
-my_samples <- list()
-comp_summary <- list()
-subgroup_comp <- list()
-```
-
-## Import Data
-
-Here we import our data and make some summary plots.
-
-### EEG Data
-
-Importing the primary EEG data set. This is odd-harmonic filtered data from region-of-interest consisting of six electrodes over occipital cortex.
-
-```{r, fig.cap = "RMS data from two of the wallpaper groups, P2 and P4M.
-Odd harmonics are shown in A and B, while even harmonic data are shown in C and D, and occipital and parietal regions of interest are shown in dark nad light gray, respectively.
-The two groups elicit very different response amplitudes for odd harmonics over occipital cortex, but for even harmonics those differences are much less pronounced."}
-plt <- cowplot::ggdraw() +
-  cowplot::draw_image("../figures/occTopo_merged.png")
-print(plt)
-
-```
-
-```{r eeg_import}
-fl <- "rms_Occ_Odd.csv"
-
-# set column types
-my_cols <- cols(
-  .default = col_double(),
-  Row = col_character()
-)
-
-#import
-read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
-  rename(wg = "Row") %>%
-  gather("subject", "rms", -wg) %>%
-  glimpse() -> d_eeg 
-```
-
-It's clearly skewed, and negative display duration are impossible, so will fit a `glm` with `family = 'lognormal'`.
-
-```{r cache=TRUE, fig.cap = "Data are the root-mean-squaured (rms) over the odd-harmonic filtered waveforms."}
-d_eeg %>% ggplot(aes(x = rms)) + 
-    geom_histogram(binwidth = 0.2, fill = "skyblue", colour = "black") +
-    theme_lucid()
-```
-
-### Threshold Data
-
-Here we import that data and select the columns that we're interested in. Threshold gives the required display duration (in seconds) for the two stimuli to allow for accurate discrimination.
-
-```{r threshold_import}
-fl <- "thresholds"
-
-my_cols_threshold <- cols(
-  person = col_character(),
-  group = col_character(),
-  staircase = col_double(),
-  threshold = col_double(),
-  log_threshold = col_double()
-)
-
-read_csv("../Experiment/process_results/thresholds.csv", col_types = my_cols_threshold) %>%
-  select(subject = "person", wg = "group", threshold) %>%
-  glimpse() -> d_dispthresh
-
-rm(my_cols_threshold)
-```
-
-As above, a summary of the data.
-
-```{r, fig.cap="Histogram of the display duration thresholds."}
-d_dispthresh %>% ggplot(aes(x = threshold)) + 
-    geom_histogram(binwidth = 0.2, fill = "skyblue", colour = "black") +
-    theme_lucid()
-```
-
-Again, we have a skewed distribution, so will fit with `family = 'lognormal'`.
-
-### Control Data
-
-In addition to the primary EEG data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data from six parietal electrodes (see Figure 1.1 and the main paper).
-
-```{r import_control}
-
-fl <- "rms_Occ_Even.csv"
-
-read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
-  rename(wg = "Row") %>%
-  gather("subject", "rms", -wg) -> d_eeg_occ_even
-
-fl <- "rms_Par_Odd.csv"
-
-read_csv(paste("data/", fl, sep = ""), col_types = my_cols) %>%
-  rename(wg = "Row") %>%
-  gather("subject", "rms", -wg) -> d_eeg_par_odd
-
-rm(my_cols, fl)
 ```

### 2020-05-29 analysis/scripts/import_subgroup_comparisons.R

@@ -31,25 +31,30 @@ for (ii in 1:nrow(index_dat)) {
     }
 } 
 
+
+# list of comparisons to remove
+to_remove <- list(
+    "PM-CM", "CM-PM",
+    "PMM-CMM", "CMM-PMM",
+    "P3M1-P31M", "P31M-P3M1")
+
 subgroups %>% 
     mutate(
-        normal_colour = as.numeric(normal),
-        normal_colour = if_else(is.na(normal_colour), 0, normal_colour),
-        normal_colour = if_else(normal_colour== 1, 1, 0),
+        colour_code = as.numeric(normal)+1,
+        colour_code = if_else(group == subgroup, 3, colour_code),
+        colour_code = if_else(paste(group, subgroup, sep="-") %in% to_remove, 4, colour_code), 
         index = kableExtra::cell_spec(
             index, 
-            color = kableExtra::spec_color(as.numeric(normal), end = 0.7, na_color = "#FFFFFF",  option = "magma")),
+            color = kableExtra::spec_color(colour_code, end = 0.7, na_color = "#FFFFFF",  option = "magma")),
         ) %>%
     select(group, subgroup, index) %>% 
     pivot_wider(names_from = group, values_from = index) -> index_table
 
-subgroups %>% 
-    select(group, subgroup, normal) %>% 
-    pivot_wider(names_from = group, values_from = normal) -> normal_table
 
 # filter by index to remove all non-valid relations
 # also remove all indentity relationships
 subgroups %>% filter(
     is.finite(index),
-    group != subgroup) %>%
+    group != subgroup,
+    !(paste(group, subgroup, sep="-") %in% to_remove)) %>%
     mutate(label = paste(group, "$\\rightarrow$", subgroup)) -> subgroups

### 2020-05-29 analysis/scripts/import_subgroup_comparisons.R
Added sym in group csv file data.

@@ -45,7 +45,7 @@ subgroups %>%
         colour_code = if_else(paste(group, subgroup, sep="-") %in% to_remove, 4, colour_code), 
         index = kableExtra::cell_spec(
             index, 
-            color = kableExtra::spec_color(colour_code, end = 0.7, na_color = "#FFFFFF",  option = "magma")),
+            color = kableExtra::spec_color(colour_code, end = 0.9, na_color = "#FFFFFF")),
         ) %>%
     select(group, subgroup, index) %>% 
     pivot_wider(names_from = group, values_from = index) -> index_table

### 2020-05-29 analysis/analysis.Rmd
Removing undeeded comparisons and more table colouring
added graphical representation of symmetries in each group

@@
 
-## Subgroup Information
+## Symmetry Information
 
+### Subgroup Relationships
 Import subgroup information and display a table of the relationships that we will be investigating.
@@
 knitr::kable(index_table, caption = "subgroup index", escape = "F") %>% kableExtra::kable_styling(full_width = F)
+```
+
+### Types of Symmetry
+
+```{r, fig.cap = "Symmetries present in each group"}
+sym_dat <- read_csv("../subgroups/symmeries_in_group.csv") %>%
+  mutate(group = str_to_upper(group))
+
+sym_dat %>% pivot_longer(-group, names_to = "symmetry", values_to = "order") %>%
+  ggplot(aes(x = symmetry, y = order, fill = as.factor(order))) + geom_col() + facet_wrap(~ group) +
+  theme_bw() + 
+  scale_y_continuous(breaks = c(2,4,6)) + 
+  ggthemes::scale_fill_ptol() + 
+  theme(
+    panel.grid.minor=element_blank(), 
+    legend.position = "none")
+
 ```

### 2020-06-02 analysis/analysis.Rmd
Modelling different types of symmetry.

@@
 
-```{r, fig.cap = "Symmetries present in each group"}
+```{r, fig.cap = "Symmetries present in each group."}
 sym_dat <- read_csv("../subgroups/symmeries_in_group.csv") %>%

### 2020-06-02 analysis/analysis2.Rmd
Modelling different types of symmetry.

--- 2020-06-02 analysis/analysis.Rmd
@@
 
-I have set `echo=FALSE` so that most of the code chunks will not display. Please refer to the .Rmd file for source code.
+This will be merged into the main file later
 
@@
 
-### Fit Model to Psychophysical Data
+## Modelling by symmetry
 
-We will now fit the model to the data.
+```{r, cache=TRUE}
+d_eeg %>% rename(group = "wg") %>% left_join(sym_dat) -> dat2
 
-```{r threshold_fit_model, cache=TRUE, echo=FALSE}
-summary(m_threshold)
-```
-
-### EEG Control Data
-
-We will also fit models to the control data. As we can see from Figure 2.4, the group differences are much smaller.
-
-```{r fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals.", cache=TRUE}
-
-m_eeg_par_odd <- brm(data = d_eeg_par_odd, 
-                     rms ~ wg-1 + (1|subject),
+m_eeg2 <- brm(data = dat2, 
+             rms ~ rotation + reflection + glide + (1|subject),
              family = "lognormal", 
@@
 
-my_samples$par_odd <- get_model_samples(m_eeg_par_odd)
-my_models$par_odd <- get_model_predictions(my_samples$par_odd)
-plt1 <- plot_model_output(my_models$par_odd, d_eeg_par_odd, 
-                  "odd-harmonic data from \nparietal electrodes", "orangered3", "threshold (ms)")
+```
 
-m_eeg_occ_even <- brm(data = d_eeg_occ_even, 
-                      rms ~ wg-1 + (1|subject),
+```{r, cache=TRUE}
+d_dispthresh %>% rename(group = "wg") %>% left_join(sym_dat) -> dat3
+
+m_disthresh2 <- brm(data = dat3, 
+             threshold ~ rotation + reflection + glide + (1|subject),
              family = "lognormal", 
@@
 
-# Extract samples from model
-my_samples$occ_even <- get_model_samples(m_eeg_occ_even)
-my_models$occ_even <- get_model_predictions(my_samples$occ_even)
-plt2 <- plot_model_output(my_models$occ_even, d_eeg_occ_even, 
-                  "even-harmonic data from \noccipital electrodes", "orangered3", "threshold (ms)")
-
-plt1 + plt2 + plot_annotation(title = 'Model Fits for the ')
 ```
 
-# Subgroup Comparisons
+```{r}
+summary(m_eeg2)
 
-We will now compute the difference between sub- and super-groups.
-
-```{r subgroup_comparisons}
-source("scripts/get_subgroup_comparisons.R")
-
-subgroup_comp$eeg <- get_subgroup_comparisons2(my_samples$eeg)
-subgroup_comp$threshold <- get_subgroup_comparisons2(my_samples$threshold,)
-subgroup_comp$occ_even <- get_subgroup_comparisons2(my_samples$occ_even)
-subgroup_comp$par_odd  <- get_subgroup_comparisons2(my_samples$par_odd)
+summary(m_disthresh2)
 ```
-
-## Primary EEG Data
-
-Finally, we calculate the probability that the RMS difference between subgroup and supergroup is larger than zero given the data.
-This information is then binned so we can colour in the posterior density plots.
-
-```{r define_comparison_summary_function}
-comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
-  # compute probabilties
-  subgroup_comp_dat %>%
-    group_by(key, index, normal) %>%
-    summarise(
-      mean_value = mean(value),
-      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)))  %>%
-    mutate(
-      prob_one_tailed = if_else(prob_one_tailed==0, 0.01, prob_one_tailed),
-      measurement = measurement_lab,
-      index_labs = as.numeric(as.factor(index)),
-      lab_cols = v_cols[index_labs],
-      prob_cat = cut(prob_one_tailed, c(0, 0.5, 0.75, 0.9, 0.99, 1))) -> comp_summary_dat
-
-  return(comp_summary_dat)
-}
-```
-
-Now we will use it!
-
-```{r, fig.height=8, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
-The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being greater than zero."}
-# compute probabilties
-comp_summary$eeg <- comparison_summary(subgroup_comp$eeg, TRUE, 'rms')
-
-subgroup_comp$eeg <- full_join(
-  subgroup_comp$eeg, 
-  comp_summary$eeg, 
-  by = c("key", "index", "normal"))
-
-# Finally, plot everything:
-p1 <- plot_comparisons(
-  filter(subgroup_comp$eeg, index == 2),
-  filter(comp_summary$eeg, index == 2), 
-  1)
-
-p2 <-plot_comparisons(
-  filter(subgroup_comp$eeg, index != 2),
-  filter(comp_summary$eeg, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts
-```
-
-## Psychophysical Data
-
-We can do the same for the display duration thresholds from our psychophysics experiment.
-Here we are looking for the opposite effect, namely that display larger are larger for subgroups than for supergroups (see main paper), so we calculate the probability that differences in duration are smaller than zero.
-
-```{r ,fig.height=8, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups.
-The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being less than zero."}
-
-# compute probabilties
-comp_summary$threshold <- comparison_summary(subgroup_comp$threshold, FALSE, 'threshold')
-
-subgroup_comp$threshold <- full_join(
-  subgroup_comp$threshold, 
-  comp_summary$threshold, 
-  by = c("key", "index", "normal"))
-
-# Finally, plot everything:
-p1 <- plot_comparisons(
-  filter(subgroup_comp$threshold, index == 2),
-  filter(comp_summary$threshold, index == 2), 
-  1, TRUE)
-
-p2 <- plot_comparisons(
-  filter(subgroup_comp$threshold, index != 2),
-  filter(comp_summary$threshold, index != 2), 
-  2, TRUE)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts
-```
-
-## Control EEG Data
-
-We will now do exactly the same with the control data (odd harmonic data from parietal electrodes and even harmonic data from occipital electrodes)
-
-```{r fig.height=8, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
-The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being greater than zero."}
-
-# compute probabilties
-comp_summary$occ_even <- comparison_summary(subgroup_comp$occ_even, TRUE, 'rms')
-
-subgroup_comp$occ_even <- full_join(
-  subgroup_comp$occ_even, 
-  comp_summary$occ_even, 
-  by = c("key", "index", "normal"))
-
-p1 <- plot_comparisons(
-  filter(subgroup_comp$occ_even, index == 2),
-  filter(comp_summary$occ_even, index == 2), 
-  1)
-p2 <- plot_comparisons(
-  filter(subgroup_comp$occ_even, index != 2),
-  filter(comp_summary$occ_even, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
-both_plts + plot_annotation(title = 'Odd harmonic data from parietal electrodes')
-
-# compute probabilties
-comp_summary$par_odd <- comparison_summary(subgroup_comp$par_odd, TRUE, 'rms')
-
-subgroup_comp$par_odd <- full_join(
-  subgroup_comp$par_odd, 
-  comp_summary$par_odd, 
-  by = c("key", "index", "normal"))
-
-p1 <- plot_comparisons(
-  filter(subgroup_comp$par_odd, index == 2),
-  filter(comp_summary$par_odd, index == 2), 
-  1)
-p2 <- plot_comparisons(
-  filter(subgroup_comp$par_odd, index != 2),
-  filter(comp_summary$par_odd, index != 2), 
-  2)
-
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
-both_plts + plot_annotation(title = 'Even harmonic data from occipital electrodes')
-```
-
-## Summary
-
-We can summarise the subgroup comparison plots above by plotting ROC curves for each of our four measurements (Figure 3.5).
-
-```{r, fig.cap="This figure shows how many of our 64 comparisons are classed as having a greater-than-zero difference (less-than-zero for the display durations) for difference thresholds.
-between 0.5 an 1.0."}
-
-count_comparisons <- function(comparisons, p) {
-  x <- sum(comparisons$prob_one_tailed >= p)
-  return(x)
-}
-
-df <- tibble()
-
-for (p in seq(0.5, 0.99, 0.01)) {
-
-  df <- bind_rows(df, map_dbl(comp_summary,  count_comparisons, p))
-}
-
-df$p <- seq(0.5, 0.99, 0.01)
-
-filter(df, p == 0.95)
-
-df %>% gather("measurement", "number", -p) %>%
-  mutate(proportion = number/nrow(subgroups)) %>%
-  ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
-  theme_minimal() + ggthemes::scale_color_ptol(labels = c("primary EEG", "sensor control EEG", "harmonic control EEG", "psychophysical threshold")) +
-  scale_y_continuous(limits = c(0, 1)) +
-  xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")
-```
-
-If we take p=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/64 = 87.5% and 49/64 = 76.6% of the comparisons for the primary EEG and display durations repectively.
-This compares to the $32/64= 50.0% and 22/64 = 34.4% for the control EEG conditions.
-
-# Additional Analysis
-
-## Replicatation of Kohler et al (2016)
-
-We can lok at the groups that only contain orientation, and see if we obtain the parametric response as documented in Kohler et al.
-(2016).
-We one might predict based on those results, we get a corresponding pattern of display duration thresholds, with the time taken to detect the symmetry dreasing as we increase the amount of rotational symmetry.
-
-Curiosuly, we do not have a large difference betwee P4 and P6 in the eeg data.
-
-```{r, fig.cap = "The effect of increasing the amount of rotational symmetry.", fig.height=3}
-my_samples$eeg %>% select(P2, P3, P4, P6) %>%
-  pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>% 
-  mutate(exp = "eeg") %>%
-  bind_rows(
-    my_samples$threshold %>% select(P2, P3, P4, P6) %>% 
-      pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>%
-      mutate(exp = "display thresholds")) %>% 
-  ggplot() + aes(x = value, fill = `wallpaper group`) + 
-    geom_density(alpha = 0.35) + 
-    facet_wrap(~ exp, scale = "free_y") + 
-  theme_minimal() + 
-  scale_fill_viridis_d()
-```
-
-## Index and Normality
-
-Subgroup relations can be classified by their index, and by whether they are normal or not.
-Here we investigate the extent to which these two variables can account for the variation between the subgroup relationships.
-
-```{r, cache=TRUE}
-my_models <- list()
-comp_summary$eeg$index <- as.numeric(as.character(comp_summary$eeg$index))
-comp_summary$threshold$index <- as.numeric(as.character(comp_summary$threshold$index))
-
-my_models$eeg <- brm(data = comp_summary$eeg, mean_value ~ index * normal)
-my_models$disp <- brm(data = comp_summary$threshold, mean_value ~ index * normal)
-summary(my_models$eeg)
-summary(my_models$disp)
-```
-
-We can see that the index of the subgroup relationship has an effect on both the difference in log(rms) and the difference in log(display duration): relationships with a higher index lead to larger differences.
-
-```{r, fig.cap = "The effect of index and normality on log(rms) and log(ms)."}
-comp_summary$eeg %>% 
-  modelr::data_grid(index, normal) %>%
-  add_fitted_draws(my_models$eeg, n = 100) %>%
-  mutate(model = "primary eeg") -> d_plt_eeg
-
-comp_summary$threshold %>% 
-  modelr::data_grid(index, normal) %>%
-  add_fitted_draws(my_models$disp, n = 100) %>%
-  mutate(model = "psychophysical threshold") -> d_plt_disp
-
-d_summary <- bind_rows(comp_summary$threshold %>% mutate(model = "psychophysical threshold"),
-                       comp_summary$eeg %>% mutate(model = "primary eeg"))
-
-bind_rows(d_plt_eeg, d_plt_disp) %>%
-  ggplot(aes(x =  index, y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
-  geom_line(alpha = .15) +
-  facet_wrap(~ model, scales="free_y") + 
-  geom_jitter(
-    data = d_summary, 
-    aes(x =  index, y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
-  theme_minimal() + 
-  ggthemes::scale_colour_ptol(labels = c("non-normal", "normal")) + 
-  scale_x_continuous("index", breaks = c(2,3,4,6,8,12)) +
-  scale_y_continuous("value") +
-  theme(panel.grid.minor = element_blank(), legend.title = element_blank()) 
-```
-
-## Correlation Between Primary EEG data and Psychophysical Thresholds
-
-Finally, we will investigate whether there is a correlation between the our primary EEG measure (rms amplitude of odd harmonics over occipital cortex) and our display duration thresholds.
-As our two different measures come from different samples of participants, we are unable to do a direct comparison.
-However, we can use the results of the models discussed in Section 3 and check for a correlation between the predicted values of the two measures.
-
-```{r compare_comparisons, cache=TRUE}
-bind_rows(
-  select(comp_summary$eeg, measurement, key, mean_value),
-  select(comp_summary$threshold, measurement, key, mean_value)) %>%
-  spread(measurement, mean_value) -> d
-
-m <- brm(data = d, threshold ~ rms)
-
-bayes_R2(m)
-```
-
-We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, 95\% credible interval is 0.026 - 0.302).
-
-```{r fig.cap="Scatter plot showing the correlation between our two measures. Each line is a sample from the posterior of a Bayesian linear regression."}
-d %>% 
-  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
-  add_fitted_draws(m, n = 100) %>%
-  ggplot(aes(x = rms, y = threshold, color = as.factor(index))) +
-  geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
-  geom_point(data = d) +
-  theme_minimal() + 
-  ggthemes::scale_colour_ptol("index")
-```

### 2020-06-05 analysis/analysis.Rmd
updated subgroup table

@@
 
-glimpse(subgroups)
-
 opts <- options(knitr.kable.NA = "")
@@
 
-sym_dat %>% pivot_longer(-group, names_to = "symmetry", values_to = "order") %>%
-  ggplot(aes(x = symmetry, y = order, fill = as.factor(order))) + geom_col() + facet_wrap(~ group) +
-  theme_bw() + 
-  scale_y_continuous(breaks = c(2,4,6)) + 
-  ggthemes::scale_fill_ptol() + 
-  theme(
-    panel.grid.minor=element_blank(), 
-    legend.position = "none")
+knitr::kable(sym_dat, caption = "wallpaper group summary", escape = "F") %>% kableExtra::kable_styling(full_width = F)
 

### 2020-06-05 analysis/scripts/get_subgroup_comparisons.R

@@ -1,6 +1,7 @@
 get_subgroup_comparisons2 <- function(m_samples, n_iter) {
   
-  subgroup_comp <- tibble(key = as.character(), value = as.numeric())
+  subgroup_comp <- tibble(key = as.character(), value = as.numeric(), 
+                          d_ref = as.numeric(), d_rot = as.numeric(), d_gli = as.numeric())
   
   for (ii in 1:nrow(subgroups)) {
    subgroup_comp <- bind_rows(
@@ -9,7 +10,13 @@ get_subgroup_comparisons2 <- function(m_samples, n_iter) {
        key = subgroups$label[ii],
        index = subgroups$index[ii],
        normal = subgroups$normal[ii],
-       value = (m_samples[subgroups$subgroup[ii]] - m_samples[subgroups$group[ii]])[[1]]))
+       value = (m_samples[subgroups$subgroup[ii]] - m_samples[subgroups$group[ii]])[[1]],
+       d_ref = filter(sym_dat, group == subgroups$group[ii])$reflection -
+         filter(sym_dat, group == subgroups$subgroup[ii])$reflection,
+       d_rot = filter(sym_dat, group == subgroups$group[ii])$rotation -
+         filter(sym_dat, group == subgroups$subgroup[ii])$rotation,
+       d_gli = filter(sym_dat, group == subgroups$group[ii])$glide -
+         filter(sym_dat, group == subgroups$subgroup[ii])$glide))
   }
   
   subgroup_comp$index <- as.factor(subgroup_comp$index)

### 2020-06-05 analysis/scripts/import_subgroup_comparisons.R

@@ -38,17 +38,31 @@ to_remove <- list(
     "PMM-CMM", "CMM-PMM",
     "P3M1-P31M", "P31M-P3M1")
 
+to_remove <- append(to_remove, unlist(paste(group_names, "-", group_names, sep = "")))
+
 subgroups %>% 
+    select(group, subgroup, index) %>%
+    pivot_wider(
+        names_from = subgroup, 
+        values_from = index) %>%
+    pivot_longer(
+        -group, 
+        values_to = "index", 
+        names_to = "subgroup") %>%
+    left_join(subgroups) %>%
     mutate(
-        colour_code = as.numeric(normal)+1,
-        colour_code = if_else(group == subgroup, 3, colour_code),
-        colour_code = if_else(paste(group, subgroup, sep="-") %in% to_remove, 4, colour_code), 
+        colour_code = if_else(paste(group, subgroup, sep="-") %in% to_remove, 2, 1), 
+       index = as.character(index),
+        index = if_else(is.na(index), "-", index),
         index = kableExtra::cell_spec(
-            index, 
-            color = kableExtra::spec_color(colour_code, end = 0.9, na_color = "#FFFFFF")),
+             index, 
+             italic = if_else(normal == TRUE, TRUE, FALSE),
+             color = kableExtra::spec_color(colour_code, end = 0.95, na_color = "#FFFFFF")),
         ) %>%
     select(group, subgroup, index) %>% 
-    pivot_wider(names_from = group, values_from = index) -> index_table
+    pivot_wider(
+        names_from = group, 
+        values_from = index) -> index_table
 
 
 # filter by index to remove all non-valid relations

### 2020-06-08 analysis/analysis.Rmd
Tidying up and fixing some bugs

@@
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, fig.align='center')
+options(knitr.table.format = "html")
 
@@
 
+Information on the symmetries and subgroups contained within each wallpaper group was obtained from <cite> and is summarised in the files `symmeries_in_group.csv`, `subgroup_relations.csv`, and `subgroup_normal.csv`.
+
+### Types of Symmetry
+
+<A short note on what is counted...i.e., so the reader understands that P2 has one rotation of order 2, etc>
+
+```{r, fig.cap = "Symmetries present in each group."}
+
+sym_dat <- read_csv("../subgroups/symmeries_in_group.csv") %>%
+  mutate(group = str_to_upper(group))
+
+knitr::kable(sym_dat, caption = "wallpaper group summary") %>% 
+    kableExtra::kable_styling(full_width = F)
+```
+
 ### Subgroup Relationships
@@
 
-knitr::kable(index_table, caption = "subgroup index", escape = "F") %>% kableExtra::kable_styling(full_width = F)
+knitr::kable(index_table, caption = "Summary of subgroup relationships.
+The numbers indicate the index of the subgroup, while italics indicate normal subgroups.
+Relationships written in yellow text are not included in our analysis.", escape = "F") %>% kableExtra::kable_styling(full_width = F)
 ```
 
-### Types of Symmetry
-
-```{r, fig.cap = "Symmetries present in each group."}
-sym_dat <- read_csv("../subgroups/symmeries_in_group.csv") %>%
-  mutate(group = str_to_upper(group))
-
-knitr::kable(sym_dat, caption = "wallpaper group summary", escape = "F") %>% kableExtra::kable_styling(full_width = F)
-
-```
+We will remove identity relationships (i.e., a group is a subgroup of itself) and the three pairs of wallpapers groups that it subgroups of each other (e.g., PM is a subgroup of CM, and CM is a subgroup of PM).
+This leaves us with a total of `r nrow(subgroups)` subgroups to include in our analysis.
 
@@
 ```{r define_comparison_summary_function}
+
 comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
@@
 The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being greater than zero."}
+The fill of the density plots indicated the probability of the difference being greater than zero.", cache=TRUE}
+
 # compute probabilties
@@
 both_plts
+
 ```
@@
 
-```{r ,fig.height=8, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups.
+```{r ,fig.height=8, cache = TRUE, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups.
 The index of each relationship is indicated by the colour of the y-axis label.
@@
 
-```{r fig.height=8, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
+```{r fig.height=8, cache=TRUE, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
 The index of each relationship is indicated by the colour of the y-axis label.
@@
 
-```{r, fig.cap="This figure shows how many of our 64 comparisons are classed as having a greater-than-zero difference (less-than-zero for the display durations) for difference thresholds.
+```{r, fig.cap="This figure shows how many of our 63 comparisons are classed as having a greater-than-zero difference (less-than-zero for the display durations) for difference thresholds.
 between 0.5 an 1.0."}
@@
 
-If we take p=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/64 = 87.5% and 49/64 = 76.6% of the comparisons for the primary EEG and display durations repectively.
-This compares to the $32/64= 50.0% and 22/64 = 34.4% for the control EEG conditions.
+If we take $p$=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/63 = 89% and 49/64 = 78% of the comparisons for the primary EEG and display durations repectively.
+This compares to the 32/64= 50% and 22/64 = 35% for the control EEG conditions.
 
@@
 
+First of all, we run for the eeg rms data.
+
 ```{r, cache=TRUE}
+
+# change index and normal variables to numeric for modelling
 my_models <- list()
-comp_summary$eeg$index <- as.numeric(as.character(comp_summary$eeg$index))
-comp_summary$threshold$index <- as.numeric(as.character(comp_summary$threshold$index))
+
+comp_summary$eeg %>% ungroup() %>%
+   mutate(
+     index = as.numeric(as.character(index)),
+     normal = if_else(normal, 1, 0)) -> comp_summary$eeg
 
 my_models$eeg <- brm(data = comp_summary$eeg, mean_value ~ index * normal)
+
+summary(my_models$eeg)
+
+```
+
+And now for the display duration thresholds.
+
+```{r, cache=TRUE}
+
+# change index and normal variables to numeric for modelling
+
+comp_summary$threshold %>% ungroup() %>%
+   mutate(
+     index = as.numeric(as.character(index)),
+     normal = if_else(normal, 1, 0)) -> comp_summary$threshold
+
 my_models$disp <- brm(data = comp_summary$threshold, mean_value ~ index * normal)
-summary(my_models$eeg)
+
 summary(my_models$disp)
+
 ```
@@
   mutate(model = "primary eeg") -> d_plt_eeg
-
+# 
 comp_summary$threshold %>% 
@@
 bind_rows(
-  select(comp_summary$eeg, measurement, key, mean_value),
-  select(comp_summary$threshold, measurement, key, mean_value)) %>%
+  select(comp_summary$eeg, measurement, key, mean_value, index),
+  select(comp_summary$threshold, measurement, key, mean_value, index)) %>%
   spread(measurement, mean_value) -> d
@@
 
-bayes_R2(m)
+bR2 <- bayes_R2(m)
+
+print(bR2)
+
 ```
 
-We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, 95\% credible interval is 0.026 - 0.302).
+We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, 95\% credible interval is `r round(bR2[3],2)` - `r round(bR2[4], 2)`).
 
 ```{r fig.cap="Scatter plot showing the correlation between our two measures. Each line is a sample from the posterior of a Bayesian linear regression."}
+
 d %>%
-  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
+  modelr::data_grid(rms = seq(-0.2, 1, 0.1)) %>%
   add_fitted_draws(m, n = 100) %>%
@@
   ggthemes::scale_colour_ptol("index")
+
 ```

### 2020-06-18 analysis/analysis2.Rmd
updated subgroup table
Subgroup comparison by symmetry model.
Why are the results odd?
trying to work out why there's no main effect of rotation

@@
 
-options(mc.cores = parallel::detectCores())
+# options(mc.cores = parallel::detectCores())
 ```
@@
 ```
+
+## Modelling by Subgroup
+
+```{r}
+source("scripts/get_subgroup_comparisons.R")
+
+subgroup_comp$eeg <- get_subgroup_comparisons2(my_samples$eeg, sym_dat)
+
+subgroup_comp$eeg %>% group_by(key) %>%
+  summarise(
+    index = unique(index),
+    d_ref = unique(d_ref),
+    d_rot = unique(d_rot),
+    d_gli = unique(d_gli),
+    value = mean(value)
+  ) -> comp_info
+
+m <- brm(data = comp_info, value ~ d_rot + d_ref + d_gli)
+
+```
+
+```{r}
+summary(m)
+
+knitr::kable(comp_info) %>% kableExtra::kable_styling()
+
+comp_info %>% filter(d_ref == 0, d_gli == 0) %>% ggplot(aes(x = d_rot, y = value, label = key)) + geom_point() + geom_smooth(method = "lm") + ggtitle("rotations only: no reflections or glides")+ ggrepel::geom_label_repel()
+
+comp_info %>% filter(d_rot == 0, d_gli == 0) %>% ggplot(aes(x = d_ref, y = value, label = key)) + geom_point() + geom_smooth(method = "lm") + ggtitle("reflections only: no rotations or glides")+ ggrepel::geom_label_repel()
+
+comp_info %>% filter(d_ref == 0) %>% ggplot(aes(x = d_rot, y = value, label = key)) + geom_point() + geom_smooth(method = "lm") + ggtitle("no reflections") + ggrepel::geom_label_repel()
+
+comp_info %>% filter(d_rot == 0) %>% ggplot(aes(x = d_ref, y = value, label = key)) + geom_point() + geom_smooth(method = "lm") + ggtitle("no rotations")+ ggrepel::geom_label_repel()
+
+```

### 2020-06-18 analysis/analysis.Rmd
Moving SessionInfo to the end
parameteric orientations

@@
 
-# Loading packages and importing data
-
-##  Packages
-
-Details of packages, etc, are given below.
-
-```{r}
-sessionInfo()
-```
+# Importing data
 
@@
 
-## Import Data
-
 Here we import our data and make some summary plots.
 
-### EEG Data
+## EEG Data
 
@@
 ```
+```{r}
 
-### Threshold Data
+d_eeg %>% filter(wg %in% c("P2", "P3", "P4", "P6")) %>% ggplot(aes(x = wg, y = log(rms))) + geom_boxplot(notch = T)
+
+d_eeg %>% group_by(wg) %>% summarise(
+  mean_log_rms = mean(log(rms)),
+  sd_log_rms = sd(log(rms)),
+  median_log_rms = median(log(rms)))
+
+```
+
+## Threshold Data
 
@@
 
-### Control Data
+## Control Data
 
@@
 
-### Types of Symmetry
+## Types of Symmetry
 
@@
 
-### Subgroup Relationships
+## Subgroup Relationships
 Import subgroup information and display a table of the relationships that we will be investigating.
@@
 
+Will try fitting a new parametric model on rotations
+
+```{r}
+
+d_eeg %>% filter(wg %in% c("P2", "P3", "P4", "P6")) %>%
+  mutate(orientation = parse_number(wg))-> d_eeg_rot
+
+m_eeg_rot <- brm(data = d_eeg_rot, rms ~ orientation + (1|subject), family = "lognormal")
+
+d_eeg_rot %>% modelr::data_grid(orientation) %>%
+  add_fitted_draws(m_eeg_rot, re_formula = NA, n=200) %>%
+  ggplot() + 
+  geom_line(data = d_eeg_rot, aes(x = orientation, y = rms, group = subject), alpha = 0.1, colour = "darkred") +
+  geom_path(aes(x = orientation, y = .value, group = .draw), alpha = .2, colour = "thistle4") + 
+  theme_minimal() 
+```
+
 ## Index and Normality
@@
 my_models$eeg <- brm(data = comp_summary$eeg, mean_value ~ index * normal)
-
-summary(my_models$eeg)
 
@@
 ```
+
+# Package and Session Information
+
+Details of packages, etc, are given below.
+
+```{r}
+sessionInfo()
+```

### 2020-06-24 analysis/analysis.Rmd
Added para-orientation model

--- 2020-06-08 analysis/analysis.Rmd
@@
 ```
+
+We now import the data. We have three variables: wallpaper group (wg), subject, and root-mean-squared amplitude (rms).
 
@@
 
-It's clearly skewed, and negative display duration are impossible, so will fit a `glm` with `family = 'lognormal'`.
+If we plot the distribution of rms, we can clearly see that it is skewed. Furthermore, as negative rms amplitudes are impossible we will use a `lognormal` distribution to model these data.
 
-```{r cache=TRUE, fig.cap = "Data are the root-mean-squaured (rms) over the odd-harmonic filtered waveforms."}
+```{r cache=TRUE, fig.cap = "Data are the root-mean-squaured (rms) amplitudes over the odd-harmonic filtered waveforms."}
 d_eeg %>% ggplot(aes(x = rms)) + 
@@
 
-### Threshold Data
+The table below shows how many groups per participant we have data for.
+
+```{r}
+d_eeg %>% group_by(subject) %>% summarise(count = n()) %>% knitr::kable()
+```
+
+## Threshold Data
 
@@
 As above, a summary of the data.
+And again, we have a skewed distribution, (with negative display durations being impossible), so we will use also use a `lognormal` distribution to model the behaviour data.
 
@@
 
-Again, we have a skewed distribution, so will fit with `family = 'lognormal'`.
+The table below shows how many groups per participant we have data for.
 
-### Control Data
+```{r}
+d_dispthresh %>% group_by(subject) %>% summarise(count = n()) %>% knitr::kable()
+```
+
+## Control Data
 
@@
 
-We can lok at the groups that only contain orientation, and see if we obtain the parametric response as documented in Kohler et al.
+We can look at the groups that only contain orientation, and see if we obtain the parametric response as documented in Kohler et al.
 (2016).
-We one might predict based on those results, we get a corresponding pattern of display duration thresholds, with the time taken to detect the symmetry dreasing as we increase the amount of rotational symmetry.
+We will also investigate if we see the corresponding pattern with the display duration threshold data, with the time taken to detect the symmetry dreasing as we increase the amount of rotational symmetry.
 
-Curiosuly, we do not have a large difference betwee P4 and P6 in the eeg data.
+```{r, cache = TRUE}
+d_eeg %>% filter(wg %in% c("P2", "P3", "P4", "P6")) %>%
+  mutate(orientation = parse_number(wg))-> d_eeg_rot
 
-```{r, fig.cap = "The effect of increasing the amount of rotational symmetry.", fig.height=3}
-my_samples$eeg %>% select(P2, P3, P4, P6) %>%
-  pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>% 
-  mutate(exp = "eeg") %>%
-  bind_rows(
-    my_samples$threshold %>% select(P2, P3, P4, P6) %>% 
-      pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>%
-      mutate(exp = "display thresholds")) %>% 
-  ggplot() + aes(x = value, fill = `wallpaper group`) + 
-    geom_density(alpha = 0.35) + 
-    facet_wrap(~ exp, scale = "free_y") + 
-  theme_minimal() + 
-  scale_fill_viridis_d()
+m_eeg_rot <- brm(data = d_eeg_rot, rms ~ orientation + (1|subject), family = "lognormal")
+
+summary(m_eeg_rot)
+
+d_eeg_rot %>% modelr::data_grid(orientation) %>%
+  add_fitted_draws(m_eeg_rot, re_formula = NA, n=200) %>%
+  ggplot() + 
+  geom_line(data = d_eeg_rot, aes(x = orientation, y = rms, group = subject), alpha = 0.1, colour = "darkred") +
+  geom_path(aes(x = orientation, y = .value, group = .draw), alpha = .1, colour = "royalblue2") + 
+  theme_minimal() -> p1
+```
+```{r, cache = TRUE}
+d_dispthresh %>% filter(wg %in% c("P2", "P3", "P4", "P6")) %>%
+  mutate(orientation = parse_number(wg))-> d_disp_rot
+
+m_disp_rot <- brm(data = d_disp_rot, threshold ~ orientation + (1|subject), family = "lognormal")
+
+summary(m_disp_rot)
+
+d_disp_rot %>% modelr::data_grid(orientation) %>%
+  add_fitted_draws(m_disp_rot, re_formula = NA, n=200) %>%
+  ggplot() + 
+  geom_line(data = d_disp_rot, aes(x = orientation, y = threshold, group = subject), alpha = 0.1, colour = "darkred") +
+  geom_path(aes(x = orientation, y = .value, group = .draw), alpha = .1, colour = "royalblue2") + 
+  theme_minimal() -> p2
+```
+
+```{r}
+p1 + p2
+
 ```

### 2020-07-23 Writeup/manuscript.tex
some sort of contribution

@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1]{Peter J. Kohler}
-\author[b]{Oda \r{A}sgard} 
-\author[b]{Alasdair D. F. Clarke}
+\author[a, b, 1]{Peter J. Kohler}
+\author[c]{Alasdair D. F. Clarke}
 
-\affil[a]{Stanford University, Department of Psychology}
-\affil[b]{University of Essex, Department of Psychology, Colchester, UK, CO4 3SQ}
+\affil[a]{York University, Department of Psychology, Toronto, ON M3J 1P3, Canada}
+\affil[b]{Centre for Vision Research, York University, Toronto, ON, M3J 1P3, Canada}
+\affil[c]{Stanford University, Department of Psychology, Stanford, CA 94305, United States}
+\affil[d]{University of Essex, Department of Psychology, Colchester, UK, CO4 3SQ}
 
@@
 % Please add here a significance statement to explain the relevance of your work
-\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself and is required for all research papers.}
+\significancestatement{Wallpaper groups were discovered in the mid-19th century, and the 17 groups constitute the complete set of possible ways of regularly tiling the 2D-plane.
+In recent years wallpaper groups have found use in the vision science community, as an ideal stimulus set for studying the perception of symmetries in textures.
+Here we present brain imaging and psychophysical data on the complete set of wallpaper groups and show the hierarchical organization among wallpaper groups in reflected in both visual cortex activity and performance on a symmetry detection task.}
 
@@
 These 17 “wallpaper” groups \cite{RN1562,RN1563,RN1425} obey a hierarchy, determined by mathematical group theory, where simpler groups are subgroups of more complex ones\cite{RN1711}.
-Here we measure symmetry-specific Steady-State Visual Evoked Potentials (SSVEPs), and find that the partial ordering of SSVEPs from visual cortex in human participants ($n=25$), preserve the hierarchical relationships between the wallpaper groups in $88.3\%$ (53 out of 60) of cases.
+Here we probe representations of symmetries in wallpaper groups using two methods: Steady-State Visual Evoked Potentials (SSVEPs) recorded using EEG and psychophysical detection thresholds.
+We find that the hierarchical relationships between the wallpaper groups are almost perfectly preserved in both behavior and response amplitudes in visual cortex.
 This remarkable consistency between the structure of symmetry representations and mathematical group theory, is likely generated over visual development, through implicit learning of regularities in the environment.
@@
 \centering
-\includegraphics[width=1\linewidth]{../ac_models_eeg_data/plots/subgroup_comprms_Occ_Odd.png}
+\includegraphics[width=1\linewidth]{../figures/figure1.png}
+\caption{Posterior distributions for the difference in mean RMS EEG response.
+Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
+We can see that 55/64 subgroup relationships have $p(\Delta|data)>0.9$.}
+\label{fig:durations_rotations}
+\end{figure}
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comprms_Occ_Odd.png}
 \caption{Posterior distributions for the difference in mean RMS EEG response.

### 2020-07-23 analysis/analysis.Rmd
fig 3 and 4 for paper

@@
   xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")
+
+ggsave("plots/model_roc_style.pdf", height = 5, width = 8)
 ```

### 2020-07-29 Writeup/manuscript.tex
significance, abstract, intro

@@
 
-\title{Template for preparing your research report submission to PNAS using Overleaf}
+\title{The human visual system preserves the hierarchy of 2-dimensional pattern regularity}
 
@@
 In recent years wallpaper groups have found use in the vision science community, as an ideal stimulus set for studying the perception of symmetries in textures.
-Here we present brain imaging and psychophysical data on the complete set of wallpaper groups and show the hierarchical organization among wallpaper groups in reflected in both visual cortex activity and performance on a symmetry detection task.}
+Here we present brain imaging and psychophysical data on the complete set of wallpaper groups and show the hierarchical organization among wallpaper groups in reflected in both representations in visual cortex and performance on a symmetry detection task.
+This shows that the visual system is highly sensitive to reguarities in textures, and suggest that symmetries may play an important role in texture perception.}
 
 % Please include corresponding author, author contribution and author declaration information
-\authorcontributions{Please provide details of author contributions here.}
-\authordeclaration{Please declare any conflict of interest here.}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: author.two\@email.com}
+\authorcontributions{PJK and ADFC designed the study, PJK collected EEG data, ADFC collected psychophysical data, PJK and ADFC wrote the paper.}
+\authordeclaration{The authors have no conflicts of interests to declare}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: \href{mailto:pjkohler@syorku.ca}{pjkohler@yorku.ca}}
 
@@
 In a 2-D image, the four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in 17 distinct ways.
-These 17 “wallpaper” groups \cite{RN1562,RN1563,RN1425} obey a hierarchy, determined by mathematical group theory, where simpler groups are subgroups of more complex ones\cite{RN1711}.
-Here we probe representations of symmetries in wallpaper groups using two methods: Steady-State Visual Evoked Potentials (SSVEPs) recorded using EEG and psychophysical detection thresholds.
-We find that the hierarchical relationships between the wallpaper groups are almost perfectly preserved in both behavior and response amplitudes in visual cortex.
+These 17 “wallpaper” groups \cite{RN1562,RN1563,RN1425} obey a hierarchy, determined by mathematical group theory, in which simpler groups are subgroups of more complex ones\cite{RN1711}.
+Here we probe representations of symmetries in wallpaper groups using two methods: (1) Steady-State Visual Evoked Potentials (SSVEPs) recorded using EEG and (2) symmetry detection thresholds measured psychophysically.
+We find that hierarchical relationships between the wallpaper groups are almost perfectly preserved in both behavior and response amplitudes in visual cortex.
 This remarkable consistency between the structure of symmetry representations and mathematical group theory, is likely generated over visual development, through implicit learning of regularities in the environment.
@@
 Since then, symmetry has been shown to contribute to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection \cite{RN1337}.
-Most of this work has focused on mirror symmetry,
-
-In the two spatial dimensions relevant for images, the four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in 17 distinct ways, the “wallpaper” groups \cite{RN1562,RN1563,RN1425}.
+Most of this work has focused on mirror symmetry or \textit{reflection}, with much less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection}.
+In the two spatial dimensions relevant for images, these four fundamental symmetries can be combined in 17 distinct ways, the “wallpaper” groups \cite{RN1562,RN1563,RN1425}.
+Previous work has focused on four of the wallpaper groups, and used functional MRI to show that rotation symmetries within wallpapers are represented parametrically in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}.
+This effect is also robust in electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\cite{RN1725} or event-related paradigms \cite{RN1959}.
+Here we extend on this work by collecting SSVEPs and psychophysical data from human participants viewing the complete set of wallpaper groups.
+We measure responses in visual cortex to 16 out of the 17 wallpaper groups, with the 17th serving as a control stimulus, with the goal of providing a more complete picture of how wallpaper groups are represented in the human visual system.
 The 17 wallpaper groups obey a hierarchy of complexity, determined by mathematical group theory, where simpler groups are sub-groups of more complex ones \cite{RN1711}.
-Here we use Steady-State Visual Evoked Potentials (SSVEPs) to measure responses to 16 out of 17 wallpaper groups and show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups.
-Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases.
-Visual cortex thus encodes all the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.
+The two datasets presented here puts on in the position of being able to assess the extent to which both behavior and brain responses follow that hierachy.
+The results show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds preserving these relationships at a level that is far beyond chance.
+Visual cortex thus appears to encode all of the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.
 Given that most participants had no knowledge of group theory, the ordered structure of visual responses to wallpaper groups is likely learned implicitly from regularities in the visual environment.
@@
 Because the P1 group serves a control stimulus in this approach, the experiment was restricted to the remaining 16 groups.
-This design allows us to isolate responses to structural features beyond the shared power spectrum, including any symmetries other than translation, in the odd harmonics of the image update frequency \cite{RN1725,RN1540} [NORCIA 2002 MISSING].
+This design allows us to isolate responses to structural features beyond the shared power spectrum, including any symmetries other than translation, in the odd harmonics of the image update frequency \cite{RN1725,RN1540, RN1954}.
 Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1.
@@
 If wallpapers are super-textures, their prevalence may be a direct consequence of the strategy the human visual system uses for encoding visual textures.
+
+\section*{snippets}
+Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases.
 

### 2020-07-29 Writeup/manuscript.tex
Fixed some typos

@@
 Their informed consent was obtained before the experiment under a protocol that was approved by the Institutional Review Board of Stanford University.
-11 participants (8 females mean age 20.73±1.21) took part in the psychophysics experiment.
+11 participants (8 females, mean age 20.73±1.21) took part in the psychophysics experiment.
 All participants had normal or corrected-to-normal vision.
@@
 The rest of the block consisted of four interleaved staircases (using the QUEST algorithm \cite{watson1983quest}, full details given in the SI) of 30 trials each.
-on average, a block of trials took around 10 minutes to complete.
+On average, a block of trials took around 10 minutes to complete.
 
@@
 \subsection*{Bayesian Analysis of EEG and Psychophysical data}
-Bayesian analysis was carried out using R (v3.6.1) \cite{R} with the \texttt{brms} package (v2.9.0 \cite{burkner2017} and rStan (v2.19.2 \cite{rStan}).
+Bayesian analysis was carried out using R (v3.6.1) \cite{R} with the \texttt{brms} package (v2.9.0) \cite{burkner2017} and rStan (v2.19.2 \cite{rStan}).
 The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16 level factor, and random effects for participant.
-The EEG data and display thresholds were modelled using log-normal distributions with weakly informative ($N(0, 2)$) priors.
+The EEG data and display thresholds were modelled using log-normal distributions with weakly informative, $ \mathcal{N}(0, 2)$, priors.
 After fitting the model to the data, samples were drawn from the posterior distribution for each mean of the EEG response (display duration) for each wallpaper group.
 These samples were then recombined to calculate the distribution of differences for each pair of subgroup and super-group.
-These distributions were then summarised by computing the conditional probability that of obtaining a positive (negative) difference, $p(\Delta | data)$.
+These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | data)$.
 
-For further technical details, please see the supplementary materials where the full R code and model specification can be found.
+For further technical details, please see the supplementary materials where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
 

### 2020-07-30 Writeup/manuscript.tex
updated results

@@
 We measure responses in visual cortex to 16 out of the 17 wallpaper groups, with the 17th serving as a control stimulus, with the goal of providing a more complete picture of how wallpaper groups are represented in the human visual system.
-The 17 wallpaper groups obey a hierarchy of complexity, determined by mathematical group theory, where simpler groups are sub-groups of more complex ones \cite{RN1711}.
-The two datasets presented here puts on in the position of being able to assess the extent to which both behavior and brain responses follow that hierachy.
-The results show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds preserving these relationships at a level that is far beyond chance.
-Visual cortex thus appears to encode all of the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.
-Given that most participants had no knowledge of group theory, the ordered structure of visual responses to wallpaper groups is likely learned implicitly from regularities in the visual environment.
 
-\section*{Results}
-
-The visual stimuli for our experiment were multiple exemplar images belonging to each of the 17 wallpaper groups, generated from random-noise textures, as described in detail elsewhere \cite{RN1725}.
-To isolate brain activity specific to the symmetry structure in the images from activity associated with modulation of low-level features, we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group.
-Because all wallpapers are periodic images due to their lattice tiling structure, the phase-scrambled images are also a wallpaper group (P1).
-P1 contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \cite{RN1425}.
-Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of P1 exemplars, that were matched in terms of their Fourier power spectrum.
-Because the P1 group serves a control stimulus in this approach, the experiment was restricted to the remaining 16 groups.
-This design allows us to isolate responses to structural features beyond the shared power spectrum, including any symmetries other than translation, in the odd harmonics of the image update frequency \cite{RN1725,RN1540, RN1954}.
-Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1.
 A wallpaper group is a topologically discrete group of isometries of the Euclidean plane, i.e.
@@
 Because P2 is repeated three times in P6, P2 is a subgroup of P6 with index 3 \cite{RN1425}.
+The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are sub-groups of more complex ones \cite{RN1711}.
+
+The two datasets presented here puts on in the position of being able to assess the extent to which both behavior and brain responses follow that hierachy.
+The results show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds preserving these relationships at a level that is far beyond chance.
+Visual cortex thus appears to encode all of the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.
+Given that most participants had no knowledge of group theory, the ordered structure of visual responses to wallpaper groups is likely learned implicitly from regularities in the visual environment.
+
+\section*{Results}
+
+The stimuli used in our two experiments were multiple exemplar images belonging to each of the wallpaper groups, generated from random-noise textures, as described in detail elsewhere \cite{RN1725}.
+Exemplar images from group P1 was used as control stimuli, and each exemplar from the other 16 groups had a power-spectrum matched P1 exemplar.
+The matched P1 exemplars were generated by phase-scrambling the exemplar images.
+Because all wallpapers are periodic due to their lattice tiling structure, the phase-scrambled images all belong to group P1 regardless of group membership of the original exemplar.
+P1 contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \cite{RN1425}.
+In our SSVEP experiment, this stimulus set allowed us to isolate brain activity specific to the symmetry structure in the exemplar images from activity associated with modulation of low-level features, by alternating exemplar images and control exemplars.
+In this design, responses to structural features beyond the shared power spectrum, including any symmetries other than translation, are isolated in the odd harmonics of the image update frequency \cite{RN1725,RN1540, RN1954}.
+Thus, the combined magnitude of the odd harmonic response components can be used as a measure of the overall strength of the visual cortex response.
+
+The psychophysical experiment took a distinct but related approach.
+In each trial an exemplar image was shown with its matched control, one image after the other, and the order varied pseudo-randomly such that in half the trials the original exemplar was shown first, and in the other half the control image was shown first.
+After each trial, participants were told to indicate whether the first or second image contained more structure, and the duration of both images was controlled by a staircase procedure so that a threshold duration could be computed for each wallpaper group.
+
+A summary of our results can be seen in Figure 1 ...
 
@@
 Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases.
+
+we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group. 
+Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of P1 exemplars, that were matched in terms of their Fourier power spectrum.
+
+Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1.
 

### 2020-07-30 analysis/analysis.Rmd
updated figures

--- 2020-06-24 analysis/analysis.Rmd
@@
   mutate(proportion = number/nrow(subgroups)) %>%
-  ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
-  theme_minimal() + ggthemes::scale_color_ptol(labels = c("primary EEG", "sensor control EEG", "harmonic control EEG", "psychophysical threshold")) +
+  ggplot(aes(x = p, y = proportion, colour = measurement)) + 
+  geom_path(size = 1.5) +
+  theme_minimal() + 
+  ggthemes::scale_color_ptol(labels = c("primary EEG", "sensor control EEG", "harmonic control EEG", "psychophysical threshold")) +
+  theme(legend.justification=c(0,0), legend.position=c(0,0), legend.background = element_rect()) +
   scale_y_continuous(limits = c(0, 1)) +
   xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")
+
+ggsave("plots/model_roc_style.pdf", height = 4, width = 4)
 ```

### 2020-07-30 Writeup/manuscript.tex
updated figures
adjusting figure margins and dimensions

@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics[width=1\linewidth]{../figures/figure1.png}
+\caption{Examplar stimulus for each of the 16 wallpaper groups, with corresponding <something EEG> and a boxplot showing the distribution of display duration thresholds. }
+\label{fig:durations_rotations}
+\end{figure}
+
 The two datasets presented here puts on in the position of being able to assess the extent to which both behavior and brain responses follow that hierachy.
@@
 \centering
-\includegraphics[width=1\linewidth]{../figures/figure1.png}
+\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.pdf}
 \caption{Posterior distributions for the difference in mean RMS EEG response.
@@
 \centering
-\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comprms_Occ_Odd.png}
+\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_psychophysical.pdf}
+\caption{Posterior distributions for the difference in mean display duration threshold.
+Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
+We can see that 55/64 subgroup relationships have $p(\Delta|data)>0.9$.}
+\label{fig:durations_rotations}
+\end{figure}
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=1\linewidth]{../analysis/plots/model_roc_style.pdf}
 \caption{Posterior distributions for the difference in mean RMS EEG response.

### 2020-07-30 analysis/analysis.Rmd
adjusting figure margins and dimensions

@@
 The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being greater than zero.", cache=TRUE}
+The fill of the density plots indicated the probability of the difference being greater than zero."}
 
@@
 both_plts
-
+ggsave("plots/subgroup_comp_eeg_rms.pdf", width = 6, height = 9)
 ```
@@
 
-```{r ,fig.height=8, cache = TRUE, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups.
+```{r ,fig.height=8, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups.
 The index of each relationship is indicated by the colour of the y-axis label.
@@
 both_plts
+ggsave("plots/subgroup_comp_psychophysical.pdf", width = 6, height = 6)
 ```
@@
 
-```{r fig.height=8, cache=TRUE, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
+```{r fig.height=8, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
 The index of each relationship is indicated by the colour of the y-axis label.

### 2020-07-30 analysis/scripts/get_subgroup_comparisons.R

@@ -210,7 +210,7 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
             face = if_else(l$normal==0, "bold", "plain"),
             colour = l$lab_cols),
             axis.title.y = element_blank(),
-            plot.margin=unit(c(1,1,1.5,1.2),"cm"))  
+            plot.margin=unit(c(0,0,0,0),"cm"))  
        
      if (is_thresholds) {
         plt <- plt + coord_cartesian((xlim = c(-2.2, 0.7))) + 

### 2020-07-30 Writeup/manuscript.tex
resizing figures

@@
 
-\begin{figure}%[tbhp]
+\begin{figure}[tbhp]
 \centering
-\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.pdf}
+\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.png}
 \caption{Posterior distributions for the difference in mean RMS EEG response.
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that 55/64 subgroup relationships have $p(\Delta|data)>0.9$.}
-\label{fig:durations_rotations}
+We can see that xx/64 subgroup relationships have $p(\Delta|data)>0.9$.}
+\label{fig:eeg_comp}
 \end{figure}
 
-\begin{figure}%[tbhp]
+\begin{figure}[tbhp]
 \centering
-\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_psychophysical.pdf}
+\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_psychophysical.png}
 \caption{Posterior distributions for the difference in mean display duration threshold.
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that 55/64 subgroup relationships have $p(\Delta|data)>0.9$.}
-\label{fig:durations_rotations}
-\end{figure}
-
-\begin{figure}%[tbhp]
-\centering
-\includegraphics[width=1\linewidth]{../analysis/plots/model_roc_style.pdf}
-\caption{Posterior distributions for the difference in mean RMS EEG response.
-Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that 55/64 subgroup relationships have $p(\Delta|data)>0.9$.}
-\label{fig:durations_rotations}
+We can see that xx/64 subgroup relationships have $p(\Delta|data)>0.9$.}
+\label{fig:durations_comp}
 \end{figure}
@@
 If wallpapers are super-textures, their prevalence may be a direct consequence of the strategy the human visual system uses for encoding visual textures.
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=1\linewidth]{../analysis/plots/model_roc_style.pdf}
+\caption{}}
+\label{fig:durations_rotations}
+\end{figure}
 

### 2020-07-30 analysis/analysis.Rmd
resizing figures

@@
 both_plts
-ggsave("plots/subgroup_comp_eeg_rms.pdf", width = 6, height = 9)
+ggsave("plots/subgroup_comp_eeg_rms.png", width = 5.5, height = 8)
 ```
@@
 both_plts
-ggsave("plots/subgroup_comp_psychophysical.pdf", width = 6, height = 6)
+ggsave("plots/subgroup_comp_psychophysical.png",  width = 5.5, height = 8)
 ```

### 2020-07-30 analysis/scripts/get_subgroup_comparisons.R

@@ -200,7 +200,7 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
         geom_line() +       
         geom_density_ridges(colour = "black", alpha = 0.75, bandwidth = 0.05) +
         ggstance::geom_linerangeh(colour = "black", size = 0.5, aes(y = key, xmin= -5, xmax= 5)) + 
-        scale_x_continuous('pdf for difference between subgroups', expand = c(0,0), breaks = seq(-2, 1, 0.5)) +
+        scale_x_continuous('pdf for subgroup difference', expand = c(0,0), breaks = seq(-2, 1, 0.5)) +
         scale_y_discrete(labels = function(x) TeX(levels(d$key))) + #labels = lapply(levels(subgroup_comp$key), TeX)
         theme_minimal() +        
         ggthemes::scale_colour_ptol(drop = FALSE) +
@@ -210,7 +210,7 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
             face = if_else(l$normal==0, "bold", "plain"),
             colour = l$lab_cols),
             axis.title.y = element_blank(),
-            plot.margin=unit(c(0,0,0,0),"cm"))  
+            plot.margin=unit(c(0,0.25,0,0.25),"cm"))  
        
      if (is_thresholds) {
         plt <- plt + coord_cartesian((xlim = c(-2.2, 0.7))) + 
@@ -221,9 +221,9 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
     }
 
     if (fig_n == 1) {
-        plt <- plt + guides(colour = FALSE, fill = guide_legend(title = element_blank()))
+        plt <- plt + guides(colour = FALSE, fill = guide_legend(title = element_blank(), nrow = 2))
     } else {
-        plt <- plt + guides(fill = FALSE, colour = guide_legend(title.postion = "top"))
+        plt <- plt + guides(fill = FALSE, colour = guide_legend(title.postion = "top", nrow = 2))
     }
 
     

### 2020-08-31 Writeup/manuscript.tex
Results almost done, Discussion halfway done

@@
 The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are sub-groups of more complex ones \cite{RN1711}.
+The full set of subgroup relationships is listed in Section 1.4.2 of the Supplementary Material.
 
@@
 \includegraphics[width=1\linewidth]{../figures/figure1.png}
-\caption{Examplar stimulus for each of the 16 wallpaper groups, with corresponding <something EEG> and a boxplot showing the distribution of display duration thresholds. }
+\caption{Examples of each of the 16 wallpaper groups are shown in the left- and right-most column of the figures, next to the corresponding SSVEP (center-left) and psychological (center-right) data from each group.
+The SSVEP data are odd-harmonic-filtered cycle-average waveforms.
+In each cycle, a P1 exemplar was shown for the first 600 ms, followed by the original exemplar for the last 600 ms.
+Errorbars are standard error of the mean.
+Psychophysical data are presented as boxplots reflecting the distribution of display duration thresholds.
+The 16 groups are ordered by the strength of the SSVEP response, to highlight the range of response amplitudes.}
 \label{fig:durations_rotations}
@@
 In each trial an exemplar image was shown with its matched control, one image after the other, and the order varied pseudo-randomly such that in half the trials the original exemplar was shown first, and in the other half the control image was shown first.
-After each trial, participants were told to indicate whether the first or second image contained more structure, and the duration of both images was controlled by a staircase procedure so that a threshold duration could be computed for each wallpaper group.
+After each trial, participants were told to indicate whether the first or second image contained more structure, and the duration of both images was controlled by a staircase procedure so that a display duration threshold could be computed for each wallpaper group.
 
-A summary of our results can be seen in Figure 1 ...
+A summary of our brain imaging and psychophysical measurements is presented with examples of the wallpaper groups in Figure 1.
+For our primary SSVEP analysis, we only considered EEG data from a pre-determined region-of-interest (ROI) consisting of six electrodes over occipital cortex (see Supplementary Figure 1.1).
+SSVEP data from this ROI was filtered so that only the odd harmonics that capture the symmetry response contribute to the waveforms.
+While waveform amplitude is quite variable among the 16 groups, all groups have a sustained negative-going response that begins at about the same time for all groups, ~180 ms after the transition from the P1 control exemplar to the original exemplar.
+To reduced the amplitude of the symmetry-specific response to a single number that could be used in further analyses and compared to the psychophysical data, we computed the root-mean-square (RMS) over the odd-harmonic-filtered waveforms.
+The data in Figure 1 are shown in descending order according to RMS.
+The psychophysical results, shown in box plots in Figure 1, were also quite variable between groups, and there seems to be a general pattern where wallpaper groups near the top of the figure, that have lower SSVEP amplitudes, also have longer psychophysical display duration thresholds.
+
+We now wanted to quantify the degree to which our two measurements were consistent with the subgroup relationships.
+We hypothesized that more complex groups would (a) produce symmetric-specific SSVEPs with higher amplitudes and (b) shorter display duration thresholds.
+[PERHAPS WE NEED TO JUSTIFY THESE ASSUMPTIONS] We tested each of these hypotheses using the same approach.
+We first fitted a Bayesian model with wallpaper group as a factor and participant as a random effect.
+We fit the model separately for SSVEP RMS and psychophysical data, and then computed posterior distributions for the difference between supergroup and subgroup.
+These difference distributions could allowed us to compute the conditional probability that the supergroup would produce (a) larger RMS and (b) a shorter threshold durations, when compared to the subgroup.
+The posterior distributions are shown in Figure 2 for the SSVEP data, and in Figure 3 for the psychophysical data, which distributions color-coded according to conditional probability.
+For both data sets our hypothesis is confirmed: For the overwhelming majority of the 64 subgroup relationships, supergroups are more likely to produce larger symmetry specific SSVEPs and shorter threshold durations, and in most cases the conditional probability of this happening is extremely high.
+
+We also ran a control analysis using (1) odd-harmonic SSVEP data from a six-electrode ROI over parietal cortex (see Supplementary Figure 1.1) and (2) even-harmonic SSVEP data from the same occipital ROI that was used in our primary analysis.
+By comparing these two control analysis to our primary SSVEP analysis, we can address the specify of our effects in terms of location (occipital cortex vs parietal cortex) and harmonic (odd vs even).
+For both control analyses (plotted in Supplementary Figures 3.3 and 3.4), the correspondence between data and subgroup relationships was weaker than in the primary analysis.
+We can quantify the strength of the assocation between the data and the subgroup relationships, by asking what proportion of subgroup relationships that reach or exceed a range of probability thresholds.
+This is plotted in Figure 4, for our psychophysical data, our primary SSVEP analysis and our two control SSVEP analyses.
+It it that odd-harmonic SSVEP data from an occipital ROI and display duration thresholds both have a strong association with the subgroup relations, that for a clear majority of the subgroups survive even at the highest threshold we consider ($p(\Delta|data)>0.99$), and that the assocation is far weaker for the two control analyses.
+
+A subset of the SSVEP data was previously published as part of an earlier demonstration of parametric responses to rotation symmetry in wallpaper groups \cite{RN1725}.
+We replicate that result using our Bayesian approach, and find the same parametric effect in the psychophysical data (Supplementary Figure 4.1 - NEEDS TO BE LABELED IN THE SM).
+We also conducted analyses looking for effects of index and normality in our two datasets.
+
+We conclude that the hierarchy of complexity among wallpaper groups is reflected in precise detail in both brain responses and performance in a psychophysical task.
 
@@
 \includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.png}
-\caption{Posterior distributions for the difference in mean RMS EEG response.
+\caption{Posterior distributions for the difference in mean RMS SSVEP response.
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
@@
 The RMS measure of SSVEP amplitude, preserves the complex hierarchy of subgroup relationships among the wallpaper groups \cite{RN1711}.
-Out of a total of 60 relationships, 53 were preserved in a significant number of participants, and 49 were significant even at a stricter threshold (p < 0.002).
-The ordering was highly stable in individual participants, with an average preservation rate of ~21 of 25 participants across all 60 relationships (see Figure 3).
 This remarkable consistency was specific to the odd harmonics of the stimulus frequency, that capture the symmetry-specific response \cite{RN1725} and to electrodes in an ROI over occipital cortex.
-When the same analysis was done on the even harmonics of the occipital cortex ROI, the ordering of responses was much less apparent (see Figure S2) and preservation rates much lower (see Figure S4).
-The odd harmonics from electrodes in an ROI over parietal cortex, showed even weaker evidence of preserving the hierarchy among sub-groups (see Figure S5).
-Importantly, no relationships were preserved in either of these control analyses that were not also preserved in the main analysis of the odd harmonics in the occipital cortex ROI.
+When the same analysis was done using the odd harmonics from a parietal cortex ROI (Supplementary Figure 3.4) or using the even harmonics of the occipital cortex ROI (Supplementary Figure 3.4), the data was much less consistent with the subgroup relationships (yellow and green lines, Figure 4).
+
 The current data provide a complete description of the visual system’s response to symmetries in the 2-D plane.
-Our design does not allow us to independently measure the response to P1, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 2), we can conclude that the relationships between P1 and all other groups, where P1 is the subgroup, are also preserved by the visual system.
+Our design does not allow us to independently measure the response to P1, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where P1 is the subgroup, are also preserved by the visual system.
 The subgroup relationships are not obvious perceptually, and most participants had no knowledge of group theory.
@@
 
-\begin{figure}%[tbhp]
+\begin{figure}[tbhp]
 \centering
@@
 \end{figure}
-
-\section*{snippets}
-Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases.
-
-we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group. 
-Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of P1 exemplars, that were matched in terms of their Fourier power spectrum.
-
-Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1.
 
@@
 
-\subsection*{EEG Analysis}
+\subsection*{SSVEP Analysis}
 Response waveforms were generated for each group by selective filtering in the frequency domain.
 For each participant, the average Fourier coefficients from the two sessions were averaged over trials and sessions.
-The Steady-State Visual Evoked Potentials paradigm we used allowed us to separate symmetry-related responses from non-specific contrast transient responses.
+The Steady-State Visual Evoked Potentials (SSVEP) paradigm we used allowed us to separate symmetry-related responses from non-specific contrast transient responses.
 Previous work has demonstrated that symmetry-related responses are predominantly found in the odd harmonics of the stimulus frequency, whereas the even harmonics consist mainly of responses unrelated to symmetry, that arise from the contrast change associated with the appearance of the second image[2-4].
@@
 
+\section*{snippets}
+Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases.
+
+we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group. 
+Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of P1 exemplars, that were matched in terms of their Fourier power spectrum.
+
+Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1.
+
 \showmatmethods{} % Display the Materials and Methods section

### 2020-09-01 Writeup/manuscript.tex
updated results

@@
 
-A subset of the SSVEP data was previously published as part of an earlier demonstration of parametric responses to rotation symmetry in wallpaper groups \cite{RN1725}.
+SSVEP data from four of the wallpaper groups (P2, P3, P4 and P6) was previously published as part of our earlier demonstration of parametric responses to rotation symmetry in wallpaper groups\cite{RN1725}.
 We replicate that result using our Bayesian approach, and find the same parametric effect in the psychophysical data (Supplementary Figure 4.1 - NEEDS TO BE LABELED IN THE SM).
-We also conducted analyses looking for effects of index and normality in our two datasets.
-
-We conclude that the hierarchy of complexity among wallpaper groups is reflected in precise detail in both brain responses and performance in a psychophysical task.
+We also conducted analyses looking for effects of index and normality in our two datasets, and found that ...
+Finally, we conducted a correlation analysis comparing SSVEP and behavioral data, and found a small ($R^2 = 0.44$) but above-zero correlation, as indicated by our confidence intervals.
+There are several factors that might explain the relatively weak correlation, most prominently the fact that the same individuals did not participate in each of the two experiments.
+Nevertheless, we find the relationship between the two datasets interesting, because it suggests that our psychophysical and SSVEP measurements are tapping into the same underlying mechanisms.
 

### 2020-09-01 analysis/analysis.Rmd
Some small edits to supp mat

@@
 ```{r}
-d_eeg %>% group_by(subject) %>% summarise(count = n()) %>% knitr::kable()
+d_eeg %>% group_by(subject) %>% 
+  summarise(count = n()) %>% 
+  knitr::kable() %>% 
+  kableExtra::kable_styling(full_width = F)
 ```
@@
 
-```{r fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals.", cache=TRUE}
+```{r fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals.", cache=TRUE, warning=FALSE}
 
@@
 
-filter(df, p == 0.95)
+filter(df, p == 0.95) %>% knitr::kable() %>% kableExtra::kable_styling(full_width = F)
 
@@
   ggplot() + 
-  geom_line(data = d_disp_rot, aes(x = orientation, y = threshold, group = subject), alpha = 0.1, colour = "darkred") +
+  geom_line(data = d_disp_rot, aes(x = orientation, y = threshold, group = subject), alpha = 0.1, colour = "red") +
   geom_path(aes(x = orientation, y = .value, group = .draw), alpha = .1, colour = "royalblue2") + 
@@
 
-```{r}
+```{r rotations-replication, fig.cap="Red lines show empirical data, blue lines show the model fit."}
 p1 + p2
-
 ```

### 2020-09-01 analysis/scripts/get_subgroup_comparisons.R

@@ -24,165 +24,6 @@ get_subgroup_comparisons2 <- function(m_samples, n_iter) {
   return(subgroup_comp)
 }
 
-get_subgroup_comparisons <- function(m_samples, n_iter)
-{
-    m_samples %>%
-      as_tibble() %>% 
-      mutate_all(exp) %>%
-      mutate(
-        'P4M$\\rightarrow$ P4G' = P4M - P4G,
-        'P4M$\\rightarrow$ P4'  = P4M - P4,
-        'P4M$\\rightarrow$ CMM' = P4M - CMM,
-        'P4M$\\rightarrow$ PMM' = P4M - PMM, 
-        'P6M$\\rightarrow$ P6' = P6M - P6,
-        'P6M$\\rightarrow$ P31M' = P6M - P31M,
-        'P6M$\\rightarrow$ P3M1' = P6M - P3M1,
-        'PMM$\\rightarrow$ PMG' = PMM - PMG,
-        'PMM$\\rightarrow$ PM' = PMM - PM,
-        'PMM$\\rightarrow$ P2' = PMM - P2,
-        'P3M1$\\rightarrow$ P3'= P3M1- P3,
-        'CMM$\\rightarrow$ PMG' = CMM - PMG,
-        'CMM$\\rightarrow$ PGG' = CMM - PGG,
-        'CMM$\\rightarrow$ CM' = CMM - CM,
-        'CMM$\\rightarrow$ P2' = CMM - P2,
-        'P4G$\\rightarrow$ P4' = P4G - P4,
-        'P4G$\\rightarrow$ CMM' = P4G - CMM,
-        'P4G$\\rightarrow$ PGG' = P4G - PGG,
-        'PMG$\\rightarrow$ PGG' = PMG - PGG, 
-        'PMG$\\rightarrow$ PM' = PMG - PM,
-        'PMG$\\rightarrow$ PG' = PMG - PG,
-        'PMG$\\rightarrow$ P2' = PMG - P2,
-        'P31M$\\rightarrow$ P3'= P31M- P3,
-        'P6$\\rightarrow$ P3'  = P6  - P3,
-        'CM$\\rightarrow$ PG'  = CM  - PG,
-        'P4$\\rightarrow$ P2'  = P4  - P2,
-        'PM$\\rightarrow$ PG'  = PM  - PG, 
-        'PGG$\\rightarrow$ PG' = PGG - PG, 
[…]

### 2020-09-01 analysis/scripts/import_subgroup_comparisons.R

@@ -1,6 +1,5 @@
 library(tidyverse)
 
-
 index_dat  <- read_csv("../subgroups/subgroup_relations.csv")
 normal_dat <- read_csv("../subgroups/subgroup_normal.csv")
 
@@ -30,7 +29,6 @@ for (ii in 1:nrow(index_dat)) {
             normal = normal_dat[ii,jj] == 1)
     }
 } 
-
 
 # list of comparisons to remove
 to_remove <- list(
@@ -71,4 +69,4 @@ subgroups %>% filter(
     is.finite(index),
     group != subgroup,
     !(paste(group, subgroup, sep="-") %in% to_remove)) %>%
-    mutate(label = paste(group, "$\\rightarrow$", subgroup)) -> subgroups
+    mutate(label = paste(group, "% -> %", subgroup)) -> subgroups

### 2020-09-01 analysis/scripts/get_subgroup_comparisons.R
Rightarrows

@@ -28,6 +28,8 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
 {
     d$key <- fct_drop(d$key)
     l$key <- fct_drop(l$key)
+    
+    x_labels <- expression()
 
     prob_relation <- mean(d$value)
 

### 2020-09-01 analysis/scripts/import_subgroup_comparisons.R

@@ -69,4 +69,4 @@ subgroups %>% filter(
     is.finite(index),
     group != subgroup,
     !(paste(group, subgroup, sep="-") %in% to_remove)) %>%
-    mutate(label = paste(group, "% -> %", subgroup)) -> subgroups
+    mutate(label = paste(group, "->", subgroup)) -> subgroups

### 2020-09-01 analysis/analysis.Rmd
Tidy up sup matt a little

@@
 
-Finally, we calculate the probability that the RMS difference between subgroup and supergroup is larger than zero given the data.
+Finally, we calculate the probability that the RMS difference between subgroup and super-group is larger than zero given the data.
 This information is then binned so we can colour in the posterior density plots.
@@
 The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being greater than zero."}
+The fill of the density plots indicated the probability of the difference being greater than zero.", warning=FALSE}
 
@@
 The index of each relationship is indicated by the colour of the y-axis label.
-The fill of the density plots indicated the probability of the difference being less than zero."}
+The fill of the density plots indicated the probability of the difference being less than zero.", warning=FALSE}
 

### 2020-09-01 Writeup/manuscript.tex
The figures are back!
Added figure caption

@@
 
-\templatetype{pnasresearcharticle} % Choose template 
+%\templatetype{pnasresearcharticle} % Choose template 
 % {pnasresearcharticle} = Template for a two-column research article
@@
 % {pnasinvited} %= Template for a PNAS invited submission
+\usepackage{graphicx}
 
@@
 
-\maketitle
+% \maketitle
 \thispagestyle{firststyle}
@@
 
-\begin{figure}[tb]
+\begin{figure}[tbhp]
 \centering
@@
 This is plotted in Figure 4, for our psychophysical data, our primary SSVEP analysis and our two control SSVEP analyses.
-It it that odd-harmonic SSVEP data from an occipital ROI and display duration thresholds both have a strong association with the subgroup relations, that for a clear majority of the subgroups survive even at the highest threshold we consider ($p(\Delta|data)>0.99$), and that the assocation is far weaker for the two control analyses.
+It it that odd-harmonic SSVEP data from an occipital ROI and display duration thresholds both have a strong association with the subgroup relations, that for a clear majority of the subgroups survive even at the highest threshold we consider ($p(\Delta>0|data)>0.99$), and that the assocation is far weaker for the two control analyses.
 
@@
 \includegraphics[width=1\linewidth]{../analysis/plots/model_roc_style.pdf}
-\caption{}}
+\caption{This plot shows the proportion of subgroup relations that satisfy $p(\Delta >0 | data) > x$.
+We can see that if we take $x = 0.95$ as our threshold, the subgroup relations are preserved in 56/63 = 89\% and 49/64 = 78\% of the comparisons for the primary EEG and display durations repectively.
+This compares to the 32/64= 50\% and 22/64 = 35\% for the control EEG conditions.}
 \label{fig:durations_rotations}

### 2020-09-04 analysis/analysis.Rmd
minor edits

@@
 
-The table below shows how many groups per participant we have data for.
-
-```{r}
-d_eeg %>% group_by(subject) %>% 
-  summarise(count = n()) %>% 
-  knitr::kable() %>% 
-  kableExtra::kable_styling(full_width = F)
-```
-
 ## Threshold Data
@@
 ```{r}
-d_dispthresh %>% group_by(subject) %>% summarise(count = n()) %>% knitr::kable()
+d_dispthresh %>% 
+  group_by(subject) %>% 
+  summarise(count = n()) %>% 
+  knitr::kable() %>% 
+  kableExtra::kable_styling(full_width = F)
 ```
@@
 ### Subgroup Relationships
-Import subgroup information and display a table of the relationships that we will be investigating.
+Import subgroup information and display a table of the relationships that we will be investigating. Relationships taken from Coxeter & Moser (1972).
 
@@
 
-Now we will use it!
-
 ```{r, fig.height=8, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
@@
 
-## Replicatation of Kohler et al (2016)
+## Replication of Kohler et al (2016)
 
-We can look at the groups that only contain orientation, and see if we obtain the parametric response as documented in Kohler et al.
-(2016).
+We can look at the groups that only contain rotations, and see if we obtain the parametric response as documented in Kohler et al. (2016).
+
+```{r, cache = TRUE}
+d_eeg %>% filter(wg %in% c("P2", "P3", "P4", "P6")) %>%
+  mutate(rotation = parse_number(wg))-> d_eeg_rot
+
+m_eeg_rot <- brm(data = d_eeg_rot, rms ~ rotation + (1|subject), family = "lognormal")
+
+summary(m_eeg_rot)
+
+d_eeg_rot %>% modelr::data_grid(rotation) %>%
+  add_fitted_draws(m_eeg_rot, re_formula = NA, n=200) %>%
+  ggplot() + 
+  geom_line(data = d_eeg_rot, aes(x = rotation, y = rms, group = subject), alpha = 0.1, colour = "darkred") +
+  geom_path(aes(x = rotation, y = .value, group = .draw), alpha = .1, colour = "royalblue2") + 
+  theme_minimal() -> p1
+```
+
 We will also investigate if we see the corresponding pattern with the display duration threshold data, with the time taken to detect the symmetry dreasing as we increase the amount of rotational symmetry.
@@
 ```{r, cache = TRUE}
-d_eeg %>% filter(wg %in% c("P2", "P3", "P4", "P6")) %>%
-  mutate(orientation = parse_number(wg))-> d_eeg_rot
+d_dispthresh %>% filter(wg %in% c("P2", "P3", "P4", "P6")) %>%
+  mutate(rotation = parse_number(wg))-> d_disp_rot
 
-m_eeg_rot <- brm(data = d_eeg_rot, rms ~ orientation + (1|subject), family = "lognormal")
-
-summary(m_eeg_rot)
-
-d_eeg_rot %>% modelr::data_grid(orientation) %>%
-  add_fitted_draws(m_eeg_rot, re_formula = NA, n=200) %>%
-  ggplot() + 
-  geom_line(data = d_eeg_rot, aes(x = orientation, y = rms, group = subject), alpha = 0.1, colour = "darkred") +
-  geom_path(aes(x = orientation, y = .value, group = .draw), alpha = .1, colour = "royalblue2") + 
-  theme_minimal() -> p1
-```
-```{r, cache = TRUE}
-d_dispthresh %>% filter(wg %in% c("P2", "P3", "P4", "P6")) %>%
-  mutate(orientation = parse_number(wg))-> d_disp_rot
-
-m_disp_rot <- brm(data = d_disp_rot, threshold ~ orientation + (1|subject), family = "lognormal")
+m_disp_rot <- brm(data = d_disp_rot, threshold ~ rotation + (1|subject), family = "lognormal")
 
@@
 
-d_disp_rot %>% modelr::data_grid(orientation) %>%
+d_disp_rot %>% modelr::data_grid(rotation) %>%
   add_fitted_draws(m_disp_rot, re_formula = NA, n=200) %>%
   ggplot() + 
-  geom_line(data = d_disp_rot, aes(x = orientation, y = threshold, group = subject), alpha = 0.1, colour = "red") +
-  geom_path(aes(x = orientation, y = .value, group = .draw), alpha = .1, colour = "royalblue2") + 
+  geom_line(data = d_disp_rot, aes(x = rotation, y = threshold, group = subject), alpha = 0.1, colour = "red") +
+  geom_path(aes(x = rotation, y = .value, group = .draw), alpha = .1, colour = "royalblue2") + 
   theme_minimal() -> p2
@@
 
+summary(my_models$eeg)
 ```
@@
 summary(my_models$disp)
+```
+
+It is unclear from the above summary tables whether either variable has a clear effect. We can use leave-one-out metrics to compare models.
+
+```{r}
+summary(lm(data = comp_summary$threshold, mean_value ~ index * normal))
+summary(lm(data = comp_summary$eeg, mean_value ~ index * normal))
 
@@
 print(bR2)
-
 ```
@@
   ggthemes::scale_colour_ptol("index")
-
 ```

### 2020-09-04 analysis/scripts/import_subgroup_comparisons.R

@@ -69,4 +69,4 @@ subgroups %>% filter(
     is.finite(index),
     group != subgroup,
     !(paste(group, subgroup, sep="-") %in% to_remove)) %>%
-    mutate(label = paste(group, "->", subgroup)) -> subgroups
+    mutate(label = paste(group, "≤", subgroup)) -> subgroups

### 2020-09-11 Writeup/manuscript.tex
Fixing title page
minor edits
wrapping figures.
edits to the Discussion ... close now

@@
 
-%\templatetype{pnasresearcharticle} % Choose template 
+\templatetype{pnasresearcharticle} % Choose template 
 % {pnasresearcharticle} = Template for a two-column research article
@@
 % {pnasinvited} %= Template for a PNAS invited submission
-\usepackage{graphicx}
+% \usepackage{graphicx}
 
@@
 
-% \maketitle
+\maketitle
  \thispagestyle{firststyle}
@@
 
-\begin{figure}[tbhp]
+\begin{figure}[ptb]
 \centering
@@
 The 16 groups are ordered by the strength of the SSVEP response, to highlight the range of response amplitudes.}
-\label{fig:durations_rotations}
+\label{fig:example_stimuli}
 \end{figure}
@@
 We hypothesized that more complex groups would (a) produce symmetric-specific SSVEPs with higher amplitudes and (b) shorter display duration thresholds.
-[PERHAPS WE NEED TO JUSTIFY THESE ASSUMPTIONS] We tested each of these hypotheses using the same approach.
+\textcolor{red}{PERHAPS WE NEED TO JUSTIFY THESE ASSUMPTIONS?} We tested each of these hypotheses using the same approach.
 We first fitted a Bayesian model with wallpaper group as a factor and participant as a random effect.
@@
 SSVEP data from four of the wallpaper groups (P2, P3, P4 and P6) was previously published as part of our earlier demonstration of parametric responses to rotation symmetry in wallpaper groups\cite{RN1725}.
-We replicate that result using our Bayesian approach, and find the same parametric effect in the psychophysical data (Supplementary Figure 4.1 - NEEDS TO BE LABELED IN THE SM).
-We also conducted analyses looking for effects of index and normality in our two datasets, and found that ...
-Finally, we conducted a correlation analysis comparing SSVEP and behavioral data, and found a small ($R^2 = 0.44$) but above-zero correlation, as indicated by our confidence intervals.
+We replicate that result using our Bayesian approach, and find the same parametric effect in the psychophysical data (Supplementary Figure 4.1).
+We also conducted an analysis \textcolor{red}{DOES THIS NEED TO BE SPELLED OUT MORE} looking for effects of index and normality in our two datasets, and found that subgroup relationships with higher indices tended to produce stronger effects for both SSVEP RMS and symmetry detection thresholds, as shown by the fact that a leave-one-out test \textcolor{red}{...
+MORE DETAILS NEEDED HERE}.
+The effect of index did not (\textcolor{red}{???} reach significance.
+
+Finally, we conducted a correlation analysis comparing SSVEP and psychophysical data, and found a small ($R^2 = 0.44$) but above-zero correlation, as indicated by our confidence intervals.
 There are several factors that might explain the relatively weak correlation, most prominently the fact that the same individuals did not participate in each of the two experiments.
@@
 
-\begin{figure}[tbhp]
+\begin{figure}[tbh]
 \centering
@@
 Here we show that beyond merely responding to the elementary symmetry operations of reflection \cite{RN1170} and rotation \cite{RN1725}, the visual system explicitly represents hierarchical structure of the 17 wallpaper groups, and thus the compositions of all four of the fundamental symmetry transformations (rotation, reflection, translation, glide reflection) which comprise regular textures.
-The RMS measure of SSVEP amplitude, preserves the complex hierarchy of subgroup relationships among the wallpaper groups \cite{RN1711}.
-This remarkable consistency was specific to the odd harmonics of the stimulus frequency, that capture the symmetry-specific response \cite{RN1725} and to electrodes in an ROI over occipital cortex.
-When the same analysis was done using the odd harmonics from a parietal cortex ROI (Supplementary Figure 3.4) or using the even harmonics of the occipital cortex ROI (Supplementary Figure 3.4), the data was much less consistent with the subgroup relationships (yellow and green lines, Figure 4).
+The SSVEP amplitude and the symmetry detection threshold both preserve the complex hierarchy of subgroup relationships among the wallpaper groups \cite{RN1711}.
+For the SSVEP, this remarkable consistency was specific to the odd harmonics of the stimulus frequency, that capture the symmetry-specific response \cite{RN1725} and to electrodes in an ROI over occipital cortex.
+When the same analysis was done using the odd harmonics from electrodes over parietal cortex (Supplementary Figure 3.3) or even harmonics from electrodes over occipital cortex (Supplementary Figure 3.4), the data was much less consistent with the subgroup relationships (yellow and green lines, Figure 4).
 
 The current data provide a complete description of the visual system’s response to symmetries in the 2-D plane.
-Our design does not allow us to independently measure the response to P1, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where P1 is the subgroup, are also preserved by the visual system.
-The subgroup relationships are not obvious perceptually, and most participants had no knowledge of group theory.
+Our design precludes us from independently measure the response to P1, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where P1 is the subgroup, are also preserved by the visual system.
+The subgroup relationships are in many cases not obvious perceptually, and most participants had no knowledge of group theory.
 Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships.
-Furthermore, behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, they are generally error-prone when asked to assign exemplar images to the appropriate group \cite{RN172}.
+Furthermore, previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, they are generally error-prone when asked to assign exemplar images to the appropriate group \cite{RN172} \textcolor{red}{ALASDAIR PERHAPS YOU WANT TO SAY MORE OR DIFFERENT HERE}.
+
 The correspondence between responses in the visual system and group theory that we demonstrate here, may reflect a form of implicit learning that depends on the structure of the natural world.
@@
 Our results suggest that the principle of efficient coding extends to a much higher level of structural redundancy – that of symmetries in visual images.
+
 The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective and occlusion.
 Near-regular textures, however abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \cite{RN1519}.
-The correspondence between brain data and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space.
+The correspondence between visual cortex responses and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space.
 This framework resembles norm-based encoding strategies that have been proposed for other stimulus classes, most notably faces \cite{RN435}, and leads to the prediction that adaptation to wallpaper patterns should distort perception of near-regular textures, similar to the aftereffects found for faces \cite{RN1768}.
-Field biologist have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \cite{RN1775}.
+Field biologists have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \cite{RN1775}.
 In the norm-based encoding framework, wallpaper groups can be considered super-textures, exaggerated examples of the near-regular textures that surround us.
@@
 Wallpaper groups are visually compelling and have been widely used in human artistic expression going back to the Neolithic age \cite{RN1949}.
-If wallpapers are super-textures, their prevalence may be a direct consequence of the strategy the human visual system uses for encoding visual textures.
+If wallpapers are in fact super-textures, it would imply that this prevalence may be a direct consequence of the strategy the human visual system uses for encoding textures.
 
-\begin{figure}[tbhp]
+\begin{figure}[tbp]
 \centering
@@
 
-\section*{snippets}
-Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases.
+%\section*{snippets}
+%Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases.
 
-we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group. 
-Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of P1 exemplars, that were matched in terms of their Fourier power spectrum.
+%we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group. 
+%Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of P1 exemplars, that were matched in terms of their Fourier power spectrum.
 
-Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1.
+%Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1.
 

### 2020-09-11 analysis/analysis.Rmd
some house-keeping

@@
 
-## EEG Data
+## SSVEP Data
 
-Importing the primary EEG data set. This is odd-harmonic filtered data from region-of-interest consisting of six electrodes over occipital cortex.
+Importing the primary SSVEP data set. This is odd-harmonic filtered data from region-of-interest consisting of six electrodes over occipital cortex.
 
-```{r, fig.cap = "RMS data from two of the wallpaper groups, P2 and P4M.
+```{r, fig.cap = "SSVEP RMS data from two of the wallpaper groups, P2 and P4M.
 Odd harmonics are shown in A and B, while even harmonic data are shown in C and D, and occipital and parietal regions of interest are shown in dark nad light gray, respectively.
@@
 
-In addition to the primary EEG data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data from six parietal electrodes (see Figure 1.1 and the main paper).
+In addition to the primary SSVEP data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data from six parietal electrodes (see Figure 1.1 and the main paper).
 
@@
 * Extract mcmc samples from the posterior distributions.
-* Use these samples to estimate the difference between sub- and super-groups.
+* Use these samples to estimate the difference between sub- and supergroups.
 
@@
 
-### Fit Model to EEG Data
+### Fit Model to SSVEP Data
 
@@
 
-We will now compute the difference between sub- and super-groups.
+We will now compute the difference between sub- and supergroups.
 
@@
 
-## Primary EEG Data
+## Primary SSVEP Data
 
-Finally, we calculate the probability that the RMS difference between subgroup and super-group is larger than zero given the data.
+Finally, we calculate the probability that the SSVEP RMS difference between subgroup and supergroup is larger than zero given the data.
 This information is then binned so we can colour in the posterior density plots.
@@
 
-```{r, fig.height=8, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
+```{r, fig.height=8, fig.cap="Distributions of the difference in mean log(rms) between sub- and supergroups.
 The index of each relationship is indicated by the colour of the y-axis label.
@@
 
-```{r ,fig.height=8, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups.
+```{r ,fig.height=8, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and supergroups.
 The index of each relationship is indicated by the colour of the y-axis label.
@@
 
-## Control EEG Data
+## Control SSVEP Data
 
@@
 
-```{r fig.height=8, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups.
+```{r fig.height=8, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and supergroups.
 The index of each relationship is indicated by the colour of the y-axis label.
@@
   theme_minimal() + 
-  ggthemes::scale_color_ptol(labels = c("primary EEG", "sensor control EEG", "harmonic control EEG", "psychophysical threshold")) +
+  ggthemes::scale_color_ptol(labels = c("primary SSVEP", "sensor control SSVEP", "harmonic control SSVEP", "psychophysical threshold")) +
   theme(legend.justification=c(0,0), legend.position=c(0,0), legend.background = element_rect()) +
@@
 
-If we take $p$=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/63 = 89% and 49/64 = 78% of the comparisons for the primary EEG and display durations repectively.
-This compares to the 32/64= 50% and 22/64 = 35% for the control EEG conditions.
+If we take $p$=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/63 = 89% and 49/64 = 78% of the comparisons for the primary SSVEP and display durations repectively.
+This compares to the 32/64= 50% and 22/64 = 35% for the control SSVEP conditions.
 
@@
 
-First of all, we run for the eeg rms data.
+First of all, we run for the SSVEP RMS data.
 
@@
 
-## Correlation Between Primary EEG data and Psychophysical Thresholds
+## Correlation Between Primary SSVEP data and Psychophysical Thresholds
 
-Finally, we will investigate whether there is a correlation between the our primary EEG measure (rms amplitude of odd harmonics over occipital cortex) and our display duration thresholds.
+Finally, we will investigate whether there is a correlation between the our primary SSVEP measure (RMS amplitude of odd harmonics over occipital cortex) and our display duration thresholds.
 As our two different measures come from different samples of participants, we are unable to do a direct comparison.

### 2020-09-11 Writeup/manuscript.tex
edits to the intro

@@
 In the two spatial dimensions relevant for images, these four fundamental symmetries can be combined in 17 distinct ways, the “wallpaper” groups \cite{RN1562,RN1563,RN1425}.
-Previous work has focused on four of the wallpaper groups, and used functional MRI to show that rotation symmetries within wallpapers are represented parametrically in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}.
-This effect is also robust in electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\cite{RN1725} or event-related paradigms \cite{RN1959}.
-Here we extend on this work by collecting SSVEPs and psychophysical data from human participants viewing the complete set of wallpaper groups.
-We measure responses in visual cortex to 16 out of the 17 wallpaper groups, with the 17th serving as a control stimulus, with the goal of providing a more complete picture of how wallpaper groups are represented in the human visual system.
+Previous work on a subset of four wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}.
+This effect is also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\cite{RN1725} or event-related paradigms \cite{RN1959}.
+Here we extend this work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups.
+We measure responses in visual cortex to 16 out of the 17 wallpaper groups, with the 17th serving as a control stimulus.
+Our goal is to provide a more complete picture of how wallpaper groups are represented in the human visual system.
 
@@
 
-The two datasets presented here puts on in the position of being able to assess the extent to which both behavior and brain responses follow that hierachy.
-The results show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds preserving these relationships at a level that is far beyond chance.
+The two datasets presented here puts on in the position of being able to assess the extent to which both behavior and brain responses follow that subgroup hierachy.
+Based on previous imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \cite{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
+For the psychophysical data, we hypothesized that more complex groups would lead to shorter symmetry detection thresholds, based on previous data showing that under a fixation presentation time, discriminability increases with the number of symmetry axes in the pattern \cite{wagemans_1991}.
+Our results confirm both hypotheses, and show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds following these relationships at a level that is far beyond chance.
 Visual cortex thus appears to encode all of the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.
@@
 
-We now wanted to quantify the degree to which our two measurements were consistent with the subgroup relationships.
-We hypothesized that more complex groups would (a) produce symmetric-specific SSVEPs with higher amplitudes and (b) shorter display duration thresholds.
-\textcolor{red}{PERHAPS WE NEED TO JUSTIFY THESE ASSUMPTIONS?} We tested each of these hypotheses using the same approach.
+We now wanted to test our two assumptions about how SSVEP amplitudes and duration thresholds would follow subgroup relationships, and thereby quantify the degree to which our two measurements were consistent with group theory.
+We tested each of the two hypotheses using the same approach.
 We first fitted a Bayesian model with wallpaper group as a factor and participant as a random effect.

### 2020-10-01 analysis/analysis.Rmd
Table 1.4 description.

@@
 
-<A short note on what is counted...i.e., so the reader understands that P2 has one rotation of order 2, etc>
+The following table lists the non-translational isometries in the 17 wallpaper groups.
+Isometries are listed up to conjugacy, meaning for instance that two identical rotations about different points are not counted separately, and nor are two reflections about lines which differ by a translation in the wallpaper group.
 

### 2020-10-14 Writeup/manuscript.tex
minor edits

@@
 Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships.
-Furthermore, previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, they are generally error-prone when asked to assign exemplar images to the appropriate group \cite{RN172} \textcolor{red}{ALASDAIR PERHAPS YOU WANT TO SAY MORE OR DIFFERENT HERE}.
+Furthermore, previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, \textcolor{green}{when asked to sort exampler images into visually similar subsets, they tend to make between four and 12 sets.
+It is common for examplars from two distinct symmetry groups to be judged as belonging to the same subset, and also for examplars from the same group to be placed into disjoint similarity subsets.}
 

### 2020-10-14 analysis/analysis.Rmd
minor edits

@@
   group_by(subject) %>% 
-  summarise(count = n()) %>% 
+  summarise(count = n(), .groups = "drop") %>% 
   knitr::kable() %>% 
@@
       mean_value = mean(value),
-      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)))  %>%
+      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)),
+      .groups = "drop")  %>%
     mutate(
@@
 
-It is unclear from the above summary tables whether either variable has a clear effect. We can use leave-one-out metrics to compare models.
-
-```{r}
-summary(lm(data = comp_summary$threshold, mean_value ~ index * normal))
-summary(lm(data = comp_summary$eeg, mean_value ~ index * normal))
-
-```
+It is unclear from the above summary tables whether either variable has a clear effect.
 

### 2021-01-12 analysis/analysis.Rmd
fixed some typos in sup mat

@@
 
-We will remove identity relationships (i.e., a group is a subgroup of itself) and the three pairs of wallpapers groups that it subgroups of each other (e.g., PM is a subgroup of CM, and CM is a subgroup of PM).
+We will remove identity relationships (i.e., a group is a subgroup of itself) and the three pairs of wallpapers groups that are subgroups of each other (e.g., PM is a subgroup of CM, and CM is a subgroup of PM).
 This leaves us with a total of `r nrow(subgroups)` subgroups to include in our analysis.
@@
   + maximal random effect structure.
-* Check model fit (r̂ statistics and trace plots).
+* Check model fit ($r^2$ statistics and trace plots).
 * Extract mcmc samples from the posterior distributions.
@@
 Our independent variable is a categorical factor with 16 levels.
-We will drop the intercept from our model and instead fit a coefficent for each factor level ($y \sim x - 0$).
-As our dependant variable will be log-transformed, we can use the priors below:
+We will drop the intercept from our model and instead fit a coefficient for each factor level ($y \sim x - 0$).
+As our defendant variable will be log-transformed, we can use the priors below:
 
@@
 
-We can see that i) our priors are relatively weak as the predictions span several orders of magnitide, and ii) our empirical data falls within this range.
+We can see that i) our priors are relatively weak as the predictions span several orders of magnitude, and ii) our empirical data falls within this range.
 
@@
 comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
-  # compute probabilties
+  # compute probabilities
   subgroup_comp_dat %>%
@@
 
-# compute probabilties
+# compute probabilities
 comp_summary$eeg <- comparison_summary(subgroup_comp$eeg, TRUE, 'rms')
@@
 
-# compute probabilties
+# compute probabilities
 comp_summary$threshold <- comparison_summary(subgroup_comp$threshold, FALSE, 'threshold')
@@
 
-# compute probabilties
+# compute probabilities
 comp_summary$occ_even <- comparison_summary(subgroup_comp$occ_even, TRUE, 'rms')
@@
 
-# compute probabilties
+# compute probabilities
 comp_summary$par_odd <- comparison_summary(subgroup_comp$par_odd, TRUE, 'rms')
@@
 
-If we take $p$=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/63 = 89% and 49/64 = 78% of the comparisons for the primary SSVEP and display durations repectively.
-This compares to the 32/64= 50% and 22/64 = 35% for the control SSVEP conditions.
+If we take $p$=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/63 = 89% and 49/63 = 78% of the comparisons for the primary SSVEP and display durations respectively.
+This compares to the 32/63= 51% and 22/63 = 35% for the control SSVEP conditions.
 
@@
 
-We will also investigate if we see the corresponding pattern with the display duration threshold data, with the time taken to detect the symmetry dreasing as we increase the amount of rotational symmetry.
+We will also investigate if we see the corresponding pattern with the display duration threshold data, with the time taken to detect the symmetry decreasing as we increase the amount of rotational symmetry.
 

### 2021-01-20 analysis/analysis.Rmd
removing normality

@@
 
-## Index and Normality
+## Index
 
-Subgroup relations can be classified by their index, and by whether they are normal or not.
-Here we investigate the extent to which these two variables can account for the variation between the subgroup relationships.
+Subgroup relations can be classified by their index. Here we investigate the extent to which index can account for the variation between the subgroup relationships.
 
@@
    mutate(
-     index = as.numeric(as.character(index)),
-     normal = if_else(normal, 1, 0)) -> comp_summary$eeg
+     index = as.numeric(as.character(index))) -> comp_summary$eeg
 
-my_models$eeg <- brm(data = comp_summary$eeg, mean_value ~ index * normal)
+my_models$eeg <- brm(
+  data = comp_summary$eeg, 
+  scale(mean_value) ~ index)
 
@@
    mutate(
-     index = as.numeric(as.character(index)),
-     normal = if_else(normal, 1, 0)) -> comp_summary$threshold
+     index = as.numeric(as.character(index))) -> comp_summary$threshold
 
-my_models$disp <- brm(data = comp_summary$threshold, mean_value ~ index * normal)
+my_models$disp <- brm(
+  data = comp_summary$threshold, 
+  scale(mean_value) ~ index)
 
@@
 ```
-
-It is unclear from the above summary tables whether either variable has a clear effect.
 

### 2021-01-21 analysis/scripts/import_subgroup_comparisons.R
fix arrows

@@ -50,7 +50,7 @@ subgroups %>%
     left_join(subgroups) %>%
     mutate(
         colour_code = if_else(paste(group, subgroup, sep="-") %in% to_remove, 2, 1), 
-       index = as.character(index),
+        index = as.character(index),
         index = if_else(is.na(index), "-", index),
         index = kableExtra::cell_spec(
              index, 
@@ -69,4 +69,4 @@ subgroups %>% filter(
     is.finite(index),
     group != subgroup,
     !(paste(group, subgroup, sep="-") %in% to_remove)) %>%
-    mutate(label = paste(group, "≤", subgroup)) -> subgroups
+    mutate(label = paste(group, "->", subgroup)) -> subgroups

### 2021-01-21 analysis/analysis.Rmd
unscaled variables

@@
   data = comp_summary$eeg, 
-  scale(mean_value) ~ index)
+  mean_value ~ index)
 
@@
   data = comp_summary$threshold, 
-  scale(mean_value) ~ index)
+  mean_value ~ index)
 
@@
 
-```{r, fig.cap = "The effect of index and normality on log(rms) and log(ms)."}
+```{r, fig.cap = "The effect of index on log(rms) and log(ms)."}
 comp_summary$eeg %>%
@@
  bind_rows(d_plt_eeg, d_plt_disp) %>%
-   ggplot(aes(x =  index, y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
+   ggplot(aes(x =  index, y = .value, group = interaction(.draw, normal))) +
    geom_line(alpha = .15) +
@@
      data = d_summary, 
-     aes(x =  index, y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
+     aes(x =  index, y = mean_value, group = 1), width = 0.1, height = 0, alpha = 0.33) +
    theme_minimal() + 

### 2021-01-21 Writeup/manuscript.tex

@@
 
-\begin{abstract}
-A century of vision research has demonstrated that symmetry contributes to numerous domains of visual perception \cite{RN1311, RN1166,RN1682,RN1824}.
-In a 2-D image, the four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in 17 distinct ways.
-These 17 “wallpaper” groups \cite{RN1562,RN1563,RN1425} obey a hierarchy, determined by mathematical group theory, in which simpler groups are subgroups of more complex ones\cite{RN1711}.
-Here we probe representations of symmetries in wallpaper groups using two methods: (1) Steady-State Visual Evoked Potentials (SSVEPs) recorded using EEG and (2) symmetry detection thresholds measured psychophysically.
-We find that hierarchical relationships between the wallpaper groups are almost perfectly preserved in both behavior and response amplitudes in visual cortex.
-This remarkable consistency between the structure of symmetry representations and mathematical group theory, is likely generated over visual development, through implicit learning of regularities in the environment.
+\begin{abstract}Symmetries are present at many scales in images of natural scenes.
+A large body of literature has demonstrated contributions of symmetry to numerous domains of visual perception.
+The four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in exactly 17 distinct ways.
+These \textit{wallpaper groups} represent the complete set of symmetries in 2D images.
+The goal of the current study is to provide a complete description of responses to symmetry in the human visual system, by collecting both brain and behavioral data using all 17 groups.
+This allows us to probe the hierarchy of complexity among wallpaper groups, in which simpler groups are subgroups of more complex ones.
+We find that this hierarchy is preserved almost perfectly in both behavior and brain activity: A multi-level Bayesian GLM indicates that for most of the 63 subgroup relationships, subgroups produce lower amplitude responses in visual cortex (posterior probability: >0.95 for 56 of 63) and require longer presentation durations to be reliably detected (posterior probability: >0.95 for 49 of 63).
+This systematic pattern is seen only in visual cortex and only in components of the brain response known to be symmetric-specific.
+Our results show that representations of symmetries in the human brain are precise and rich in detail, and that this precision is reflected in behavior.
+These findings expand our understanding of symmetry perception, and open up new avenues for research on how fine-grained representations of regular textures contribute to natural vision.
 \end{abstract}
@@
 The importance of symmetry for visual perception has been known at least since the gestalt movement of the early 20th century.
-Since then, symmetry has been shown to contribute to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection \cite{RN1337}.
+Since then, symmetry has been shown to contribute to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection.
 Most of this work has focused on mirror symmetry or \textit{reflection}, with much less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection}.

### 2021-01-21 analysis/analysis.Rmd
minor sup mat edits

@@
 
-Here we import our data and make some summary plots.
+First we import our data and make some summary plots.
 
@@
 
-The table below shows how many groups per participant we have data for.
-
-```{r}
-d_dispthresh %>% 
-  group_by(subject) %>% 
-  summarise(count = n(), .groups = "drop") %>% 
-  knitr::kable() %>% 
-  kableExtra::kable_styling(full_width = F)
-```
-
 ## Control Data
 
-In addition to the primary SSVEP data set, we are also importing two control data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data from six parietal electrodes (see Figure 1.1 and the main paper).
+In addition to the primary SSVEP data set, we are also importing two control EEG data sets which are (a) even harmonic data from the same occipital electrodes, and (b) odd harmonic data from six parietal electrodes (see Figure 1.1 and the main paper).
 
@@
 
-Information on the symmetries and subgroups contained within each wallpaper group was obtained from <cite> and is summarised in the files `symmeries_in_group.csv`, `subgroup_relations.csv`, and `subgroup_normal.csv`.
+Information on the symmetries and subgroups contained within each wallpaper group was obtained from Coxeter & Moses (1980, Generators and Relations
+for Discrete Groups) and is summarised in the files `symmeries_in_group.csv`, `subgroup_relations.csv`, and `subgroup_normal.csv`.
 
@@
 
-knitr::kable(sym_dat, caption = "wallpaper group summary") %>% 
+knitr::kable(sym_dat, caption = "Wallpaper group summary") %>% 
     kableExtra::kable_styling(full_width = F)
@@
   + maximal random effect structure.
-* Check model fit ($r^2$ statistics and trace plots).
 * Extract mcmc samples from the posterior distributions.
@@
 We will drop the intercept from our model and instead fit a coefficient for each factor level ($y \sim x - 0$).
-As our defendant variable will be log-transformed, we can use the priors below:
+As our dependent variable will be log-transformed, we can use the priors below:
 

### 2021-01-27 Writeup/manuscript.tex
final version of manuscript

@@
 In recent years wallpaper groups have found use in the vision science community, as an ideal stimulus set for studying the perception of symmetries in textures.
-Here we present brain imaging and psychophysical data on the complete set of wallpaper groups and show the hierarchical organization among wallpaper groups in reflected in both representations in visual cortex and performance on a symmetry detection task.
-This shows that the visual system is highly sensitive to reguarities in textures, and suggest that symmetries may play an important role in texture perception.}
+Here we present brain imaging and psychophysical data using the complete set of wallpaper groups and show that the hierarchical organization among wallpaper groups is reflected both in representations in visual cortex and performance on a symmetry detection task.
+This shows that the visual system is highly sensitive to regularities in textures, and suggest that symmetries may play an important role in texture perception.}
 
@@
 The importance of symmetry for visual perception has been known at least since the gestalt movement of the early 20th century.
-Since then, symmetry has been shown to contribute to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection.
+Since then, symmetry has been shown to contribute to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection \cite{RN1337}.
 Most of this work has focused on mirror symmetry or \textit{reflection}, with much less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection}.
-In the two spatial dimensions relevant for images, these four fundamental symmetries can be combined in 17 distinct ways, the “wallpaper” groups \cite{RN1562,RN1563,RN1425}.
-Previous work on a subset of four wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}.
-This effect is also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\cite{RN1725} or event-related paradigms \cite{RN1959}.
+In the two spatial dimensions relevant for images, these four fundamental symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \cite{RN1562,RN1563,RN1425}.
+Previous work on a subset of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}.
+This effect was also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\cite{RN1725} or event-related paradigms \cite{RN1959}.
 Here we extend this work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups.
@@
 In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \cite{RN1425}.
+The full list of subgroup relationships is listed in Section 1.4.2 of the Supplementary Material.
 Subgroup relationships between wallpaper groups can be distinguished by their indices.
 The index of a subgroup relationship is the number of cosets, i.e.
-the number of times the subgroup is found in the outer group \cite{RN1425}.
-As an example, let us consider groups P6 and P2.
-If we ignore the translations in two directions that both groups share, group P6 consists of the set of rotations {$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$}, in which P2 {$0^{\circ}$, $180^{\circ}$} is contained.
-P2 is thus a subgroup of P6, and the full P6 set can be generated by every combination of P2 and rotations {$0^{\circ}$, $120^{\circ}$, $240^{\circ}$}.
-Because P2 is repeated three times in P6, P2 is a subgroup of P6 with index 3 \cite{RN1425}.
-The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are sub-groups of more complex ones \cite{RN1711}.
-The full set of subgroup relationships is listed in Section 1.4.2 of the Supplementary Material.
+the number of times the subgroup is found in the supergroup \cite{RN1425}.
+As an example, let us consider groups \textit{P2} and \textit{P6}.
+If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained.
+\textit{P2} is thus a subgroup of \textit{P6}, and \textit{P6} can be generated by combining \textit{P2} with rotations \{$0^{\circ}$, $120^{\circ}$, $240^{\circ}$\}.
+Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \cite{RN1425}.
+The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \cite{RN1711}.
 
@@
 The SSVEP data are odd-harmonic-filtered cycle-average waveforms.
-In each cycle, a P1 exemplar was shown for the first 600 ms, followed by the original exemplar for the last 600 ms.
+In each cycle, a \textit{P1} exemplar was shown for the first 600 ms, followed by the original exemplar for the last 600 ms.
 Errorbars are standard error of the mean.
@@
 
-The two datasets presented here puts on in the position of being able to assess the extent to which both behavior and brain responses follow that subgroup hierachy.
-Based on previous imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \cite{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
-For the psychophysical data, we hypothesized that more complex groups would lead to shorter symmetry detection thresholds, based on previous data showing that under a fixation presentation time, discriminability increases with the number of symmetry axes in the pattern \cite{wagemans_1991}.
+The two datasets we present here make it possible to assess the extent to which both behavior and brain responses follow the hierachy of complexity expressed by the subgroup relationships.
+Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \cite{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
+For the psychophysical data, we hypothesized that more complex groups would lead to shorter symmetry detection thresholds, based on previous data showing that under a fixed presentation time, discriminability increases with the number of symmetry axes in the pattern \cite{wagemans_1991}.
 Our results confirm both hypotheses, and show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds following these relationships at a level that is far beyond chance.
-Visual cortex thus appears to encode all of the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.
-Given that most participants had no knowledge of group theory, the ordered structure of visual responses to wallpaper groups is likely learned implicitly from regularities in the visual environment.
+The human visual system thus appears to encode all of the fundamental symmetries using a representational structure that closely approximates the subgroup relationships from group theory.
 
@@
 
-The stimuli used in our two experiments were multiple exemplar images belonging to each of the wallpaper groups, generated from random-noise textures, as described in detail elsewhere \cite{RN1725}.
-Exemplar images from group P1 was used as control stimuli, and each exemplar from the other 16 groups had a power-spectrum matched P1 exemplar.
-The matched P1 exemplars were generated by phase-scrambling the exemplar images.
-Because all wallpapers are periodic due to their lattice tiling structure, the phase-scrambled images all belong to group P1 regardless of group membership of the original exemplar.
-P1 contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \cite{RN1425}.
+The stimuli used in our two experiments were generated from random-noise textures, which made it possible to generate multiple exemplars from each of the wallpaper groups, as described in detail elsewhere \cite{RN1725}.
+We generated control stimuli matched to each exemplar in the main stimulus set, by scrambling the phase but maintaining the power spectrum.
+All wallpaper groups are inherently periodic because of their repeating lattice structure.
+Phase scrambling maintains this periodicity so the phase-scrambled control images all belong to group \textit{P1} regardless of group membership of the original exemplar.
+\textit{P1} contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \cite{RN1425}.
 In our SSVEP experiment, this stimulus set allowed us to isolate brain activity specific to the symmetry structure in the exemplar images from activity associated with modulation of low-level features, by alternating exemplar images and control exemplars.
@@
 In each trial an exemplar image was shown with its matched control, one image after the other, and the order varied pseudo-randomly such that in half the trials the original exemplar was shown first, and in the other half the control image was shown first.
-After each trial, participants were told to indicate whether the first or second image contained more structure, and the duration of both images was controlled by a staircase procedure so that a display duration threshold could be computed for each wallpaper group.
+After each trial, participants were told to indicate whether the first or second image contained more structure, and the duration of both images was controlled by a staircase procedure so that a threshold duration for symmetry detection could be computed for each wallpaper group.
 
-A summary of our brain imaging and psychophysical measurements is presented with examples of the wallpaper groups in Figure 1.
+Examples of the wallpaper groups and a summary of our brain imaging and psychophysical measurements is in Figure 1.
 For our primary SSVEP analysis, we only considered EEG data from a pre-determined region-of-interest (ROI) consisting of six electrodes over occipital cortex (see Supplementary Figure 1.1).
 SSVEP data from this ROI was filtered so that only the odd harmonics that capture the symmetry response contribute to the waveforms.
-While waveform amplitude is quite variable among the 16 groups, all groups have a sustained negative-going response that begins at about the same time for all groups, ~180 ms after the transition from the P1 control exemplar to the original exemplar.
-To reduced the amplitude of the symmetry-specific response to a single number that could be used in further analyses and compared to the psychophysical data, we computed the root-mean-square (RMS) over the odd-harmonic-filtered waveforms.
+While waveform amplitude is quite variable among the 16 groups, all groups have a sustained negative-going response that begins at about the same time for all groups, ~180 ms after the transition from the \textit{P1} control exemplar to the original exemplar.
+To reduce the amplitude of the symmetry-specific response to a single number that could be used in further analyses and compared to the psychophysical data, we computed the root-mean-square (RMS) over the odd-harmonic-filtered waveforms.
 The data in Figure 1 are shown in descending order according to RMS.
-The psychophysical results, shown in box plots in Figure 1, were also quite variable between groups, and there seems to be a general pattern where wallpaper groups near the top of the figure, that have lower SSVEP amplitudes, also have longer psychophysical display duration thresholds.
+The psychophysical results, shown in box plots in Figure 1, were also quite variable between groups, and there seems to be a general pattern where wallpaper groups near the top of the figure, that have lower SSVEP amplitudes, also have longer psychophysical threshold durations.
 
-We now wanted to test our two assumptions about how SSVEP amplitudes and duration thresholds would follow subgroup relationships, and thereby quantify the degree to which our two measurements were consistent with group theory.
-We tested each of the two hypotheses using the same approach.
+We now wanted to test our two hypotheses about how SSVEP amplitudes and threshold durations would follow subgroup relationships, and thereby quantify the degree to which our two measurements were consistent with the group theoretical hierarchy of complexity.
+We tested each hypothesis using the same approach.
 We first fitted a Bayesian model with wallpaper group as a factor and participant as a random effect.
 We fit the model separately for SSVEP RMS and psychophysical data, and then computed posterior distributions for the difference between supergroup and subgroup.
-These difference distributions could allowed us to compute the conditional probability that the supergroup would produce (a) larger RMS and (b) a shorter threshold durations, when compared to the subgroup.
+These difference distributions allowed us to compute the conditional probability that the supergroup would produce (a) larger RMS and (b) a shorter threshold durations, when compared to the subgroup.
 The posterior distributions are shown in Figure 2 for the SSVEP data, and in Figure 3 for the psychophysical data, which distributions color-coded according to conditional probability.
-For both data sets our hypothesis is confirmed: For the overwhelming majority of the 64 subgroup relationships, supergroups are more likely to produce larger symmetry specific SSVEPs and shorter threshold durations, and in most cases the conditional probability of this happening is extremely high.
+For both data sets our hypothesis is confirmed: For the overwhelming majority of the 63 subgroup relationships, supergroups are more likely to produce larger symmetry-specific SSVEPs and shorter symmetry detection threshold durations, and in most cases the conditional probability of this happening is extremely high.
 
@@
 By comparing these two control analysis to our primary SSVEP analysis, we can address the specify of our effects in terms of location (occipital cortex vs parietal cortex) and harmonic (odd vs even).
-For both control analyses (plotted in Supplementary Figures 3.3 and 3.4), the correspondence between data and subgroup relationships was weaker than in the primary analysis.
+For both control analyses (plotted in Supplementary Figures 3.3 and 3.4), the correspondence between data and subgroup relationships was substantially weaker than in the primary analysis.
 We can quantify the strength of the assocation between the data and the subgroup relationships, by asking what proportion of subgroup relationships that reach or exceed a range of probability thresholds.
 This is plotted in Figure 4, for our psychophysical data, our primary SSVEP analysis and our two control SSVEP analyses.
-It it that odd-harmonic SSVEP data from an occipital ROI and display duration thresholds both have a strong association with the subgroup relations, that for a clear majority of the subgroups survive even at the highest threshold we consider ($p(\Delta>0|data)>0.99$), and that the assocation is far weaker for the two control analyses.
+It shows that odd-harmonic SSVEP data from the occipital ROI and symmetry detection threshold durations both have a strong association with the subgroup relationships such that a clear majority of the subgroups survive even at the highest threshold we consider ($p(\Delta>0|data)>0.99$).
+The assocation is far weaker for the two control analyses.
 
-SSVEP data from four of the wallpaper groups (P2, P3, P4 and P6) was previously published as part of our earlier demonstration of parametric responses to rotation symmetry in wallpaper groups\cite{RN1725}.
-We replicate that result using our Bayesian approach, and find the same parametric effect in the psychophysical data (Supplementary Figure 4.1).
-We also conducted an analysis \textcolor{red}{DOES THIS NEED TO BE SPELLED OUT MORE} looking for effects of index and normality in our two datasets, and found that subgroup relationships with higher indices tended to produce stronger effects for both SSVEP RMS and symmetry detection thresholds, as shown by the fact that a leave-one-out test \textcolor{red}{...
-MORE DETAILS NEEDED HERE}.
-The effect of index did not (\textcolor{red}{???} reach significance.
+SSVEP data from four of the wallpaper groups (\textit{P2}, \textit{P3}, \textit{P4} and \textit{P6}) was previously published as part of our earlier demonstration of parametric responses to rotation symmetry in wallpaper groups\cite{RN1725}.
+We replicate that result using our Bayesian approach, and find an analogous parametric effect in the psychophysical data (see Supplementary Figure 4.1).
+We also conducted an analysis testing for an effect of index in our two datasets, and found that subgroup relationships with higher indices tended to produce greater pairwise differences between the subgroup and supergroup, for both SSVEP RMS and symmetry detection thresholds (see Supplementary Figure 4.2).
+The effect of index is relatively weak, but the fact that there is a measurable index effect can nonetheless be taken as preliminary evidence that representations of symmetries in wallpaper groups may be compositional.
 
-Finally, we conducted a correlation analysis comparing SSVEP and psychophysical data, and found a small ($R^2 = 0.44$) but above-zero correlation, as indicated by our confidence intervals.
-There are several factors that might explain the relatively weak correlation, most prominently the fact that the same individuals did not participate in each of the two experiments.
-Nevertheless, we find the relationship between the two datasets interesting, because it suggests that our psychophysical and SSVEP measurements are tapping into the same underlying mechanisms.
+Finally, we conducted a correlation analysis comparing SSVEP and psychophysical data, and found a modest ($R^2 = 0.44$) but above-zero correlation.
+The correlation between the two datasets is interesting because it suggests that our psychophysical and SSVEP measurements are at least to some extent tapping into the same underlying mechanisms.
 
@@
 \includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.png}
-\caption{Posterior distributions for the difference in mean RMS SSVEP response.
+\caption{Posterior distributions for the difference in mean SSVEP RMS amplitude.
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that xx/64 subgroup relationships have $p(\Delta|data)>0.9$.}
+We can see that 56/63 subgroup relationships have $p(\Delta|data)>0.95$.}
 \label{fig:eeg_comp}
@@
 \includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_psychophysical.png}
-\caption{Posterior distributions for the difference in mean display duration threshold.
+\caption{Posterior distributions for the difference in mean symmetry detection threshold durations.
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that xx/64 subgroup relationships have $p(\Delta|data)>0.9$.}
+We can see that 49/63 subgroup relationships have $p(\Delta|data)>0.95$.}
 \label{fig:durations_comp}
@@
 
-Here we show that beyond merely responding to the elementary symmetry operations of reflection \cite{RN1170} and rotation \cite{RN1725}, the visual system explicitly represents hierarchical structure of the 17 wallpaper groups, and thus the compositions of all four of the fundamental symmetry transformations (rotation, reflection, translation, glide reflection) which comprise regular textures.
-The SSVEP amplitude and the symmetry detection threshold both preserve the complex hierarchy of subgroup relationships among the wallpaper groups \cite{RN1711}.
-For the SSVEP, this remarkable consistency was specific to the odd harmonics of the stimulus frequency, that capture the symmetry-specific response \cite{RN1725} and to electrodes in an ROI over occipital cortex.
-When the same analysis was done using the odd harmonics from electrodes over parietal cortex (Supplementary Figure 3.3) or even harmonics from electrodes over occipital cortex (Supplementary Figure 3.4), the data was much less consistent with the subgroup relationships (yellow and green lines, Figure 4).
+Here we show that beyond merely responding to the elementary symmetry operations of reflection \cite{RN1170} and rotation \cite{RN1725}, the visual system explicitly represents hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
+Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \cite{RN1711}.
+For the SSVEP, this remarkable consistency was specific to the odd harmonics of the stimulus frequency that are known to capture the symmetry-specific response \cite{RN1725} and to electrodes in a region-of-interest (ROI) over occipital cortex.
+When the same analysis was done using the odd harmonics from electrodes over parietal cortex (Supplementary Figure 3.3) or even harmonics from electrodes over occipital cortex (Supplementary Figure 3.4), the data was substantially less consistent with the subgroup relationships (yellow and green lines, Figure 4).
 
 The current data provide a complete description of the visual system’s response to symmetries in the 2-D plane.
-Our design precludes us from independently measure the response to P1, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where P1 is the subgroup, are also preserved by the visual system.
+Our design precludes us from independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
 The subgroup relationships are in many cases not obvious perceptually, and most participants had no knowledge of group theory.
 Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships.
-Furthermore, previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, \textcolor{green}{when asked to sort exampler images into visually similar subsets, they tend to make between four and 12 sets.
-It is common for examplars from two distinct symmetry groups to be judged as belonging to the same subset, and also for examplars from the same group to be placed into disjoint similarity subsets.}
+Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, they tend to sort them into fewer groups than there actually are (4-12 groups) and it is common for exemplars from different wallpaper groups to be sorted in the same group \cite{RN172}.
+The more controlled two-interval forced choice approach used in the current behavioral experiment allows us to show that more granular representations of wallpaper groups are measurable in behavior.
+The relatively modest correlation we observe between behavior and brain imaging data may suggest that the relationship between the two measurements is more complex than what can be captured in our summary statistics.
+Future work in which the same individuals participate in both behavioral and brain imaging experiments will help tease apart these complexities and enhance our understanding of the brain mechanisms that drive both measurements.
+
+We also find an of index for both our brain imaging measurements and our symmetry detection thresholds.
+This means that the visual system not only represents the hierarchical relationship captured by individual subgroups, but also distinguishes between subgroups depending on how many times the subgroup is repeated in the supergroup, with more repetitions leading to larger pairwise differences.
+Our measured effect of index is relatively weak.
+This is perhaps because the index analysis does not take into account the \textit{type} of isometries that differentiate the subgroup and supergroup.
+The effect of symmetry type can be observed by contrasting the measured SSVEP amplitudes and detection thresholds for groups \textit{PM} and \textit{PG} in Figure 1.
+The two groups are comparable except \textit{PM} contains reflection and \textit{PG} contains glide reflection, and the former clearly elicits higher amplitudes and lower thresholds.
+The charting of how different symmetry types contribute to the representational hierarchy will be an important goal for future work.
 
@@
 The ordered structure of responses to wallpaper groups could be driven by a central tenet of neural coding, that of efficiency.
-If coding is to be efficient, neural resources should be distributed in such a way that the structure of the environment is captured with minimum redundancy considering the visual geometric optics, the capabilities of the subsequent neural coding stages and the behavioral goals of the organism \cite{RN1758, RN1760, RN1757, RN1756}.
+If coding is to be efficient, neural resources should be distributed to capture the structure of the environment with minimum redundancy considering the visual geometric optics, the capabilities of the subsequent neural coding stages and the behavioral goals of the organism \cite{RN1758, RN1760, RN1757, RN1756}.
 Early work within the efficient coding framework suggested that natural images had a $1/f$ spectrum and that the corresponding redundancy between pixels in natural images could be coded efficiently with a sparse set of oriented filter responses, such as those present in the early visual pathway \cite{RN1740, RN1446}.
@@
 The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective and occlusion.
-Near-regular textures, however abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \cite{RN1519}.
+Near-regular textures, however, abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \cite{RN1519}.
 The correspondence between visual cortex responses and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space.
@@
 Field biologists have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \cite{RN1775}.
-In the norm-based encoding framework, wallpaper groups can be considered super-textures, exaggerated examples of the near-regular textures that surround us.
+In the norm-based encoding framework, wallpaper groups can be considered \textit{supertextures}, exaggerated examples of the near-regular textures that surround us.
 Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \cite{RN1764}.
 Wallpaper groups are visually compelling and have been widely used in human artistic expression going back to the Neolithic age \cite{RN1949}.
-If wallpapers are in fact super-textures, it would imply that this prevalence may be a direct consequence of the strategy the human visual system uses for encoding textures.
+If wallpapers are in fact supertextures, this prevalence may be a direct result of the strategy the human visual system has adopted for texture encoding.
 
@@
 \includegraphics[width=1\linewidth]{../analysis/plots/model_roc_style.pdf}
-\caption{This plot shows the proportion of subgroup relations that satisfy $p(\Delta >0 | data) > x$.
-We can see that if we take $x = 0.95$ as our threshold, the subgroup relations are preserved in 56/63 = 89\% and 49/64 = 78\% of the comparisons for the primary EEG and display durations repectively.
-This compares to the 32/64= 50\% and 22/64 = 35\% for the control EEG conditions.}
+\caption{This plot shows the proportion of subgroup relationships that satisfy $p(\Delta >0 | data) > x$.
+We can see that if we take $x = 0.95$ as our threshold, the subgroup relationships are preserved in 56/63 = 89\% and 49/64 = 78\% of the comparisons for the primary SSVEP and threshold duration datasets, repectively.
+This compares to the 32/64= 50\% and 22/64 = 35\% for the SSVEP control datasets.}
 \label{fig:durations_rotations}
@@
 The use of noise textures as the starting point for stimulus generation allowed the creation of an almost infinite number of distinct exemplars of each wallpaper group.
-For each exemplar image, phase-randomized control exemplars were generated that had the same power spectrum as the exemplar images for each group.
+To make individual exemplars as similar as possible we replaced the power spectrum of each exemplar with the median across exemplars within a group.
+We then generated control exemplars that had the same power spectrum as the exemplar images by randomizing the phase of each exemplar image.
 The phase scrambling eliminates rotation, reflection and glide-reflection symmetries within each exemplar, but the phase-scrambled images inherent the spectral periodicity arising from the periodic tiling.
-This means that all control exemplars, regardless of which wallpaper group they are derived from, degenerate to another symmetry group, namely P1.
-P1 is the simplest of the wallpaper groups, and contains only translations of a region whose shape derives from the lattice.
-Because the different wallpaper groups have different lattices, P1 controls matched to different groups have different power spectra.
+This means that all control exemplars, regardless of which wallpaper group they are derived from, are transformed into another symmetry group, namely \textit{P1}.
+\textit{P1} is the simplest of the wallpaper groups, and contains only translations of a region whose shape derives from the lattice.
+Because the different wallpaper groups have different lattices, \textit{P1} controls matched to different groups have different power spectra.
 Our experimental design takes these differences into account by comparing the neural responses evoked by each wallpaper group to responses evoked by the matched control exemplars.
@@
 Stimulus presentation was controlled using in-house software.
-
 For the psychophysics experiment, the stimuli were shown on a $48 \times 27$cm VIEWPixx/3D LCD Display monitor, model VPX-VPX-2005C, resolution $1920 \times 1080$ pixels, with a viewing distance of approximately 40cm and linear gamma.
@@
 \subsection*{EEG Procedure}
-Visual Evoked Potentials were measured using a steady-state design, in which P1 control images alternated with test images from each of the 16 other wallpaper groups[2].
-Exemplar images were always preceded by their matched P1 control image.
-A single 0.83 Hz stimulus cycle consisted of a control P1 image followed by an exemplar image, each shown for 600 ms.
+Visual Evoked Potentials were measured using a steady-state design, in which \textit{P1} control images alternated with exemplar images from each of the 16 other wallpaper groups.
+Exemplar images were always preceded by their matched \textit{P1} control image.
+A single 0.83 Hz stimulus cycle consisted of a control \textit{P1} image followed by an exemplar image, each shown for 600 ms.
 A trial consisted of 10 such cycles (12 sec) over which 10 different exemplar images and matched controls from the same rotation group were presented.
@@
 Two times per trial, an image pair was shown at reduced contrast, and the participants were instructed to press a button on a response pad.
-We adjusted the contrast reduction such that average accuracy for each participant was kept at ~85\% correct, so that the difficulty of the vigilance task was kept constant.
+We adjusted the contrast reduction such that average accuracy for each participant was kept at ~85\% correct, in order to keep the difficulty of the vigilance at a constant level.
 
 \subsection*{Psychophysics Procedure}
-The experiment consisted of 16 blocks, one for each of the wallpaper groups (excluding $P1$).
-In each trial, participants were presented with two stimuli (one of which was the wallpaper group for the current block of trials, the other being $P1$), one after the other (inter stimuli interval of 700ms).
+The experiment consisted of 16 blocks, one for each of the wallpaper groups (excluding \textit{P1}).
+We used a two-interval forced choice approach.
+In each trial, participants were presented with two stimuli (one of which was the wallpaper group for the current block of trials, the other being \textit{P1}), one after the other (inter-stimulus interval of 700ms).
 After each stimuli had been presented, it was masked with white noise for 300ms.
-After both stimuli had been presented, participants made a response on the keyboard to indicate whether they thought the first or second contained the most symmetry.
-Each block started with 10 practise trials, (stimulus display duration of 500ms) to allow participants to familiarise themselves with the current block's wallpaper pattern.
+After both stimuli had been presented, participants made a response on the keyboard to indicate whether they thought the first or second image contained more symmetry.
+Each block started with 10 practice trials, (stimulus display duration of 500ms) to allow participants to familiarise themselves with the current block's wallpaper pattern.
 If they achieved an accuracy of 9/10 in these trials they progressed to the rest of the block, otherwise they carried out another set of 10 practise trials.
@@
 \subsection*{EEG Acquisition and Preprocessing}
-Electroencephalogram Acquisition and Preprocessing.
-The time-locked Steady-State Visual Evoked Potentials were collected with 128-sensor HydroCell Sensor Nets (Electrical Geodesics, Eugene, OR) and were band-pass filtered from 0.3 to 50 Hz.
+Steady-State Visual Evoked Potentials (SSVEPs) were collected with 128-sensor HydroCell Sensor Nets (Electrical Geodesics, Eugene, OR) and were band-pass filtered from 0.3 to 50 Hz.
 Raw data were evaluated off line according to a sample-by-sample thresholding procedure to remove noisy sensors that were replaced by the average of the six nearest spatial neighbors.
@@
 This was typically the case for epochs containing artifacts, such as blinks or eye movements.
-The use of steady-state stimulation drives cortical responses at specific frequencies directly tied to the stimulus frequency.
+Steady-state stimulation will drive cortical responses at specific frequencies directly tied to the stimulus frequency.
 It is thus appropriate to quantify these responses in terms of both phase and amplitude.
 Therefore, a Fourier analysis was applied on every remaining epoch using a discrete Fourier transform with a rectangular window.
-The use of epochs two-cycles (i.e., 2.4 s) long, was motivated by the need to have a relatively high resolution in the frequency domain, $\delta$f = 0.42 Hz.
+The use of two-cycle long epochs (i.e., 2.4 s) was motivated by the need to have a relatively high resolution in the frequency domain, $\delta$f = 0.42 Hz.
 For each frequency bin, the complex-valued Fourier coefficients were then averaged across all epochs within each trial.
@@
 For each participant, the average Fourier coefficients from the two sessions were averaged over trials and sessions.
-The Steady-State Visual Evoked Potentials (SSVEP) paradigm we used allowed us to separate symmetry-related responses from non-specific contrast transient responses.
-Previous work has demonstrated that symmetry-related responses are predominantly found in the odd harmonics of the stimulus frequency, whereas the even harmonics consist mainly of responses unrelated to symmetry, that arise from the contrast change associated with the appearance of the second image[2-4].
+The SSVEP paradigm we used allowed us to separate symmetry-related responses from non-specific contrast transient responses.
+Previous work has demonstrated that symmetry-related responses are predominantly found in the odd harmonics of the stimulus frequency, whereas the even harmonics consist mainly of responses unrelated to symmetry, that arise from the contrast change associated with the appearance of the second image \cite{RN1954, RN1725}.
 This functional distinction of the harmonics allowed us to generate a single-cycle waveform containing the response specific to symmetry, by filtering out the even harmonics in the spectral domain, and then back-transforming the remaining signal, consisting only of odd harmonics, into the time-domain.
 For our main analysis, we averaged the odd harmonic single-cycle waveforms within a six-electrode region of interest (ROI) over occipital cortex (electrodes 70, 74, 75, 81, 82, 83).
-These waveforms, averaged over participants, are shown in Figure 2 in the main paper.
-The same analysis was done for the even harmonics (see Figure S1) and for the odd harmonics within a six electrode ROI over parietal cortex (electrodes 53, 54, 61, 78, 79, 86; see Figure S2).
+These waveforms, averaged over participants, are shown in Figure 1.
+The same analysis was done for the even harmonics and for the odd harmonics within a six electrode ROI over parietal cortex (electrodes 53, 54, 61, 78, 79, 86; see Supplementary Figure 1.1).
 The root-mean square values of these waveforms, for each individual participant, were used to determine whether each of the wallpaper subgroup relationships were preserved in the brain data.
 
-\subsection*{Bayesian Analysis of EEG and Psychophysical data}
+\subsection*{Defining the list of subgroup relationships}
+In order to get the complete list of subgroup relationships, we digitized Table 4 from Coxeter \cite{RN1711} (shown in Supplementary Table 1.2).
+After removing identity relationships (i.e.
+each group is a subgroup of itself) and the three pairs of wallpapers groups that are subgroups of each other (e.g.
+\textit{PM} is a subgroup of \textit{CM}, and \textit{CM} is a subgroup of \textit{PM}) we were left with a total of 63 unambiguous subgroups that were included in our analysis.
+
+\subsection*{Bayesian Analysis of SSVEP and Psychophysical data}
 Bayesian analysis was carried out using R (v3.6.1) \cite{R} with the \texttt{brms} package (v2.9.0) \cite{burkner2017} and rStan (v2.19.2 \cite{rStan}).
 The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16 level factor, and random effects for participant.
-The EEG data and display thresholds were modelled using log-normal distributions with weakly informative, $ \mathcal{N}(0, 2)$, priors.
-After fitting the model to the data, samples were drawn from the posterior distribution for each mean of the EEG response (display duration) for each wallpaper group.
-These samples were then recombined to calculate the distribution of differences for each pair of subgroup and super-group.
+The SSVEP data and symmetry detection threshold durations were modelled using log-normal distributions with weakly informative, $ \mathcal{N}(0, 2)$, priors.
+After fitting the model to the data, samples were drawn from the posterior distribution of the two datasets, for each wallpaper group.
+These samples were then recombined to calculate the distribution of differences for each of the 63 pairs of subgroup and supergroup.
 These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | data)$.
-
 For further technical details, please see the supplementary materials where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
@@
 %we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group. 
-%Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of P1 exemplars, that were matched in terms of their Fourier power spectrum.
+%Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of \textit{P1} exemplars, that were matched in terms of their Fourier power spectrum.
 
@@
 
-\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
+\acknow{PJK acknowledges funding support from the Canada First Research Excellence Fund and the National Sciences and Engineering Research Council of Canada.}
 

### 2021-01-27 analysis/analysis.Rmd

@@
 ```{r, fig.cap = "SSVEP RMS data from two of the wallpaper groups, P2 and P4M.
-Odd harmonics are shown in A and B, while even harmonic data are shown in C and D, and occipital and parietal regions of interest are shown in dark nad light gray, respectively.
+Odd harmonics are shown in A and B, while even harmonic data are shown in C and D, and occipital and parietal regions of interest are shown in dark and light gray, respectively.
 The two groups elicit very different response amplitudes for odd harmonics over occipital cortex, but for even harmonics those differences are much less pronounced."}
@@
 *  Fit a multi-level model using the lognormal distribution.
-  + independent variable (wallpaper group) and one dependent variable (rms or threshold).
-  + maximal random effect structure.
+  + Independent variable (wallpaper group) and one dependent variable (rms or threshold).
+  + Maximal random effect structure.
 * Extract mcmc samples from the posterior distributions.
@@
 
-```{r, fig.cap ="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals."}
+```{r, fig.cap ="The density plot shows the distribution of the empirical data, while the red line shows the 66% and 95% prediction intervals."}
 

### 2021-02-03 Writeup/manuscript.tex
Figure 1 text wrap
less critical discussion of correlation

@@
 % Please add here a significance statement to explain the relevance of your work
-\significancestatement{Wallpaper groups were discovered in the mid-19th century, and the 17 groups constitute the complete set of possible ways of regularly tiling the 2D-plane.
+\significancestatement{Wallpaper groups were discovered in the mid-19th century, and the 17 groups constitute the complete set of possible ways of regularly tiling the 2-D plane.
 In recent years wallpaper groups have found use in the vision science community, as an ideal stimulus set for studying the perception of symmetries in textures.
@@
 If this is the case, add \parshape=0 to the end of the list environment.
-\dropcap{S}ymmetries are present at many scales in images of natural scenes, due to a complex interplay of physical forces that govern pattern formation in nature.
-The importance of symmetry for visual perception has been known at least since the gestalt movement of the early 20th century.
-Since then, symmetry has been shown to contribute to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection \cite{RN1337}.
+\dropcap{S}ymmetries are present at many scales in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature.
+Symmetrical patterns have been created and appreciated by human cultures throughout history and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception.
+Symmetry contributes to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection \cite{RN1337}.
 Most of this work has focused on mirror symmetry or \textit{reflection}, with much less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection}.
-In the two spatial dimensions relevant for images, these four fundamental symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \cite{RN1562,RN1563,RN1425}.
-Previous work on a subset of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}.
+In the two spatial dimensions relevant for images, these four symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \cite{RN1562,RN1563,RN1425}.
+Previous work on a subset of four of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}.
 This effect was also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\cite{RN1725} or event-related paradigms \cite{RN1959}.
@@
 
-\begin{figure}[ptb]
-\centering
+\begin{figure}[h!tb]
 \includegraphics[width=1\linewidth]{../figures/figure1.png}
@@
 In each trial an exemplar image was shown with its matched control, one image after the other, and the order varied pseudo-randomly such that in half the trials the original exemplar was shown first, and in the other half the control image was shown first.
-After each trial, participants were told to indicate whether the first or second image contained more structure, and the duration of both images was controlled by a staircase procedure so that a threshold duration for symmetry detection could be computed for each wallpaper group.
+After each trial, participants were instructed to indicate whether the first or second image contained more structure.
+The duration of both images was controlled by a staircase procedure so that a threshold duration for symmetry detection could be computed for each wallpaper group.
 
-Examples of the wallpaper groups and a summary of our brain imaging and psychophysical measurements is in Figure 1.
+Examples of the wallpaper groups and a summary of our brain imaging and psychophysical measurements are shown in Figure 1.
 For our primary SSVEP analysis, we only considered EEG data from a pre-determined region-of-interest (ROI) consisting of six electrodes over occipital cortex (see Supplementary Figure 1.1).
@@
 
-Finally, we conducted a correlation analysis comparing SSVEP and psychophysical data, and found a modest ($R^2 = 0.44$) but above-zero correlation.
-The correlation between the two datasets is interesting because it suggests that our psychophysical and SSVEP measurements are at least to some extent tapping into the same underlying mechanisms.
+Finally, we conducted a correlation analysis comparing SSVEP and psychophysical data, and found a reliable correlation ($R^2 = 0.44$, Bayesian confidence interval [0.28, 0.55]).
+The correlation reflects an inverse relationship: For subgroup relationships where the supergroup produces a much \textit{larger} SSVEP amplitude than the subgroup, the supergroup also tends to produce a much \textit{smaller} symmetry detection threshold.
+This is consistent with our hypotheses about how the two measurements relate to symmetry representations in the brain, and suggests that our brain imaging and psychophysical measurements are at least to some extent tapping into the same underlying mechanisms.
 
@@
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that 56/63 subgroup relationships have $p(\Delta|data)>0.95$.}
+We can see that 55/63 subgroup relationships have $p(\Delta|data)>0.99$.}
 \label{fig:eeg_comp}
@@
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that 49/63 subgroup relationships have $p(\Delta|data)>0.95$.}
+We can see that 43/63 subgroup relationships have $p(\Delta|data)>0.99$.}
 \label{fig:durations_comp}
@@
 
-Here we show that beyond merely responding to the elementary symmetry operations of reflection \cite{RN1170} and rotation \cite{RN1725}, the visual system explicitly represents hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
+Here we show that beyond merely responding to the elementary symmetry operations of reflection \cite{RN1170} and rotation \cite{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
 Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \cite{RN1711}.
@@
 
-The current data provide a complete description of the visual system’s response to symmetries in the 2-D plane.
+The current data provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane.
 Our design precludes us from independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
@@
 The more controlled two-interval forced choice approach used in the current behavioral experiment allows us to show that more granular representations of wallpaper groups are measurable in behavior.
-The relatively modest correlation we observe between behavior and brain imaging data may suggest that the relationship between the two measurements is more complex than what can be captured in our summary statistics.
-Future work in which the same individuals participate in both behavioral and brain imaging experiments will help tease apart these complexities and enhance our understanding of the brain mechanisms that drive both measurements.
+
+We observe a reliable correlation between our brain imaging and psychophysical data.
+This suggests that the two measurements reflect the same underlying symmetry representations in visual cortex.
+While it should be noted that the correlation is relatively modest ($R^2 = 0.44$), we note that this may be partly due to the fact that the same individuals did not participate in the two experiments.
+Future work in which behavioral and brain imaging data are collected from the same participants, will help further establish the connection between the two measurements, and tease apart any additional complexity that may not have been captured by the summary statistics we applied here.
 
@@
 Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \cite{RN1764}.
-Wallpaper groups are visually compelling and have been widely used in human artistic expression going back to the Neolithic age \cite{RN1949}.
+Wallpaper groups are visually compelling and symmetries have been widely used in human artistic expression going back to the Neolithic age \cite{RN1949}.
 If wallpapers are in fact supertextures, this prevalence may be a direct result of the strategy the human visual system has adopted for texture encoding.

### 2021-02-04 analysis/analysis.Rmd
Figure 1 text wrap
changes some text about R2

@@
 
-We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, 95\% credible interval is `r round(bR2[3],2)` - `r round(bR2[4], 2)`).
+The confidence interval indicates that we can be confident that $R^2>0$ (i.e, 95\% credible interval is `r round(bR2[3],2)` - `r round(bR2[4], 2)`).
 

### 2021-02-04 Writeup/manuscript.tex
OSF updates

@@
 These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | data)$.
-For further technical details, please see the supplementary materials where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
+For further technical details, please see the Supplementary Materials at \url{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
 

### 2021-02-19 Writeup/kohler_clarke_2021.tex
reformatting latex

--- 2021-02-04 Writeup/manuscript.tex
@@
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
-% Use the lineno option to display guide line numbers if required.
+\documentclass[11pt, twoside]{article}
+\usepackage{boisik}
+\usepackage[OT1]{fontenc}
+\usepackage[affil-it]{authblk}
+\usepackage{graphicx}
+\usepackage{multicol}
+\usepackage{hyperref}
+\usepackage{natbib}
+\usepackage{geometry}
+ \geometry{a4paper, total={170mm,257mm}, left=20mm, top=20mm}
 
-\templatetype{pnasresearcharticle} % Choose template 
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} %= Template for a one-column mathematics article
-% {pnasinvited} %= Template for a PNAS invited submission
-% \usepackage{graphicx}
+\title{\huge The human visual system preserves the hierarchy of 2-dimensional pattern regularity}
+\author[1,2,3]{Peter J. Kohler}
+\author[4]{Alasdair D. F. Clarke}
 
-\title{The human visual system preserves the hierarchy of 2-dimensional pattern regularity}
+\affil[1]{\small York University, Department of Psychology, Toronto, ON M3J 1P3, Canada}
+\affil[2]{\small Centre for Vision Research, York University, Toronto, ON, M3J 1P3, Canada}
+\affil[3]{\small Stanford University, Department of Psychology, Stanford, CA 94305, United States}
+\affil[4]{\small University of Essex, Department of Psychology, Colchester, UK, CO4 3SQ}
 
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a, b, 1]{Peter J. Kohler}
-\author[c]{Alasdair D. F. Clarke}
+\date{}
 
-\affil[a]{York University, Department of Psychology, Toronto, ON M3J 1P3, Canada}
-\affil[b]{Centre for Vision Research, York University, Toronto, ON, M3J 1P3, Canada}
-\affil[c]{Stanford University, Department of Psychology, Stanford, CA 94305, United States}
-\affil[d]{University of Essex, Department of Psychology, Colchester, UK, CO4 3SQ}
+\begin{document}
 
-% Please give the surname of the lead author for the running footer
-\leadauthor{Kohler}
+\maketitle
 
-% Please add here a significance statement to explain the relevance of your work
-\significancestatement{Wallpaper groups were discovered in the mid-19th century, and the 17 groups constitute the complete set of possible ways of regularly tiling the 2-D plane.
-In recent years wallpaper groups have found use in the vision science community, as an ideal stimulus set for studying the perception of symmetries in textures.
-Here we present brain imaging and psychophysical data using the complete set of wallpaper groups and show that the hierarchical organization among wallpaper groups is reflected both in representations in visual cortex and performance on a symmetry detection task.
-This shows that the visual system is highly sensitive to regularities in textures, and suggest that symmetries may play an important role in texture perception.}
-
-% Please include corresponding author, author contribution and author declaration information
-\authorcontributions{PJK and ADFC designed the study, PJK collected EEG data, ADFC collected psychophysical data, PJK and ADFC wrote the paper.}
-\authordeclaration{The authors have no conflicts of interests to declare}
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: \href{mailto:pjkohler@syorku.ca}{pjkohler@yorku.ca}}
-
-% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}
-
-\begin{abstract}Symmetries are present at many scales in images of natural scenes.
+\begin{abstract}
+Symmetries are present at many scales in images of natural scenes.
 A large body of literature has demonstrated contributions of symmetry to numerous domains of visual perception.
@@
 This allows us to probe the hierarchy of complexity among wallpaper groups, in which simpler groups are subgroups of more complex ones.
-We find that this hierarchy is preserved almost perfectly in both behavior and brain activity: A multi-level Bayesian GLM indicates that for most of the 63 subgroup relationships, subgroups produce lower amplitude responses in visual cortex (posterior probability: >0.95 for 56 of 63) and require longer presentation durations to be reliably detected (posterior probability: >0.95 for 49 of 63).
+We find that this hierarchy is preserved almost perfectly in both behavior and brain activity: A multi-level Bayesian GLM indicates that for most of the 63 subgroup relationships, subgroups produce lower amplitude responses in visual cortex (posterior probability: $>0.95$ for 56 of 63) and require longer presentation durations to be reliably detected (posterior probability: $>0.95$ for 49 of 63).
 This systematic pattern is seen only in visual cortex and only in components of the brain response known to be symmetric-specific.
@@
 
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\begin{document}
-
-\maketitle
- \thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
-
-% If your first paragraph (i.e.
-with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
-If this is the case, add \parshape=0 to the end of the list environment.
-\dropcap{S}ymmetries are present at many scales in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature.
+Symmetries are present at many scales in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature.
 Symmetrical patterns have been created and appreciated by human cultures throughout history and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception.
@@
 
-\begin{figure}[h!tb]
-\includegraphics[width=1\linewidth]{../figures/figure1.png}
+\begin{figure}[hp]
+\centering
+\includegraphics[width=0.75\linewidth]{../figures/figure1.png}
 \caption{Examples of each of the 16 wallpaper groups are shown in the left- and right-most column of the figures, next to the corresponding SSVEP (center-left) and psychological (center-right) data from each group.
@@
 
-\begin{figure}[tbh]
+\begin{figure}[hp]
 \centering
-\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.png}
+\includegraphics[width=0.8\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.png}
 \caption{Posterior distributions for the difference in mean SSVEP RMS amplitude.
@@
 
-\begin{figure}[tbh]
+\begin{figure}[hptb]
 \centering
-\includegraphics[width=1\linewidth]{../analysis/plots/subgroup_comp_psychophysical.png}
+\includegraphics[width=0.8\linewidth]{../analysis/plots/subgroup_comp_psychophysical.png}
 \caption{Posterior distributions for the difference in mean symmetry detection threshold durations.
@@
 \centering
-\includegraphics[width=1\linewidth]{../analysis/plots/model_roc_style.pdf}
+\includegraphics[width=0.5\linewidth]{../analysis/plots/model_roc_style.pdf}
 \caption{This plot shows the proportion of subgroup relationships that satisfy $p(\Delta >0 | data) > x$.
@@
 \end{figure}
-
-% \section*{Guide to using this template on Overleaf}
-
-% Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout.
-For more detailed information please see the \href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information for Authors}.
-
-% If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}.
-You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.
-
-% \subsection*{Author Affiliations}
-
-% Include department, institution, and complete address, with the ZIP/postal code, for each author.
-Use lower case letters to match authors with institutions, as shown in the example.
-Authors with an ORCID ID may supply this information at submission.
-
-% \subsection*{Submitting Manuscripts}
-
-% All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}.
-If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.
-
-% \subsection*{Format}
-
-% Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References.
-Other orders and headings are permitted.
-
-% \subsection*{Manuscript Length}
-
-% PNAS generally uses a two-column format averaging 67 characters, including spaces, per line.
-The maximum length of a Direct Submission research article is six pages and a Direct Submission Plus research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.
-When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for Direct Submission Plus.
-
-% \subsection*{Data Archival}
-
-% PNAS must be able to archive the data essential to a published article.
-Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.
-
-% % \begin{SCfigure*}[\sidecaptionrelwidth][t]
-% % \centering
-% % % \includegraphics[width=11.4cm,height=11.4cm]{frog}
-% % \caption{This caption would be placed at the side of the figure, rather than below it.}\label{fig:side}
-% % \end{SCfigure*}
-
-% \subsection*{Digital Figures}
-
-% Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size.
-Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
-
-% Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
-
-% Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
-To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.
-
-% \subsection*{Tables}
-% In addition to including your tables within this manuscript file, PNAS requires that each table be uploaded to the submission separately as a “Table” file.
-Please ensure that each table .tex file contains a preamble, the \verb|\begin{document}| command, and the \verb|\end{document}| command.
-This is necessary so that the submission system can convert each file to PDF.
-
-% \subsection*{Single column equations}
-
-% Authors may use 1- or 2-column equations in their article, according to their preference.
-
-% To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.
-
-% Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used.
-
-% \begin{figure*}[bt!]
-% \begin{align*}
-% (x+y)^3&=(x+y)(x+y)^2\\
-%        &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
-%        &=x^3+3x^2y+3xy^3+x^3. 
-% \end{align*}
-% \end{figure*}
-
-% \begin{table}%[tbhp]
-% \centering
-% \caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
-% \begin{tabular}{lrrr}
-% Species & CBS & CV & G3 \\
-% \midrule
-% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
-% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
-% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
-% \bottomrule
-% \end{tabular}
-
-% \addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
-% \end{table}
-
-% \subsection*{Supporting Information (SI)}
-
-% Authors should submit SI as a single separate PDF file, combining all text, figures, tables, movie legends, and SI references.
-PNAS will publish SI uncomposed, as the authors have provided it.
-Additional details can be found here: \href{http://www.pnas.org/page/authors/journal-policies}{policy on SI}.
-For SI formatting instructions click \href{https://www.pnascentral.org/cgi-bin/main.plex?form_type=display_auth_si_instructions}{here}.
-The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
-Refer to the SI Appendix in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.
-
-% Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
-
-% \subsubsection*{SI Datasets}
-
-% Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
-
-% \matmethods{Please describe your materials and methods here.
-This can be more than one paragraph, and may contain subsections and equations as required.
-Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.
 
@@
 
-%\section*{snippets}
-%Specifically, the amplitudes of symmetry-specific responses in individual participants ($n=25$) preserve these relationships at an above-chance level in 88.3\% (53 out of 60) of cases.
-
-%we used a steady-state design, in which exemplar images belonging to 16 of the 17 wallpaper groups alternated with phase-scrambled images of the same group. 
-%Exemplars from each of the 16 groups alternated at 0.83 Hz with their corresponding set of \textit{P1} exemplars, that were matched in terms of their Fourier power spectrum.
-
-%Thus, the magnitude of the odd harmonic response components can be used as a distance metric for each group, with distance being measured relative to the simplest group, P1.
-
-\showmatmethods{} % Display the Materials and Methods section
-
-\acknow{PJK acknowledges funding support from the Canada First Research Excellence Fund and the National Sciences and Engineering Research Council of Canada.}
-
-\showacknow{} % Display the acknowledgments section
-
 % Bibliography
-\bibliographystyle{pnas-new} 
+\begin{multicols}{2}
+\small
+\bibliographystyle{plainnat} 
 \bibliography{literature}
+\end{multicols}
 

### 2021-02-19 Writeup/kohler_clarke_2021.tex
updated abstract

@@
 The four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in exactly 17 distinct ways.
-These \textit{wallpaper groups} represent the complete set of symmetries in 2D images.
-The goal of the current study is to provide a complete description of responses to symmetry in the human visual system, by collecting both brain and behavioral data using all 17 groups.
+These \textit{wallpaper groups} represent the complete set of symmetries in 2D images and have recently found use in the vision science community as an ideal stimulus set for studying the perception of symmetries in textures.
+The goal of the current study is to provide a more comprehensive description of responses to symmetry in the human visual system, by collecting both brain imaging (Steady-State Visual Evoked Potentials measured using high-density EEG) and behavioral (symmetry detection thresholds) data using the entire set of wallpaper groups.
 This allows us to probe the hierarchy of complexity among wallpaper groups, in which simpler groups are subgroups of more complex ones.
@@
 
-Symmetries are present at many scales in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature.
+Symmetries are abundant in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature.
 Symmetrical patterns have been created and appreciated by human cultures throughout history and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception.

### 2021-02-20 Writeup/kohler_clarke_2021.tex
Fixed citations and increased linespacing

@@
  \geometry{a4paper, total={170mm,257mm}, left=20mm, top=20mm}
+
+\renewcommand{\baselinestretch}{1.3}
 
@@
 Symmetrical patterns have been created and appreciated by human cultures throughout history and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception.
-Symmetry contributes to the perception of shapes \cite{RN1311,RN1682}, scenes \cite{RN1824} and surface properties \cite{RN1166}, as well as the social process of mate selection \cite{RN1337}.
+Symmetry contributes to the perception of shapes \citep{RN1311,RN1682}, scenes \citep{RN1824} and surface properties \citep{RN1166}, as well as the social process of mate selection \citep{RN1337}.
 Most of this work has focused on mirror symmetry or \textit{reflection}, with much less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection}.
-In the two spatial dimensions relevant for images, these four symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \cite{RN1562,RN1563,RN1425}.
-Previous work on a subset of four of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \cite{RN1725}.
-This effect was also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\cite{RN1725} or event-related paradigms \cite{RN1959}.
+In the two spatial dimensions relevant for images, these four symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \citep{RN1562,RN1563,RN1425}.
+Previous work on a subset of four of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \citep{RN1725}.
+This effect was also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\citep{RN1725} or event-related paradigms \citep{RN1959}.
 Here we extend this work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups.
@@
 A wallpaper group is a topologically discrete group of isometries of the Euclidean plane, i.e.
-transformations that preserve distance \cite{RN1425}.
+transformations that preserve distance \citep{RN1425}.
 Wallpaper groups differ in the number and kind of these transformations.
-In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \cite{RN1425}.
+In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \citep{RN1425}.
 The full list of subgroup relationships is listed in Section 1.4.2 of the Supplementary Material.
@@
 The index of a subgroup relationship is the number of cosets, i.e.
-the number of times the subgroup is found in the supergroup \cite{RN1425}.
+the number of times the subgroup is found in the supergroup \citep{RN1425}.
 As an example, let us consider groups \textit{P2} and \textit{P6}.
@@
 \textit{P2} is thus a subgroup of \textit{P6}, and \textit{P6} can be generated by combining \textit{P2} with rotations \{$0^{\circ}$, $120^{\circ}$, $240^{\circ}$\}.
-Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \cite{RN1425}.
-The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \cite{RN1711}.
+Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \citep{RN1425}.
+The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
 
@@
 The two datasets we present here make it possible to assess the extent to which both behavior and brain responses follow the hierachy of complexity expressed by the subgroup relationships.
-Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \cite{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
-For the psychophysical data, we hypothesized that more complex groups would lead to shorter symmetry detection thresholds, based on previous data showing that under a fixed presentation time, discriminability increases with the number of symmetry axes in the pattern \cite{wagemans_1991}.
+Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \citep{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
+For the psychophysical data, we hypothesized that more complex groups would lead to shorter symmetry detection thresholds, based on previous data showing that under a fixed presentation time, discriminability increases with the number of symmetry axes in the pattern \citep{wagemans_1991}.
 Our results confirm both hypotheses, and show that activity in human visual cortex is remarkably consistent with the hierarchical relationships between the wallpaper groups, with SSVEP amplitudes and psychophysical thresholds following these relationships at a level that is far beyond chance.
@@
 
-The stimuli used in our two experiments were generated from random-noise textures, which made it possible to generate multiple exemplars from each of the wallpaper groups, as described in detail elsewhere \cite{RN1725}.
+The stimuli used in our two experiments were generated from random-noise textures, which made it possible to generate multiple exemplars from each of the wallpaper groups, as described in detail elsewhere \citep{RN1725}.
 We generated control stimuli matched to each exemplar in the main stimulus set, by scrambling the phase but maintaining the power spectrum.
@@
 Phase scrambling maintains this periodicity so the phase-scrambled control images all belong to group \textit{P1} regardless of group membership of the original exemplar.
-\textit{P1} contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \cite{RN1425}.
+\textit{P1} contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \citep{RN1425}.
 In our SSVEP experiment, this stimulus set allowed us to isolate brain activity specific to the symmetry structure in the exemplar images from activity associated with modulation of low-level features, by alternating exemplar images and control exemplars.
-In this design, responses to structural features beyond the shared power spectrum, including any symmetries other than translation, are isolated in the odd harmonics of the image update frequency \cite{RN1725,RN1540, RN1954}.
+In this design, responses to structural features beyond the shared power spectrum, including any symmetries other than translation, are isolated in the odd harmonics of the image update frequency \citep{RN1725,RN1540, RN1954}.
 Thus, the combined magnitude of the odd harmonic response components can be used as a measure of the overall strength of the visual cortex response.
@@
 
-SSVEP data from four of the wallpaper groups (\textit{P2}, \textit{P3}, \textit{P4} and \textit{P6}) was previously published as part of our earlier demonstration of parametric responses to rotation symmetry in wallpaper groups\cite{RN1725}.
+SSVEP data from four of the wallpaper groups (\textit{P2}, \textit{P3}, \textit{P4} and \textit{P6}) was previously published as part of our earlier demonstration of parametric responses to rotation symmetry in wallpaper groups\citep{RN1725}.
 We replicate that result using our Bayesian approach, and find an analogous parametric effect in the psychophysical data (see Supplementary Figure 4.1).
@@
 
-Here we show that beyond merely responding to the elementary symmetry operations of reflection \cite{RN1170} and rotation \cite{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
-Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \cite{RN1711}.
-For the SSVEP, this remarkable consistency was specific to the odd harmonics of the stimulus frequency that are known to capture the symmetry-specific response \cite{RN1725} and to electrodes in a region-of-interest (ROI) over occipital cortex.
+Here we show that beyond merely responding to the elementary symmetry operations of reflection \citep{RN1170} and rotation \citep{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
+Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \citep{RN1711}.
+For the SSVEP, this remarkable consistency was specific to the odd harmonics of the stimulus frequency that are known to capture the symmetry-specific response \citep{RN1725} and to electrodes in a region-of-interest (ROI) over occipital cortex.
 When the same analysis was done using the odd harmonics from electrodes over parietal cortex (Supplementary Figure 3.3) or even harmonics from electrodes over occipital cortex (Supplementary Figure 3.4), the data was substantially less consistent with the subgroup relationships (yellow and green lines, Figure 4).
@@
 Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships.
-Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \cite{RN1253}, they tend to sort them into fewer groups than there actually are (4-12 groups) and it is common for exemplars from different wallpaper groups to be sorted in the same group \cite{RN172}.
+Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \citep{RN1253}, they tend to sort them into fewer groups than there actually are (4-12 groups) and it is common for exemplars from different wallpaper groups to be sorted in the same group \citep{RN172}.
 The more controlled two-interval forced choice approach used in the current behavioral experiment allows us to show that more granular representations of wallpaper groups are measurable in behavior.
@@
 The correspondence between responses in the visual system and group theory that we demonstrate here, may reflect a form of implicit learning that depends on the structure of the natural world.
-The environment is itself constrained by physical forces underlying pattern formation and these forces are subject to multiple symmetry constraints \cite{RN1634}.
+The environment is itself constrained by physical forces underlying pattern formation and these forces are subject to multiple symmetry constraints \citep{RN1634}.
 The ordered structure of responses to wallpaper groups could be driven by a central tenet of neural coding, that of efficiency.
-If coding is to be efficient, neural resources should be distributed to capture the structure of the environment with minimum redundancy considering the visual geometric optics, the capabilities of the subsequent neural coding stages and the behavioral goals of the organism \cite{RN1758, RN1760, RN1757, RN1756}.
-Early work within the efficient coding framework suggested that natural images had a $1/f$ spectrum and that the corresponding redundancy between pixels in natural images could be coded efficiently with a sparse set of oriented filter responses, such as those present in the early visual pathway \cite{RN1740, RN1446}.
+If coding is to be efficient, neural resources should be distributed to capture the structure of the environment with minimum redundancy considering the visual geometric optics, the capabilities of the subsequent neural coding stages and the behavioral goals of the organism \citep{RN1758, RN1760, RN1757, RN1756}.
+Early work within the efficient coding framework suggested that natural images had a $1/f$ spectrum and that the corresponding redundancy between pixels in natural images could be coded efficiently with a sparse set of oriented filter responses, such as those present in the early visual pathway \citep{RN1740, RN1446}.
 Our results suggest that the principle of efficient coding extends to a much higher level of structural redundancy – that of symmetries in visual images.
@@
 The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective and occlusion.
-Near-regular textures, however, abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \cite{RN1519}.
+Near-regular textures, however, abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \citep{RN1519}.
 The correspondence between visual cortex responses and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space.
-This framework resembles norm-based encoding strategies that have been proposed for other stimulus classes, most notably faces \cite{RN435}, and leads to the prediction that adaptation to wallpaper patterns should distort perception of near-regular textures, similar to the aftereffects found for faces \cite{RN1768}.
-Field biologists have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \cite{RN1775}.
+This framework resembles norm-based encoding strategies that have been proposed for other stimulus classes, most notably faces \citep{RN435}, and leads to the prediction that adaptation to wallpaper patterns should distort perception of near-regular textures, similar to the aftereffects found for faces \citep{RN1768}.
+Field biologists have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \citep{RN1775}.
 In the norm-based encoding framework, wallpaper groups can be considered \textit{supertextures}, exaggerated examples of the near-regular textures that surround us.
-Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \cite{RN1764}.
-Wallpaper groups are visually compelling and symmetries have been widely used in human artistic expression going back to the Neolithic age \cite{RN1949}.
+Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \citep{RN1764}.
+Wallpaper groups are visually compelling and symmetries have been widely used in human artistic expression going back to the Neolithic age \citep{RN1949}.
 If wallpapers are in fact supertextures, this prevalence may be a direct result of the strategy the human visual system has adopted for texture encoding.
@@
 \subsection*{Stimulus Generation}
-Exemplars from the different wallpaper groups were generated using a modified version of the methodology developed by Clarke and colleagues\cite{RN172} that we have described in detail elsewhere\cite{RN1725}.
+Exemplars from the different wallpaper groups were generated using a modified version of the methodology developed by Clarke and colleagues\citep{RN172} that we have described in detail elsewhere\citep{RN1725}.
 Briefly, exemplar patterns for each group were generated from random-noise textures, which were then repeated and transformed to cover the plane, according to the symmetry axes and geometric lattice specific to each group.
@@
 For the psychophysics experiment, the stimuli were shown on a $48 \times 27$cm VIEWPixx/3D LCD Display monitor, model VPX-VPX-2005C, resolution $1920 \times 1080$ pixels, with a viewing distance of approximately 40cm and linear gamma.
-Stimulus presentation was controlled using MatLab and Psychtoolbox-3 \cite{kleiner2007,brainard1997spatial}.
+Stimulus presentation was controlled using MatLab and Psychtoolbox-3 \citep{kleiner2007,brainard1997spatial}.
 The diameter of the circular aperture for the stimuli was $21.5^\circ$.
@@
 This process was repeated until the required accuracy of 9/10 was obtained.
-The rest of the block consisted of four interleaved staircases (using the QUEST algorithm \cite{watson1983quest}, full details given in the SI) of 30 trials each.
+The rest of the block consisted of four interleaved staircases (using the QUEST algorithm \citep{watson1983quest}, full details given in the SI) of 30 trials each.
 On average, a block of trials took around 10 minutes to complete.
@@
 The SSVEP paradigm we used allowed us to separate symmetry-related responses from non-specific contrast transient responses.
-Previous work has demonstrated that symmetry-related responses are predominantly found in the odd harmonics of the stimulus frequency, whereas the even harmonics consist mainly of responses unrelated to symmetry, that arise from the contrast change associated with the appearance of the second image \cite{RN1954, RN1725}.
+Previous work has demonstrated that symmetry-related responses are predominantly found in the odd harmonics of the stimulus frequency, whereas the even harmonics consist mainly of responses unrelated to symmetry, that arise from the contrast change associated with the appearance of the second image \citep{RN1954, RN1725}.
 This functional distinction of the harmonics allowed us to generate a single-cycle waveform containing the response specific to symmetry, by filtering out the even harmonics in the spectral domain, and then back-transforming the remaining signal, consisting only of odd harmonics, into the time-domain.
@@
 \subsection*{Defining the list of subgroup relationships}
-In order to get the complete list of subgroup relationships, we digitized Table 4 from Coxeter \cite{RN1711} (shown in Supplementary Table 1.2).
+In order to get the complete list of subgroup relationships, we digitized Table 4 from Coxeter \citep{RN1711} (shown in Supplementary Table 1.2).
 After removing identity relationships (i.e.
@@
 \subsection*{Bayesian Analysis of SSVEP and Psychophysical data}
-Bayesian analysis was carried out using R (v3.6.1) \cite{R} with the \texttt{brms} package (v2.9.0) \cite{burkner2017} and rStan (v2.19.2 \cite{rStan}).
+Bayesian analysis was carried out using R (v3.6.1) \citep{R} with the \texttt{brms} package (v2.9.0) \citep{burkner2017} and rStan (v2.19.2 \citep{rStan}).
 The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16 level factor, and random effects for participant.

### 2021-02-24 Writeup/kohler_clarke_2021.tex
W literature added

@@
 Future work in which behavioral and brain imaging data are collected from the same participants, will help further establish the connection between the two measurements, and tease apart any additional complexity that may not have been captured by the summary statistics we applied here.
+It has recently been demonstrated that \textit{W}, a measure of perceptual goodness derived from a holographic model of regularity \citep{van_der_helm_1996}, can predict EEG responses \citep{makin_w_2016} and perceptual discrimination performance \citep{nucci_2007} for patterns that contain symmetry and other types of regularity.
+The model was formulated based on dot patterns with symmetry axes centered on a single spatial location.
+It will important to determine if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane, and test how well it can explain behavioral and brain responses to wallpapers.
 
-We also find an of index for both our brain imaging measurements and our symmetry detection thresholds.
+We also find an effect of index for both our brain imaging measurements and our symmetry detection thresholds.
 This means that the visual system not only represents the hierarchical relationship captured by individual subgroups, but also distinguishes between subgroups depending on how many times the subgroup is repeated in the supergroup, with more repetitions leading to larger pairwise differences.
@@
 The two groups are comparable except \textit{PM} contains reflection and \textit{PG} contains glide reflection, and the former clearly elicits higher amplitudes and lower thresholds.
-The charting of how different symmetry types contribute to the representational hierarchy will be an important goal for future work.
+An important goal for future work will be to map out how different symmetry types contribute to the representational hierarchy.
 

### 2021-02-25 Writeup/kohler_clarke_2021.tex
Fixed references
There was a bibitem with am missing year field

--- 2021-02-20 Writeup/kohler_clarke_2021.tex
@@
 \usepackage{hyperref}
+\hypersetup{colorlinks=false}
 \usepackage{natbib}
@@
 \subsection*{Bayesian Analysis of SSVEP and Psychophysical data}
-Bayesian analysis was carried out using R (v3.6.1) \citep{R} with the \texttt{brms} package (v2.9.0) \citep{burkner2017} and rStan (v2.19.2 \citep{rStan}).
-The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16 level factor, and random effects for participant.
-The SSVEP data and symmetry detection threshold durations were modelled using log-normal distributions with weakly informative, $ \mathcal{N}(0, 2)$, priors.
-After fitting the model to the data, samples were drawn from the posterior distribution of the two datasets, for each wallpaper group.
-These samples were then recombined to calculate the distribution of differences for each of the 63 pairs of subgroup and supergroup.
-These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | data)$.
-For further technical details, please see the Supplementary Materials at \url{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
 
@@
 \small
-\bibliographystyle{plainnat} 
+\bibliographystyle{apalike} 
 \bibliography{literature}

### 2021-03-10 Writeup/kohler_clarke_2021.tex
Put bayesian methods back in!

--- 2021-02-24 Writeup/kohler_clarke_2021.tex
@@
 \usepackage{hyperref}
+\hypersetup{colorlinks=false}
 \usepackage{natbib}
 \usepackage{geometry}
+\usepackage{amsmath}
  \geometry{a4paper, total={170mm,257mm}, left=20mm, top=20mm}
@@
 \subsection*{Bayesian Analysis of SSVEP and Psychophysical data}
+
 Bayesian analysis was carried out using R (v3.6.1) \citep{R} with the \texttt{brms} package (v2.9.0) \citep{burkner2017} and rStan (v2.19.2 \citep{rStan}).
@@
 These samples were then recombined to calculate the distribution of differences for each of the 63 pairs of subgroup and supergroup.
-These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | data)$.
+These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | \text{data})$.
 For further technical details, please see the Supplementary Materials at \url{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.

### 2021-03-10 Writeup/kohler_clarke_2021.tex
fixed typo in citation.

@@
 \usepackage{multicol}
-\usepackage{hyperref}
-\hypersetup{colorlinks=false}
 \usepackage{natbib}
@@
 Future work in which behavioral and brain imaging data are collected from the same participants, will help further establish the connection between the two measurements, and tease apart any additional complexity that may not have been captured by the summary statistics we applied here.
-It has recently been demonstrated that \textit{W}, a measure of perceptual goodness derived from a holographic model of regularity \citep{van_der_helm_1996}, can predict EEG responses \citep{makin_w_2016} and perceptual discrimination performance \citep{nucci_2007} for patterns that contain symmetry and other types of regularity.
+It has recently been demonstrated that \textit{W}, a measure of perceptual goodness derived from a holographic model of regularity \citep{van_der_helm_1996}, can predict EEG responses \citep{makin_2016} and perceptual discrimination performance \citep{nucci_2007} for patterns that contain symmetry and other types of regularity.
 The model was formulated based on dot patterns with symmetry axes centered on a single spatial location.
@@
 These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | \text{data})$.
-For further technical details, please see the Supplementary Materials at \url{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
+For further technical details, please see the Supplementary Materials at \texttt{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
 

### 2021-03-11 Writeup/kohler_clarke_2021.tex
minor spelling mistakes corrected

@@
 
-The two datasets we present here make it possible to assess the extent to which both behavior and brain responses follow the hierachy of complexity expressed by the subgroup relationships.
+The two datasets we present here make it possible to assess the extent to which both behavior and brain responses follow the hierarchy of complexity expressed by the subgroup relationships.
 Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \citep{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
@@
 All wallpaper groups are inherently periodic because of their repeating lattice structure.
-Phase scrambling maintains this periodicity so the phase-scrambled control images all belong to group \textit{P1} regardless of group membership of the original exemplar.
+Phase scrambling maintains this periodicity, so the phase-scrambled control images all belong to group \textit{P1} regardless of group membership of the original exemplar.
 \textit{P1} contains no symmetries other than translation, while all other groups contain translation in combination with one or more of the other three fundamental symmetries (reflection, rotation, glide reflection) \citep{RN1425}.
@@
 We first fitted a Bayesian model with wallpaper group as a factor and participant as a random effect.
-We fit the model separately for SSVEP RMS and psychophysical data, and then computed posterior distributions for the difference between supergroup and subgroup.
+We fit the model separately for SSVEP RMS and psychophysical data and then computed posterior distributions for the difference between supergroup and subgroup.
 These difference distributions allowed us to compute the conditional probability that the supergroup would produce (a) larger RMS and (b) a shorter threshold durations, when compared to the subgroup.
@@
 For both control analyses (plotted in Supplementary Figures 3.3 and 3.4), the correspondence between data and subgroup relationships was substantially weaker than in the primary analysis.
-We can quantify the strength of the assocation between the data and the subgroup relationships, by asking what proportion of subgroup relationships that reach or exceed a range of probability thresholds.
+We can quantify the strength of the association between the data and the subgroup relationships, by asking what proportion of subgroup relationships that reach or exceed a range of probability thresholds.
 This is plotted in Figure 4, for our psychophysical data, our primary SSVEP analysis and our two control SSVEP analyses.
 It shows that odd-harmonic SSVEP data from the occipital ROI and symmetry detection threshold durations both have a strong association with the subgroup relationships such that a clear majority of the subgroups survive even at the highest threshold we consider ($p(\Delta>0|data)>0.99$).
-The assocation is far weaker for the two control analyses.
+The association is far weaker for the two control analyses.
 
@@
 We replicate that result using our Bayesian approach, and find an analogous parametric effect in the psychophysical data (see Supplementary Figure 4.1).
-We also conducted an analysis testing for an effect of index in our two datasets, and found that subgroup relationships with higher indices tended to produce greater pairwise differences between the subgroup and supergroup, for both SSVEP RMS and symmetry detection thresholds (see Supplementary Figure 4.2).
+We also conducted an analysis testing for an effect of index in our two datasets and found that subgroup relationships with higher indices tended to produce greater pairwise differences between the subgroup and supergroup, for both SSVEP RMS and symmetry detection thresholds (see Supplementary Figure 4.2).
 The effect of index is relatively weak, but the fact that there is a measurable index effect can nonetheless be taken as preliminary evidence that representations of symmetries in wallpaper groups may be compositional.
 
-Finally, we conducted a correlation analysis comparing SSVEP and psychophysical data, and found a reliable correlation ($R^2 = 0.44$, Bayesian confidence interval [0.28, 0.55]).
+Finally, we conducted a correlation analysis comparing SSVEP and psychophysical data and found a reliable correlation ($R^2 = 0.44$, Bayesian confidence interval [0.28, 0.55]).
 The correlation reflects an inverse relationship: For subgroup relationships where the supergroup produces a much \textit{larger} SSVEP amplitude than the subgroup, the supergroup also tends to produce a much \textit{smaller} symmetry detection threshold.
@@
 The model was formulated based on dot patterns with symmetry axes centered on a single spatial location.
-It will important to determine if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane, and test how well it can explain behavioral and brain responses to wallpapers.
+It will be important to determine if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane, and test how well it can explain behavioral and brain responses to wallpapers.
 
@@
 Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \citep{RN1764}.
-Wallpaper groups are visually compelling and symmetries have been widely used in human artistic expression going back to the Neolithic age \citep{RN1949}.
+Wallpaper groups are visually compelling, and symmetries have been widely used in human artistic expression going back to the Neolithic age \citep{RN1949}.
 If wallpapers are in fact supertextures, this prevalence may be a direct result of the strategy the human visual system has adopted for texture encoding.
@@
 This means that all control exemplars, regardless of which wallpaper group they are derived from, are transformed into another symmetry group, namely \textit{P1}.
-\textit{P1} is the simplest of the wallpaper groups, and contains only translations of a region whose shape derives from the lattice.
+\textit{P1} is the simplest of the wallpaper groups and contains only translations of a region whose shape derives from the lattice.
 Because the different wallpaper groups have different lattices, \textit{P1} controls matched to different groups have different power spectra.
@@
 In each trial, participants were presented with two stimuli (one of which was the wallpaper group for the current block of trials, the other being \textit{P1}), one after the other (inter-stimulus interval of 700ms).
-After each stimuli had been presented, it was masked with white noise for 300ms.
+After each stimulus had been presented, it was masked with white noise for 300ms.
 After both stimuli had been presented, participants made a response on the keyboard to indicate whether they thought the first or second image contained more symmetry.
@@
 Bayesian analysis was carried out using R (v3.6.1) \citep{R} with the \texttt{brms} package (v2.9.0) \citep{burkner2017} and rStan (v2.19.2 \citep{rStan}).
-The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16 level factor, and random effects for participant.
+The data from each experiment were modelled using a Bayesian generalised mixed effect model with wallpaper group being treated as a 16-level factor, and random effects for participant.
 The SSVEP data and symmetry detection threshold durations were modelled using log-normal distributions with weakly informative, $ \mathcal{N}(0, 2)$, priors.
@@
 These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | \text{data})$.
-For further technical details, please see the Supplementary Materials at \texttt{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predicitive checks, and model diagnoastics, can be found.
+For further technical details, please see the Supplementary Materials at \texttt{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predictive checks, and model diagnostics, can be found.
 

### 2021-05-04 Writeup/kohler_clarke_2021.tex
now with linenumbers
Added color package
\textcolor{ForestGreen}{example text goes here}
Adding notes on naming convention

@@
 \usepackage{amsmath}
+\usepackage{lineno}
+\usepackage[dvipsnames]{xcolor}
+
+\linenumbers
  \geometry{a4paper, total={170mm,257mm}, left=20mm, top=20mm}
@@
 transformations that preserve distance \citep{RN1425}.
-Wallpaper groups differ in the number and kind of these transformations.
+Wallpaper groups differ in the number and kind of these transformations \textcolor{OliveGreen}{and we can uniquely refer to different groups using crystallographic notation.
+Full details of this naming convention can be found XXXX, but in brief, most groups are notated by $pXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotational symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains \textit{m}irror or \textit{g}lide symmetry.
+For example, $P4$ contains 4 fold rotation, while $P2M$ contains 2 fold rotation and a mirror symmetry (see Figure XXX).}
+
 In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \citep{RN1425}.

### 2021-05-04 Writeup/kohler_clarke_2021.tex
naming convention notes.

--- 2021-03-11 Writeup/kohler_clarke_2021.tex
@@
 transformations that preserve distance \citep{RN1425}.
-Wallpaper groups differ in the number and kind of these transformations.
+Wallpaper groups differ in the number and kind of these transformations \textcolor{OliveGreen}{and we can uniquely refer to different groups using crystallographic notation.
+Full details of this naming convention can be found XXXX, but in brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotational symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains \textit{m}irror or \textit{g}lide symmetry.
+For example, $P4$ contains 4 fold rotation, while $P2M$ contains 2 fold rotation and a mirror symmetry (see Figure XXX).
+Two of the groups start with a $C$ rather than a $P$, ($CM$ and $CMM$) which indicates that the symmetries are specified relative to a cell that itself contains repetition.}
+
 In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \citep{RN1425}.

### 2021-05-05 Writeup/kohler_clarke_2021.tex
most of my answers to R1

--- 2021-03-11 Writeup/kohler_clarke_2021.tex
@@
 \usepackage{graphicx}
+\usepackage{textcomp}
 \usepackage{multicol}
@@
 \usepackage{amsmath}
+\usepackage{lineno}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{hyperref}
+
+\linenumbers
  \geometry{a4paper, total={170mm,257mm}, left=20mm, top=20mm}
@@
 transformations that preserve distance \citep{RN1425}.
-Wallpaper groups differ in the number and kind of these transformations.
+The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation.
+In brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotational symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains \textit{m}irror or \textit{g}lide symmetry.
+For example, $P4$ contains 4 fold rotation, while $P2M$ contains 2 fold rotation and a mirror symmetry (see Figure XXX).
+Two of the groups start with a $C$ rather than a $P$, ($CM$ and $CMM$) which indicates that the symmetries are specified relative to a cell that itself contains repetition.
+Full details of the naming convention and examples of the wallpaper groups can be found on \href{https://en.wikipedia.org/wiki/Wallpaper_group}{wikipedia}}.
+
 In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \citep{RN1425}.
@@
 The two datasets we present here make it possible to assess the extent to which both behavior and brain responses follow the hierarchy of complexity expressed by the subgroup relationships.
-Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \citep{RN1170, RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
+Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \citep{Sasaki_2005, Tyler_2005,RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
 For the psychophysical data, we hypothesized that more complex groups would lead to shorter symmetry detection thresholds, based on previous data showing that under a fixed presentation time, discriminability increases with the number of symmetry axes in the pattern \citep{wagemans_1991}.
@@
 
-Here we show that beyond merely responding to the elementary symmetry operations of reflection \citep{RN1170} and rotation \citep{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
+Here we show that beyond merely responding to the elementary symmetry operations of reflection \citep{Sasaki_2005, Tyler_2005} and rotation \citep{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
 Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \citep{RN1711}.
@@
 Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships.
-Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \citep{RN1253}, they tend to sort them into fewer groups than there actually are (4-12 groups) and it is common for exemplars from different wallpaper groups to be sorted in the same group \citep{RN172}.
+\textcolor{ForestGreen}{Previous brain-imaging studies have found evidence of parametric responses with the number of reflection symmetry folds \cite{keefe_2018, Sasaki_2005, makin_2016} and with the order of rotation symmetry \cite{RN1725}.
+Our study is the first demonstration that the brain encodes symmetry in this parametric fashion across every possible combination of different \textit{symmetry types}, and that this parametric encoding is also reflected in behavior.} Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \citep{RN1253}, they tend to sort them into fewer groups than there actually are (4-12 groups) and it is common for exemplars from different wallpaper groups to be sorted in the same group \citep{RN172}.
 The more controlled two-interval forced choice approach used in the current behavioral experiment allows us to show that more granular representations of wallpaper groups are measurable in behavior.
+
+\textcolor{ForestGreen}{A large literature exists on the \textit{Sustained Posterior Negativity} (SPN), a characteristic negative-going waveform that is known to reflect responses to symmetry and other forms of regularity and structure \citep{makin_2016}.
+The SPN scales with the proportion of reflection symmetry in displays that contain a mixture of symmetry and noise \cite{makin_formation_2020, palumbo_scaling_2015}, and both reflection, rotation and translation can produce a measurable SPN \cite{makin_electrophysiological_2013}.
+It has recently been demonstrated that a holographic model of regularity \citep{van_der_helm_1996}, can predict both SPN amplitude \citep{makin_2016} and perceptual discrimination performance \citep{nucci_2007} for dot patterns that contain symmetry and other types of regularity.
+The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods of isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure 1), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study that has been published on the wallpaper groups also produced SPN-like waveforms \citep{RN1959}.
+Future work should more firmly establish the connection and determine if the SPN can capture responses similarly precise symmetry responses as the SSVEPs presented here.
+It would also be worthwhile to ask if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane.}
 
@@
 This suggests that the two measurements reflect the same underlying symmetry representations in visual cortex.
-While it should be noted that the correlation is relatively modest ($R^2 = 0.44$), we note that this may be partly due to the fact that the same individuals did not participate in the two experiments.
-Future work in which behavioral and brain imaging data are collected from the same participants, will help further establish the connection between the two measurements, and tease apart any additional complexity that may not have been captured by the summary statistics we applied here.
-It has recently been demonstrated that \textit{W}, a measure of perceptual goodness derived from a holographic model of regularity \citep{van_der_helm_1996}, can predict EEG responses \citep{makin_2016} and perceptual discrimination performance \citep{nucci_2007} for patterns that contain symmetry and other types of regularity.
-The model was formulated based on dot patterns with symmetry axes centered on a single spatial location.
-It will be important to determine if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane, and test how well it can explain behavioral and brain responses to wallpapers.
+\textcolor{ForestGreen}{It should be noted that the correlation is relatively modest ($R^2 = 0.44$).
+This may be partly due to the fact that different individuals participated in the two experiments.
+It may also be related to the fact that participants where not doing a symmetry-related task during the SSVEP experiment, but instead monitored the stimuli for brief, intermittent contrast changes.
+Previous brain imaging studies have found enhanced reflection symmetry responses when participants performed a symmetry-related task \cite{makin_formation_2020, Sasaki_2005, keefe_2018}.
+We did not manipulate task during our SSVEP recordings, but it is possible that adding a symmetry-related task to our SSVEP experiment would have produced measurements that reflected subgroup relationships to an even higher extent.
+We note, however, that our performance is already close to ceiling (see Figure 3).
+It is possible that adding a symmetry-related task would enhance SSVEP amplitudes across all wallpaper groups without improving the discriminality of individual groups (similar to what was observed by Keefe and his co-authors \cite{keefe_2018}).
+SPN measurements suggest that task-driven processing may be important for detecting symmetries that have been subject to perspective distortion \cite{makin_conditions_2015} although it should be noted that this effect was much less clear in a subsequent functional MRI study \cite{keefe_2018}.
+Future work in which behavioral and brain imaging data are collected from the same participants, and behavior is manipulated in the SSVEP task, will help further establish the connection between the two measurements, and elucidate the potential contribution of task-related top-down processing to the current results.}
 
@@
 
-The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective and occlusion.
-Near-regular textures, however, abound in the visual world, and can be approximated as deformed versions of the wallpaper groups \citep{RN1519}.
+The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective \textcolor{ForestGreen}{(see above)} and occlusion.
+Near-regular textures, however, abound in the visual world, \textcolor{ForestGreen}{and can be modeled} as deformed versions of the wallpaper groups \citep{RN1519}.
 The correspondence between visual cortex responses and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space.
@@
 \subsection*{Participants}
-Twenty-five participants (11 females, mean age 28.7±13.3) took part in the EEG experiment.
+Twenty-five participants (11 females, mean age 28.7 \textpm\ 3.3) took part in the EEG experiment.
 Their informed consent was obtained before the experiment under a protocol that was approved by the Institutional Review Board of Stanford University.
-11 participants (8 females, mean age 20.73±1.21) took part in the psychophysics experiment.
+11 participants (8 females, mean age 20.73 \textpm\ 1.21) took part in the psychophysics experiment.
 All participants had normal or corrected-to-normal vision.

### 2021-05-05 Writeup/kohler_clarke_2021.tex
remove boxes around hyperlinks
better hyperlinks

@@
 \usepackage{hyperref}
+\usepackage{xcolor}
+\hypersetup{
+    colorlinks,
+    linkcolor={red!50!black},
+    citecolor={blue!50!black},
+    urlcolor={blue!80!black}
+}
 
@@
 These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | \text{data})$.
-For further technical details, please see the Supplementary Materials at \texttt{https://osf.io/f3ex8/} where the full R code, model specification, prior and posterior predictive checks, and model diagnostics, can be found.
+For further technical details, please see \href{https://osf.io/f3ex8/}{the Supplementary Materials} where the full R code, model specification, prior and posterior predictive checks, and model diagnostics, can be found.
 

### 2021-05-11 Writeup/kohler_clarke_2021.tex
add new figure 1 to paper

@@
 
+\begin{figure}[hptb]
+\centering
+\includegraphics[width=0.75\linewidth]{../figures/figure1.pdf}
+\caption{\textcolor{ForestGreen}{Subgroup relationships within indices 2 (solid lines) and 3 (dashed line) are shown in (A).
+All other relationships can we deduced by identifying the shortest path through the hierarchy, and multiplying the subgroup indices.
+For example, \textit{P1} is related to \textit{P6} through \textit{P6}\textrightarrow\textit{P3} (index 2) and \textit{P3}\textrightarrow\textit{P1} (index 3) so \textit{P1} is also a subgroup of \textit{P6} with index $3 \times 2 = 6$.
+We also show enlarged versions of some of the subgroups relationships involving \textit{PMM} (B, shown in blue) and \textit{P6} (C, shown in red) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
+\end{figure}
+
 \begin{figure}[hp]
 \centering
-\includegraphics[width=0.75\linewidth]{../figures/figure1.png}
+\includegraphics[width=0.75\linewidth]{../figures/figure2.png}
 \caption{Examples of each of the 16 wallpaper groups are shown in the left- and right-most column of the figures, next to the corresponding SSVEP (center-left) and psychological (center-right) data from each group.
@@
 
-Examples of the wallpaper groups and a summary of our brain imaging and psychophysical measurements are shown in Figure 1.
+Examples of the wallpaper groups and a summary of our brain imaging and psychophysical measurements are shown in Figure 2.
 For our primary SSVEP analysis, we only considered EEG data from a pre-determined region-of-interest (ROI) consisting of six electrodes over occipital cortex (see Supplementary Figure 1.1).
@@
 To reduce the amplitude of the symmetry-specific response to a single number that could be used in further analyses and compared to the psychophysical data, we computed the root-mean-square (RMS) over the odd-harmonic-filtered waveforms.
-The data in Figure 1 are shown in descending order according to RMS.
-The psychophysical results, shown in box plots in Figure 1, were also quite variable between groups, and there seems to be a general pattern where wallpaper groups near the top of the figure, that have lower SSVEP amplitudes, also have longer psychophysical threshold durations.
+The data in Figure 2 are shown in descending order according to RMS.
+The psychophysical results, shown in box plots in Figure 2, were also quite variable between groups, and there seems to be a general pattern where wallpaper groups near the top of the figure, that have lower SSVEP amplitudes, also have longer psychophysical threshold durations.
 
@@
 These difference distributions allowed us to compute the conditional probability that the supergroup would produce (a) larger RMS and (b) a shorter threshold durations, when compared to the subgroup.
-The posterior distributions are shown in Figure 2 for the SSVEP data, and in Figure 3 for the psychophysical data, which distributions color-coded according to conditional probability.
+The posterior distributions are shown in Figure 3 for the SSVEP data, and in Figure 4 for the psychophysical data, which distributions color-coded according to conditional probability.
 For both data sets our hypothesis is confirmed: For the overwhelming majority of the 63 subgroup relationships, supergroups are more likely to produce larger symmetry-specific SSVEPs and shorter symmetry detection threshold durations, and in most cases the conditional probability of this happening is extremely high.
@@
 We can quantify the strength of the association between the data and the subgroup relationships, by asking what proportion of subgroup relationships that reach or exceed a range of probability thresholds.
-This is plotted in Figure 4, for our psychophysical data, our primary SSVEP analysis and our two control SSVEP analyses.
+This is plotted in Figure 5, for our psychophysical data, our primary SSVEP analysis and our two control SSVEP analyses.
 It shows that odd-harmonic SSVEP data from the occipital ROI and symmetry detection threshold durations both have a strong association with the subgroup relationships such that a clear majority of the subgroups survive even at the highest threshold we consider ($p(\Delta>0|data)>0.99$).
@@
 \centering
-\includegraphics[width=0.8\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.png}
+\includegraphics[width=0.75\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.png}
 \caption{Posterior distributions for the difference in mean SSVEP RMS amplitude.
@@
 \centering
-\includegraphics[width=0.8\linewidth]{../analysis/plots/subgroup_comp_psychophysical.png}
+\includegraphics[width=0.75\linewidth]{../analysis/plots/subgroup_comp_psychophysical.png}
 \caption{Posterior distributions for the difference in mean symmetry detection threshold durations.
@@
 For the SSVEP, this remarkable consistency was specific to the odd harmonics of the stimulus frequency that are known to capture the symmetry-specific response \citep{RN1725} and to electrodes in a region-of-interest (ROI) over occipital cortex.
-When the same analysis was done using the odd harmonics from electrodes over parietal cortex (Supplementary Figure 3.3) or even harmonics from electrodes over occipital cortex (Supplementary Figure 3.4), the data was substantially less consistent with the subgroup relationships (yellow and green lines, Figure 4).
+When the same analysis was done using the odd harmonics from electrodes over parietal cortex (Supplementary Figure 3.3) or even harmonics from electrodes over occipital cortex (Supplementary Figure 3.4), the data was substantially less consistent with the subgroup relationships (yellow and green lines, Figure 5).
 
 The current data provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane.
-Our design precludes us from independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 1), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
+Our design precludes us from independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 2), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
 The subgroup relationships are in many cases not obvious perceptually, and most participants had no knowledge of group theory.
@@
 It has recently been demonstrated that a holographic model of regularity \citep{van_der_helm_1996}, can predict both SPN amplitude \citep{makin_2016} and perceptual discrimination performance \citep{nucci_2007} for dot patterns that contain symmetry and other types of regularity.
-The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods of isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure 1), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study that has been published on the wallpaper groups also produced SPN-like waveforms \citep{RN1959}.
+The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods of isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure 2), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study that has been published on the wallpaper groups also produced SPN-like waveforms \citep{RN1959}.
 Future work should more firmly establish the connection and determine if the SPN can capture responses similarly precise symmetry responses as the SSVEPs presented here.
@@
 We did not manipulate task during our SSVEP recordings, but it is possible that adding a symmetry-related task to our SSVEP experiment would have produced measurements that reflected subgroup relationships to an even higher extent.
-We note, however, that our performance is already close to ceiling (see Figure 3).
+We note, however, that our performance is already close to ceiling (see Figure 5).
 It is possible that adding a symmetry-related task would enhance SSVEP amplitudes across all wallpaper groups without improving the discriminality of individual groups (similar to what was observed by Keefe and his co-authors \cite{keefe_2018}).
@@
 This is perhaps because the index analysis does not take into account the \textit{type} of isometries that differentiate the subgroup and supergroup.
-The effect of symmetry type can be observed by contrasting the measured SSVEP amplitudes and detection thresholds for groups \textit{PM} and \textit{PG} in Figure 1.
+The effect of symmetry type can be observed by contrasting the measured SSVEP amplitudes and detection thresholds for groups \textit{PM} and \textit{PG} in Figure 2.
 The two groups are comparable except \textit{PM} contains reflection and \textit{PG} contains glide reflection, and the former clearly elicits higher amplitudes and lower thresholds.
@@
 For our main analysis, we averaged the odd harmonic single-cycle waveforms within a six-electrode region of interest (ROI) over occipital cortex (electrodes 70, 74, 75, 81, 82, 83).
-These waveforms, averaged over participants, are shown in Figure 1.
+These waveforms, averaged over participants, are shown in Figure 2.
 The same analysis was done for the even harmonics and for the odd harmonics within a six electrode ROI over parietal cortex (electrodes 53, 54, 61, 78, 79, 86; see Supplementary Figure 1.1).

### 2021-05-11 Writeup/kohler_clarke_2021.tex
minor spelling mistakes

@@
 \caption{\textcolor{ForestGreen}{Subgroup relationships within indices 2 (solid lines) and 3 (dashed line) are shown in (A).
-All other relationships can we deduced by identifying the shortest path through the hierarchy, and multiplying the subgroup indices.
+All other relationships can be infered by identifying the shortest path through the hierarchy, and multiplying the subgroup indices.
 For example, \textit{P1} is related to \textit{P6} through \textit{P6}\textrightarrow\textit{P3} (index 2) and \textit{P3}\textrightarrow\textit{P1} (index 3) so \textit{P1} is also a subgroup of \textit{P6} with index $3 \times 2 = 6$.
-We also show enlarged versions of some of the subgroups relationships involving \textit{PMM} (B, shown in blue) and \textit{P6} (C, shown in red) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
+We also show enlarged versions of some of the subgroup relationships involving \textit{PMM} (B, shown in blue) and \textit{P6} (C, shown in red) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
 \end{figure}

### 2021-05-11 Writeup/kohler_clarke_2021.tex
Fixed 63 - 64 mistake in fig4 caption

--- 2021-05-05 Writeup/kohler_clarke_2021.tex
@@
 \caption{This plot shows the proportion of subgroup relationships that satisfy $p(\Delta >0 | data) > x$.
-We can see that if we take $x = 0.95$ as our threshold, the subgroup relationships are preserved in 56/63 = 89\% and 49/64 = 78\% of the comparisons for the primary SSVEP and threshold duration datasets, repectively.
-This compares to the 32/64= 50\% and 22/64 = 35\% for the SSVEP control datasets.}
+We can see that if we take $x = 0.95$ as our threshold, the subgroup relationships are preserved in 56/63 = 89\% and \textcolor{ForestGreen}{48/63 = 76\%} of the comparisons for the primary SSVEP and threshold duration datasets, receptively.
+This compares to the \textcolor{ForestGreen}{32/63= 51\%} and \textcolor{ForestGreen}{22/63 = 35\%} for the SSVEP control datasets.}
 \label{fig:durations_rotations}

### 2021-05-13 Writeup/kohler_clarke_2021.tex
addressed R2 comments, and two of the editor queries

@@
 Symmetry contributes to the perception of shapes \citep{RN1311,RN1682}, scenes \citep{RN1824} and surface properties \citep{RN1166}, as well as the social process of mate selection \citep{RN1337}.
-Most of this work has focused on mirror symmetry or \textit{reflection}, with much less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection}.
-In the two spatial dimensions relevant for images, these four symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \citep{RN1562,RN1563,RN1425}.
+Most of this work has focused on mirror symmetry or \textit{reflection}, with less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection} \citep{wagemans_1998}.
+\textcolor{ForestGreen}{This is perhaps because that reflection has been found to be more perceptually salient \citep{mach_1959,royer_1981,palmer_1991,ogden_2016,hamada_1988} and produced more brain activity \citep{makin_2013,makin_2014,makin_2012,wright_2015} than other types of symmetry.
+The current study considers all of the four fundamental symmetries not in isolation, but when they are combined in regular textures.} In the two spatial dimensions relevant for images, symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \citep{RN1562,RN1563,RN1425}.
 Previous work on a subset of four of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \citep{RN1725}.
-This effect was also robust with electroencephalography (EEG), whether measured using Steady-State Visual Evoked Potentials (SSVEPs)\citep{RN1725} or event-related paradigms \citep{RN1959}.
-Here we extend this work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups.
+This effect was also robust \textcolor{ForestGreen}{when symmetry responses were measured with electroencephalography (EEG) using both Steady-State Visual Evoked Potentials (SSVEPs)\citep{RN1725} and Event-Related Potentials \citep{RN1959}.
+The SSVEP technique uses periodic visual stimulation to produce a periodic brain response that is confined to integer multiples of the stimulation frequency known as harmonics.
+SSVEP response harmonics can be isolated in the frequency domain and depending on the specific design, different harmonics will in some cases express different aspects of the brain response.
+\citep{RN1540}.} Here we extend on the previous work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups.
 We measure responses in visual cortex to 16 out of the 17 wallpaper groups, with the 17th serving as a control stimulus.
@@
 \includegraphics[width=0.75\linewidth]{../figures/figure1.pdf}
-\caption{\textcolor{ForestGreen}{Subgroup relationships within indices 2 (solid lines) and 3 (dashed line) are shown in (A).
+\caption{\textcolor{ForestGreen}{Subgroup relationships with indices 2 (solid lines) and 3 (dashed line) are shown in (A).
 All other relationships can be infered by identifying the shortest path through the hierarchy, and multiplying the subgroup indices.
@@
 
-The current data provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane.
-Our design precludes us from independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 2), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
+\textcolor{ForestGreen}{The current study uses 16 distinct wallpaper groups, while previous neuroimaging studies focused on a subset of 4 \citep{RN1725,RN1959}.
+This represents a significant conceptual advance, because it makes it possible to investigate the complete subgroup hierarchy among the 17 groups and ask to what extent the hierarchy is reflected in brain activity.
+Our data} provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane.
+\textcolor{ForestGreen}{We do not} independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 2), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
 The subgroup relationships are in many cases not obvious perceptually, and most participants had no knowledge of group theory.
@@
 \textcolor{ForestGreen}{Previous brain-imaging studies have found evidence of parametric responses with the number of reflection symmetry folds \cite{keefe_2018, Sasaki_2005, makin_2016} and with the order of rotation symmetry \cite{RN1725}.
-Our study is the first demonstration that the brain encodes symmetry in this parametric fashion across every possible combination of different \textit{symmetry types}, and that this parametric encoding is also reflected in behavior.} Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \citep{RN1253}, they tend to sort them into fewer groups than there actually are (4-12 groups) and it is common for exemplars from different wallpaper groups to be sorted in the same group \citep{RN172}.
-The more controlled two-interval forced choice approach used in the current behavioral experiment allows us to show that more granular representations of wallpaper groups are measurable in behavior.
+Our study is the first demonstration that the brain encodes symmetry in this parametric fashion across every possible combination of different \textit{symmetry types}, and that this parametric encoding is also reflected in behavior.} Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \citep{RN1253}, they tend to sort exemplars into fewer \textcolor{ForestGreen}{(4-12) sets than the number of wallpaper groups, often placing exemplars from different wallpaper groups in the same set \citep{RN172}.
+The two-interval forced choice approach we use in the current psychophysical experiment makes it possible to directly compare symmetry detection thresholds to the subgroup hierarchy, and reveals that not only can the 17 wallpaper groups be distinguished based on behavioral data, behavior largely follows the subgroup hierarchy.}
 
 \textcolor{ForestGreen}{A large literature exists on the \textit{Sustained Posterior Negativity} (SPN), a characteristic negative-going waveform that is known to reflect responses to symmetry and other forms of regularity and structure \citep{makin_2016}.
-The SPN scales with the proportion of reflection symmetry in displays that contain a mixture of symmetry and noise \cite{makin_formation_2020, palumbo_scaling_2015}, and both reflection, rotation and translation can produce a measurable SPN \cite{makin_electrophysiological_2013}.
+The SPN scales with the proportion of reflection symmetry in displays that contain a mixture of symmetry and noise \cite{makin_formation_2020, palumbo_scaling_2015}, and both reflection, rotation and translation can produce a measurable SPN \cite{makin_2013}.
 It has recently been demonstrated that a holographic model of regularity \citep{van_der_helm_1996}, can predict both SPN amplitude \citep{makin_2016} and perceptual discrimination performance \citep{nucci_2007} for dot patterns that contain symmetry and other types of regularity.
@@
 This may be partly due to the fact that different individuals participated in the two experiments.
-It may also be related to the fact that participants where not doing a symmetry-related task during the SSVEP experiment, but instead monitored the stimuli for brief, intermittent contrast changes.
+It may also be related to the fact that participants where not doing a symmetry-related task during the SSVEP experiment, but instead monitored the stimuli for brief changes in contrast that occured twice per trial (see Methods).
 Previous brain imaging studies have found enhanced reflection symmetry responses when participants performed a symmetry-related task \cite{makin_formation_2020, Sasaki_2005, keefe_2018}.
@@
 Their informed consent was obtained before the experiment under a protocol that was approved by the University of Essex's Ethics Committee.
+\textcolor{ForestGreen}{There was no overlap in participants between the EEG and psychophysics experiments.}
 
@@
 To control vigilance, a contrast dimming task was employed.
-Two times per trial, an image pair was shown at reduced contrast, and the participants were instructed to press a button on a response pad.
-We adjusted the contrast reduction such that average accuracy for each participant was kept at ~85\% correct, in order to keep the difficulty of the vigilance at a constant level.
+Two times per trial, an image pair \textcolor{ForestGreen}{(control {P1} plus exemplar)} was shown at reduced contrast.
+\textcolor{ForestGreen}{P}articipants were instructed to press a button on a response pad \textcolor{ForestGreen}{whenever they noticed a contrast change.
+Reaction times were not taken into account and participants were told to respond at their own pace while being as accurate as possible}.
+We adjusted the \textcolor{ForestGreen}{reduction in contrast} such that average accuracy for each participant was kept at ~85\% correct, in order to keep the difficulty of the vigilance \textcolor{ForestGreen}{task} at a constant level.
 

### 2021-05-14 Writeup/kohler_clarke_2021.tex
edits that correspond to PK responses to Editors and Reviewers

@@
 Symmetries are present at many scales in images of natural scenes.
-A large body of literature has demonstrated contributions of symmetry to numerous domains of visual perception.
+\textcolor{ForestGreen}{Humans and other animals are highly sensitive to visual symmetry, and a large} literature has demonstrated contributions of symmetry to numerous domains of visual perception.
 The four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in exactly 17 distinct ways.
@@
 This allows us to probe the hierarchy of complexity among wallpaper groups, in which simpler groups are subgroups of more complex ones.
-We find that this hierarchy is preserved almost perfectly in both behavior and brain activity: A multi-level Bayesian GLM indicates that for most of the 63 subgroup relationships, subgroups produce lower amplitude responses in visual cortex (posterior probability: $>0.95$ for 56 of 63) and require longer presentation durations to be reliably detected (posterior probability: $>0.95$ for 49 of 63).
+We find that this hierarchy is preserved almost perfectly in both behavior and brain activity: A multi-level Bayesian GLM indicates that for most of the 63 subgroup relationships, subgroups produce lower amplitude responses in visual cortex (posterior probability: $>0.95$ for 56 of 63) and require longer presentation durations to be reliably detected (posterior probability: $>0.95$ for \textcolor{ForestGreen}{48} of 63).
 This systematic pattern is seen only in visual cortex and only in components of the brain response known to be symmetric-specific.
@@
 Symmetries are abundant in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature.
-Symmetrical patterns have been created and appreciated by human cultures throughout history and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception.
-Symmetry contributes to the perception of shapes \citep{RN1311,RN1682}, scenes \citep{RN1824} and surface properties \citep{RN1166}, as well as the social process of mate selection \citep{RN1337}.
-Most of this work has focused on mirror symmetry or \textit{reflection}, with less attention being paid to the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection} \citep{wagemans_1998}.
-\textcolor{ForestGreen}{This is perhaps because that reflection has been found to be more perceptually salient \citep{mach_1959,royer_1981,palmer_1991,ogden_2016,hamada_1988} and produced more brain activity \citep{makin_2013,makin_2014,makin_2012,wright_2015} than other types of symmetry.
-The current study considers all of the four fundamental symmetries not in isolation, but when they are combined in regular textures.} In the two spatial dimensions relevant for images, symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \citep{RN1562,RN1563,RN1425}.
+\textcolor{ForestGreen}{Sensitivity to symmetry has been demonstrated in a number of species, includes bees \citep{giurfa_insect_symmetry_1996}, fish \citep{morris_swordtail_symmetry_1998,schluter_sailfin_symmetry_1998}, birds \citep{RN1337,swaddle_symmetry_finches_1994} and dolphins \citep{von_fersen_dolphin_symmetry_1992}, and may be used as a cue for mate selection in many species \citep{swaddle_asymmetry_review_1999} including humans \citep{rhodes_facial_1998}.
+Humans cultures have created and appreciated s}ymmetrical patterns throughout history, and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception.
+Symmetry contributes to the perception of shapes \citep{RN1311,RN1682}, scenes \citep{apthorp_symmetry_numerosity} and surface properties \citep{RN1166}.
+\textcolor{ForestGreen}{This literature is almost exclusively based on stimuli in which one or more symmetry axes are placed at a single point in the image.
+Focus has been on} mirror symmetry or \textit{reflection}, \textcolor{ForestGreen}{with relatively few studies} including the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection} \citep{wagemans_1998} \textcolor{ForestGreen}{- perhaps because reflection has been found to be more perceptually salient \citep{mach_1959,royer_1981,palmer_1991,ogden_2016,hamada_1988} and produce more brain activity \citep{makin_2013,makin_2014,makin_2012,wright_2015}.
+In the current study, we take a different approach by investigating visual processing of regular textures in which combinations of the four fundamental symmetries tile the 2D plane.}
+
+In the two spatial dimensions relevant for images, symmetries can be combined in 17 distinct ways, \textit{the wallpaper groups} \citep{RN1562,RN1563,RN1425}.
 Previous work on a subset of four of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \citep{RN1725}.
@@
 The SSVEP technique uses periodic visual stimulation to produce a periodic brain response that is confined to integer multiples of the stimulation frequency known as harmonics.
-SSVEP response harmonics can be isolated in the frequency domain and depending on the specific design, different harmonics will in some cases express different aspects of the brain response.
+SSVEP response harmonics can be isolated in the frequency domain and depending on the specific design, different harmonics will express different aspects of the brain response.
 \citep{RN1540}.} Here we extend on the previous work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups.
@@
 transformations that preserve distance \citep{RN1425}.
-The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation.
-In brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotational symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains \textit{m}irror or \textit{g}lide symmetry.
-For example, $P4$ contains 4 fold rotation, while $P2M$ contains 2 fold rotation and a mirror symmetry (see Figure XXX).
+The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation (CITATION?).
+In brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotation symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains reflection (m) or glide reflection (g).
+For example, $P4$ contains 4 fold rotation, while $P2M$ contains 2 fold rotation and one reflection (see Figure XXX).
 Two of the groups start with a $C$ rather than a $P$, ($CM$ and $CMM$) which indicates that the symmetries are specified relative to a cell that itself contains repetition.
@@
 the number of times the subgroup is found in the supergroup \citep{RN1425}.
-As an example, let us consider groups \textit{P2} and \textit{P6}.
+As an example, let us consider groups \textit{P2} and \textit{P6} \textcolor{ForestGreen}{(see Figure 1C)}.
 If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained.
@@
 
-Here we show that beyond merely responding to the elementary symmetry operations of reflection \citep{Sasaki_2005, Tyler_2005} and rotation \citep{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every composition of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
+Here we show that beyond merely responding to the elementary symmetry operations of reflection \citep{Sasaki_2005, Tyler_2005} and rotation \citep{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every \textcolor{ForestGreen}{combination} of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
 Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \citep{RN1711}.
@@
 It has recently been demonstrated that a holographic model of regularity \citep{van_der_helm_1996}, can predict both SPN amplitude \citep{makin_2016} and perceptual discrimination performance \citep{nucci_2007} for dot patterns that contain symmetry and other types of regularity.
-The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods of isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure 2), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study that has been published on the wallpaper groups also produced SPN-like waveforms \citep{RN1959}.
-Future work should more firmly establish the connection and determine if the SPN can capture responses similarly precise symmetry responses as the SSVEPs presented here.
+The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods of isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure 2), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study that has been published on the wallpaper groups produced SPN-like waveforms \citep{RN1959}.
+Future work should more firmly establish the connection and determine if the SPN can capture similarly precise symmetry responses as the SSVEPs presented here.
 It would also be worthwhile to ask if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane.}
@@
 It may also be related to the fact that participants where not doing a symmetry-related task during the SSVEP experiment, but instead monitored the stimuli for brief changes in contrast that occured twice per trial (see Methods).
-Previous brain imaging studies have found enhanced reflection symmetry responses when participants performed a symmetry-related task \cite{makin_formation_2020, Sasaki_2005, keefe_2018}.
-We did not manipulate task during our SSVEP recordings, but it is possible that adding a symmetry-related task to our SSVEP experiment would have produced measurements that reflected subgroup relationships to an even higher extent.
-We note, however, that our performance is already close to ceiling (see Figure 5).
-It is possible that adding a symmetry-related task would enhance SSVEP amplitudes across all wallpaper groups without improving the discriminality of individual groups (similar to what was observed by Keefe and his co-authors \cite{keefe_2018}).
-SPN measurements suggest that task-driven processing may be important for detecting symmetries that have been subject to perspective distortion \cite{makin_conditions_2015} although it should be noted that this effect was much less clear in a subsequent functional MRI study \cite{keefe_2018}.
-Future work in which behavioral and brain imaging data are collected from the same participants, and behavior is manipulated in the SSVEP task, will help further establish the connection between the two measurements, and elucidate the potential contribution of task-related top-down processing to the current results.}
+Previous brain imaging studies have found enhanced reflection symmetry responses when participants performed a symmetry-related task \citep{makin_formation_2020, Sasaki_2005, keefe_2018}.
+It is possible that adding a symmetry-related task to our SSVEP experiment would have produced measurements that reflected subgroup relationships to an even higher extent than what we observed.
+On the other hand, our results are already close to ceiling (see Figure 5) and adding a symmetry-related task may simply enhance SSVEP amplitudes overall without improving the discriminality of individual groups, as has been observed for reflection by \cite{keefe_2018}.
+Task-driven processing may be important for detecting symmetries that have been subject to perspective distortion, as suggested by SPN measurements \citep{makin_conditions_2015} and somewhat less clearly in a subsequent functional MRI study \citep{keefe_2018}.
+Future work in which behavioral and brain imaging data are collected from the same participants, and task is manipulated in the SSVEP experiment, will help further establish the connection between the two measurements, and elucidate the potential contribution of task-related top-down processing to the current results.}
 
@@
 Field biologists have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \citep{RN1775}.
-In the norm-based encoding framework, wallpaper groups can be considered \textit{supertextures}, exaggerated examples of the near-regular textures that surround us.
-Artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \citep{RN1764}.
+In the norm-based encoding framework, wallpaper groups can be considered \textit{supertextures}, exaggerated examples of the near-regular textures \textcolor{ForestGreen}{common in the natural world.
+If non-human animals employ a similar encoding strategy, they would be expected to be sensitive to symmetries in wallpaper groups.
+Recent functional MRI work in macaque monkeys offer some support for that: Macaque visual cortex responds parametrically to reflection and rotation symmetries in wallpaper groups, and the set of brain areas involved largely overlap those observed to be sensitive to symmetry in humans \citep{audurier_symmetry_2021}.} In human societies, visual artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \citep{RN1764}.
 Wallpaper groups are visually compelling, and symmetries have been widely used in human artistic expression going back to the Neolithic age \citep{RN1949}.
@@
 To control vigilance, a contrast dimming task was employed.
-Two times per trial, an image pair \textcolor{ForestGreen}{(control {P1} plus exemplar)} was shown at reduced contrast.
+Two times per trial, an image pair \textcolor{ForestGreen}{(control \textit{P1} plus exemplar)} was shown at reduced contrast.
 \textcolor{ForestGreen}{P}articipants were instructed to press a button on a response pad \textcolor{ForestGreen}{whenever they noticed a contrast change.
@@
 \subsection*{Bayesian Analysis of SSVEP and Psychophysical data}
-
 Bayesian analysis was carried out using R (v3.6.1) \citep{R} with the \texttt{brms} package (v2.9.0) \citep{burkner2017} and rStan (v2.19.2 \citep{rStan}).

### 2021-05-18 Writeup/kohler_clarke_2021.tex
small edits

@@
 transformations that preserve distance \citep{RN1425}.
-The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation (CITATION?).
+The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation.
 In brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotation symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains reflection (m) or glide reflection (g).
-For example, $P4$ contains 4 fold rotation, while $P2M$ contains 2 fold rotation and one reflection (see Figure XXX).
+For example, $P4$ contains 4 fold rotation, while $P4MM$ contains 4 fold rotation and two mirror reflections.
+By convention, many of the groups are given shortened names: for example, $P4MM$ is usually referred to as $P4M$, as the second reflection can be deduced from the presence of a 4 fold rotation alongside a reflection.
+(see Figures \ref{fig:subgroups} and \ref{fig:example_stimuli} for examples).
 Two of the groups start with a $C$ rather than a $P$, ($CM$ and $CMM$) which indicates that the symmetries are specified relative to a cell that itself contains repetition.
@@
 Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \citep{RN1425}.
-The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
+\textcolor{ForestGreen}{A more complex example are the subgroups of \textit{PMM}, which is made up of two reflections, and a rotation of $180^{\circ}$.
+As such, $P2$ and $PM$ are both subgroups of $PMM$.
+(Illustrative examples are given in Figure \ref{fig:subgroups})} The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
 
@@
 We also show enlarged versions of some of the subgroup relationships involving \textit{PMM} (B, shown in blue) and \textit{P6} (C, shown in red) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
+\label{fig:subgroups}
 \end{figure}

### 2021-05-18 Writeup/kohler_clarke_2021.tex
final changes to manuscript, abstract, etc

--- 2021-05-14 Writeup/kohler_clarke_2021.tex
@@
 \begin{abstract}
-Symmetries are present at many scales in images of natural scenes.
-\textcolor{ForestGreen}{Humans and other animals are highly sensitive to visual symmetry, and a large} literature has demonstrated contributions of symmetry to numerous domains of visual perception.
-The four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined in exactly 17 distinct ways.
-These \textit{wallpaper groups} represent the complete set of symmetries in 2D images and have recently found use in the vision science community as an ideal stimulus set for studying the perception of symmetries in textures.
-The goal of the current study is to provide a more comprehensive description of responses to symmetry in the human visual system, by collecting both brain imaging (Steady-State Visual Evoked Potentials measured using high-density EEG) and behavioral (symmetry detection thresholds) data using the entire set of wallpaper groups.
+\textcolor{ForestGreen}{Symmetries are present at many scales in natural scenes.
+Humans and other animals are highly sensitive to visual symmetry, and symmetry contributes to numerous domains of visual perception.
+The four fundamental symmetries, reflection, rotation, translation and glide reflection, can be combined into exactly 17 distinct regular textures.
+These \textit{wallpaper groups} represent the complete set of symmetries in 2D images.
+The current study seeks to provide a more comprehensive description of responses to symmetry in the human visual system, by collecting both brain imaging (Steady-State Visual Evoked Potentials measured using high-density EEG) and behavioral (symmetry detection thresholds) data using the entire set of wallpaper groups.
 This allows us to probe the hierarchy of complexity among wallpaper groups, in which simpler groups are subgroups of more complex ones.
-We find that this hierarchy is preserved almost perfectly in both behavior and brain activity: A multi-level Bayesian GLM indicates that for most of the 63 subgroup relationships, subgroups produce lower amplitude responses in visual cortex (posterior probability: $>0.95$ for 56 of 63) and require longer presentation durations to be reliably detected (posterior probability: $>0.95$ for \textcolor{ForestGreen}{48} of 63).
-This systematic pattern is seen only in visual cortex and only in components of the brain response known to be symmetric-specific.
-Our results show that representations of symmetries in the human brain are precise and rich in detail, and that this precision is reflected in behavior.
-These findings expand our understanding of symmetry perception, and open up new avenues for research on how fine-grained representations of regular textures contribute to natural vision.
+We find that both behavior and brain activity preserve the hierarchy almost perfectly: Subgroups consistently produce lower amplitude symmetry-specific responses in visual cortex and require longer presentation durations to be reliably detected.
+These findings expand our understanding of symmetry perception by showing that the human brain encodes symmetries with a high level of precision and detail.
+This opens new avenues for research on how fine-grained representations of regular textures contribute to natural vision.}
 \end{abstract}
@@
 transformations that preserve distance \citep{RN1425}.
-The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation (CITATION?).
+The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation.
 In brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotation symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains reflection (m) or glide reflection (g).
-For example, $P4$ contains 4 fold rotation, while $P2M$ contains 2 fold rotation and one reflection (see Figure XXX).
+For example, $P4$ contains rotation of order 4, while $P4MM$ contains rotation of order 4 and two reflections.
+By convention, many of the groups are given shortened names: for example, $P4MM$ is usually referred to as $P4M$, as the second reflection can be deduced from the presence of rotation of order 4 alongside a reflection.
 Two of the groups start with a $C$ rather than a $P$, ($CM$ and $CMM$) which indicates that the symmetries are specified relative to a cell that itself contains repetition.
-Full details of the naming convention and examples of the wallpaper groups can be found on \href{https://en.wikipedia.org/wiki/Wallpaper_group}{wikipedia}}.
+Full details of the naming convention can be found on \href{https://en.wikipedia.org/wiki/Wallpaper_group}{wikipedia} and examples of the wallpaper groups are shown in Figures \ref{fig:subgroups} and \ref{fig:summary}.}
 
@@
 the number of times the subgroup is found in the supergroup \citep{RN1425}.
-As an example, let us consider groups \textit{P2} and \textit{P6} \textcolor{ForestGreen}{(see Figure 1C)}.
+As an example, let us consider groups \textit{P2} and \textit{P6} \textcolor{ForestGreen}{(see Figure 1B)}.
 If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained.
@@
 Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \citep{RN1425}.
-The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
+\textcolor{ForestGreen}{Similarly, \textit{PMM} contains two reflections and rotations \{$0^{\circ}$, $180^{\circ}$\}.
+\textit{PMM} can be generated by adding an additional reflection to both \textit{P2} (\{$0^{\circ}$, $180^{\circ}$\}) and \textit{PM} (one reflection), so \textit{P2} and \textit{PM} are both subgroups of \textit{PMM} with index 2 (see Figure 1C).} The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
 
@@
 \caption{\textcolor{ForestGreen}{Subgroup relationships with indices 2 (solid lines) and 3 (dashed line) are shown in (A).
-All other relationships can be infered by identifying the shortest path through the hierarchy, and multiplying the subgroup indices.
+All other relationships can be inferred by identifying the shortest path through the hierarchy, and multiplying the subgroup indices.
 For example, \textit{P1} is related to \textit{P6} through \textit{P6}\textrightarrow\textit{P3} (index 2) and \textit{P3}\textrightarrow\textit{P1} (index 3) so \textit{P1} is also a subgroup of \textit{P6} with index $3 \times 2 = 6$.
-We also show enlarged versions of some of the subgroup relationships involving \textit{PMM} (B, shown in blue) and \textit{P6} (C, shown in red) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
+We also show enlarged versions of some of the subgroup relationships involving \textit{P6} (B, shown in red) and \textit{PMM} (C, shown in blue) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
+\label{fig:subgroups}
 \end{figure}
@@
 The 16 groups are ordered by the strength of the SSVEP response, to highlight the range of response amplitudes.}
-\label{fig:example_stimuli}
+\label{fig:summary}
 \end{figure}
@@
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that 55/63 subgroup relationships have $p(\Delta|data)>0.99$.}
+We can see that 55/63 subgroup relationships have $p(\Delta <0 | data)>0.99$.}
 \label{fig:eeg_comp}
@@
 \caption{Posterior distributions for the difference in mean symmetry detection threshold durations.
-Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that 43/63 subgroup relationships have $p(\Delta|data)>0.99$.}
+Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is smaller than zero.
+We can see that 43/63 subgroup relationships have $p(\Delta >0 | data)>0.99$.}
 \label{fig:durations_comp}
@@
 It has recently been demonstrated that a holographic model of regularity \citep{van_der_helm_1996}, can predict both SPN amplitude \citep{makin_2016} and perceptual discrimination performance \citep{nucci_2007} for dot patterns that contain symmetry and other types of regularity.
-The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods of isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure 2), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study that has been published on the wallpaper groups produced SPN-like waveforms \citep{RN1959}.
+The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods for isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure 2), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study on the wallpaper groups also found SPN-like waveforms \citep{RN1959}.
 Future work should more firmly establish the connection and determine if the SPN can capture similarly precise symmetry responses as the SSVEPs presented here.
@@
 \includegraphics[width=0.5\linewidth]{../analysis/plots/model_roc_style.pdf}
-\caption{This plot shows the proportion of subgroup relationships that satisfy $p(\Delta >0 | data) > x$.
+\caption{This plot shows the proportion of subgroup relationships that satisfy \textcolor{ForestGreen}{$p(\Delta >0 | data) > x$ for the SSVEP data and $p(\Delta <0 | data) > x$ for the psychophysical data}.
 We can see that if we take $x = 0.95$ as our threshold, the subgroup relationships are preserved in 56/63 = 89\% and \textcolor{ForestGreen}{48/63 = 76\%} of the comparisons for the primary SSVEP and threshold duration datasets, receptively.

### 2021-05-18 Writeup/kohler_clarke_2021.tex
minor additional changes

@@
 the number of times the subgroup is found in the supergroup \citep{RN1425}.
-As an example, let us consider groups \textit{P2} and \textit{P6} \textcolor{ForestGreen}{(see Figure 1B)}.
+As an example, let us consider groups \textit{P2} and \textit{P6} \textcolor{ForestGreen}{(see Figure \ref{fig:subgroups}B)}.
 If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained.
@@
 \textcolor{ForestGreen}{Similarly, \textit{PMM} contains two reflections and rotations \{$0^{\circ}$, $180^{\circ}$\}.
-\textit{PMM} can be generated by adding an additional reflection to both \textit{P2} (\{$0^{\circ}$, $180^{\circ}$\}) and \textit{PM} (one reflection), so \textit{P2} and \textit{PM} are both subgroups of \textit{PMM} with index 2 (see Figure 1C).} The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
+\textit{PMM} can be generated by adding an additional reflection to both \textit{P2} (\{$0^{\circ}$, $180^{\circ}$\}) and \textit{PM} (one reflection), so \textit{P2} and \textit{PM} are both subgroups of \textit{PMM} with index 2 (see Figure \ref{fig:subgroups}C).} The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
 
@@
 
-Examples of the wallpaper groups and a summary of our brain imaging and psychophysical measurements are shown in Figure 2.
+Examples of the wallpaper groups and a summary of our brain imaging and psychophysical measurements are shown in Figure \ref{fig:summary}.
 For our primary SSVEP analysis, we only considered EEG data from a pre-determined region-of-interest (ROI) consisting of six electrodes over occipital cortex (see Supplementary Figure 1.1).
@@
 The data in Figure 2 are shown in descending order according to RMS.
-The psychophysical results, shown in box plots in Figure 2, were also quite variable between groups, and there seems to be a general pattern where wallpaper groups near the top of the figure, that have lower SSVEP amplitudes, also have longer psychophysical threshold durations.
+The psychophysical results, shown in box plots in Figure \ref{fig:summary}, were also quite variable between groups, and there seems to be a general pattern where wallpaper groups near the top of the figure, that have lower SSVEP amplitudes, also have longer psychophysical threshold durations.
 
@@
 These difference distributions allowed us to compute the conditional probability that the supergroup would produce (a) larger RMS and (b) a shorter threshold durations, when compared to the subgroup.
-The posterior distributions are shown in Figure 3 for the SSVEP data, and in Figure 4 for the psychophysical data, which distributions color-coded according to conditional probability.
+The posterior distributions are shown in Figure \ref{fig:eeg_comp} for the SSVEP data, and in Figure \ref{fig:durations_comp} for the psychophysical data, which distributions color-coded according to conditional probability.
 For both data sets our hypothesis is confirmed: For the overwhelming majority of the 63 subgroup relationships, supergroups are more likely to produce larger symmetry-specific SSVEPs and shorter symmetry detection threshold durations, and in most cases the conditional probability of this happening is extremely high.
@@
 We can quantify the strength of the association between the data and the subgroup relationships, by asking what proportion of subgroup relationships that reach or exceed a range of probability thresholds.
-This is plotted in Figure 5, for our psychophysical data, our primary SSVEP analysis and our two control SSVEP analyses.
+This is plotted in Figure \ref{fig:roc_style}, for our psychophysical data, our primary SSVEP analysis and our two control SSVEP analyses.
 It shows that odd-harmonic SSVEP data from the occipital ROI and symmetry detection threshold durations both have a strong association with the subgroup relationships such that a clear majority of the subgroups survive even at the highest threshold we consider ($p(\Delta>0|data)>0.99$).
@@
 Our data} provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane.
-\textcolor{ForestGreen}{We do not} independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure 2), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
+\textcolor{ForestGreen}{We do not} independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure \ref{fig:summary}), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
 The subgroup relationships are in many cases not obvious perceptually, and most participants had no knowledge of group theory.
@@
 It has recently been demonstrated that a holographic model of regularity \citep{van_der_helm_1996}, can predict both SPN amplitude \citep{makin_2016} and perceptual discrimination performance \citep{nucci_2007} for dot patterns that contain symmetry and other types of regularity.
-The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods for isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure 2), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study on the wallpaper groups also found SPN-like waveforms \citep{RN1959}.
+The available evidence suggests that the SPN and our SSVEP measurements are two distinct methods for isolating the same symmetry-related brain response: When observed in the time-domain, the symmetry-selective odd-harmonic responses produce similarly sustained waveforms (see Figure \ref{fig:summary}), odd-harmonic SSVEP responses can be measured for dot patterns similar to those used to measure the SPN \citep{RN1954}, and the one event-related study on the wallpaper groups also found SPN-like waveforms \citep{RN1959}.
 Future work should more firmly establish the connection and determine if the SPN can capture similarly precise symmetry responses as the SSVEPs presented here.
@@
 It is possible that adding a symmetry-related task to our SSVEP experiment would have produced measurements that reflected subgroup relationships to an even higher extent than what we observed.
-On the other hand, our results are already close to ceiling (see Figure 5) and adding a symmetry-related task may simply enhance SSVEP amplitudes overall without improving the discriminality of individual groups, as has been observed for reflection by \cite{keefe_2018}.
+On the other hand, our results are already close to ceiling (see Figure \ref{fig:roc_style}) and adding a symmetry-related task may simply enhance SSVEP amplitudes overall without improving the discriminality of individual groups, as has been observed for reflection by \cite{keefe_2018}.
 Task-driven processing may be important for detecting symmetries that have been subject to perspective distortion, as suggested by SPN measurements \citep{makin_conditions_2015} and somewhat less clearly in a subsequent functional MRI study \citep{keefe_2018}.
@@
 This is perhaps because the index analysis does not take into account the \textit{type} of isometries that differentiate the subgroup and supergroup.
-The effect of symmetry type can be observed by contrasting the measured SSVEP amplitudes and detection thresholds for groups \textit{PM} and \textit{PG} in Figure 2.
+The effect of symmetry type can be observed by contrasting the measured SSVEP amplitudes and detection thresholds for groups \textit{PM} and \textit{PG} in Figure \ref{fig:summary}.
 The two groups are comparable except \textit{PM} contains reflection and \textit{PG} contains glide reflection, and the former clearly elicits higher amplitudes and lower thresholds.
@@
 This compares to the \textcolor{ForestGreen}{32/63= 51\%} and \textcolor{ForestGreen}{22/63 = 35\%} for the SSVEP control datasets.}
-\label{fig:durations_rotations}
+\label{fig:roc_style}
 \end{figure}
@@
 For our main analysis, we averaged the odd harmonic single-cycle waveforms within a six-electrode region of interest (ROI) over occipital cortex (electrodes 70, 74, 75, 81, 82, 83).
-These waveforms, averaged over participants, are shown in Figure 2.
+These waveforms, averaged over participants, are shown in Figure \ref{fig:summary}.
 The same analysis was done for the even harmonics and for the odd harmonics within a six electrode ROI over parietal cortex (electrodes 53, 54, 61, 78, 79, 86; see Supplementary Figure 1.1).

### 2021-06-23 Writeup/kohler_clarke_2021.tex
minor revisions, no longer tracked changes

@@
 \begin{abstract}
-\textcolor{ForestGreen}{Symmetries are present at many scales in natural scenes.
+Symmetries are present at many scales in natural scenes.
 Humans and other animals are highly sensitive to visual symmetry, and symmetry contributes to numerous domains of visual perception.
@@
 These findings expand our understanding of symmetry perception by showing that the human brain encodes symmetries with a high level of precision and detail.
-This opens new avenues for research on how fine-grained representations of regular textures contribute to natural vision.}
+This opens new avenues for research on how fine-grained representations of regular textures contribute to natural vision.
 \end{abstract}
@@
 Symmetries are abundant in natural and man-made environments, due to a complex interplay of physical forces that govern pattern formation in nature.
-\textcolor{ForestGreen}{Sensitivity to symmetry has been demonstrated in a number of species, includes bees \citep{giurfa_insect_symmetry_1996}, fish \citep{morris_swordtail_symmetry_1998,schluter_sailfin_symmetry_1998}, birds \citep{RN1337,swaddle_symmetry_finches_1994} and dolphins \citep{von_fersen_dolphin_symmetry_1992}, and may be used as a cue for mate selection in many species \citep{swaddle_asymmetry_review_1999} including humans \citep{rhodes_facial_1998}.
-Humans cultures have created and appreciated s}ymmetrical patterns throughout history, and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception.
+Sensitivity to symmetry has been demonstrated in a number of species, includes bees \citep{giurfa_insect_symmetry_1996}, fish \citep{morris_swordtail_symmetry_1998,schluter_sailfin_symmetry_1998}, birds \citep{RN1337,swaddle_symmetry_finches_1994} and dolphins \citep{von_fersen_dolphin_symmetry_1992}, and may be used as a cue for mate selection in many species \citep{swaddle_asymmetry_review_1999} including humans \citep{rhodes_facial_1998}.
+Humans cultures have created and appreciated symmetrical patterns throughout history, and since the gestalt movement of the early 20th century, symmetry has been recognized as important for visual perception.
 Symmetry contributes to the perception of shapes \citep{RN1311,RN1682}, scenes \citep{apthorp_symmetry_numerosity} and surface properties \citep{RN1166}.
-\textcolor{ForestGreen}{This literature is almost exclusively based on stimuli in which one or more symmetry axes are placed at a single point in the image.
-Focus has been on} mirror symmetry or \textit{reflection}, \textcolor{ForestGreen}{with relatively few studies} including the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection} \citep{wagemans_1998} \textcolor{ForestGreen}{- perhaps because reflection has been found to be more perceptually salient \citep{mach_1959,royer_1981,palmer_1991,ogden_2016,hamada_1988} and produce more brain activity \citep{makin_2013,makin_2014,makin_2012,wright_2015}.
-In the current study, we take a different approach by investigating visual processing of regular textures in which combinations of the four fundamental symmetries tile the 2D plane.}
+This literature is almost exclusively based on stimuli in which one or more symmetry axes are placed at a single point in the image.
+Focus has been on mirror symmetry or \textit{reflection}, with relatively few studies including the other fundamental symmetries: \textit{rotation}, \textit{translation} and \textit{glide reflection} \citep{wagemans_1998} - perhaps because reflection has been found to be more perceptually salient \citep{mach_1959,royer_1981,palmer_1991,ogden_2016,hamada_1988} and produce more brain activity \citep{makin_2013,makin_2014,makin_2012,wright_2015}.
+In the current study, we take a different approach by investigating visual processing of regular textures in which combinations of the four fundamental symmetries tile the 2D plane.
 
@@
 Previous work on a subset of four of the wallpaper groups used functional MRI to demonstrate that rotation symmetries in wallpapers elicit parametric responses in several areas in occipital cortex, beginning with visual area V3 \citep{RN1725}.
-This effect was also robust \textcolor{ForestGreen}{when symmetry responses were measured with electroencephalography (EEG) using both Steady-State Visual Evoked Potentials (SSVEPs)\citep{RN1725} and Event-Related Potentials \citep{RN1959}.
+This effect was also robust when symmetry responses were measured with electroencephalography (EEG) using both Steady-State Visual Evoked Potentials (SSVEPs)\citep{RN1725} and Event-Related Potentials \citep{RN1959}.
 The SSVEP technique uses periodic visual stimulation to produce a periodic brain response that is confined to integer multiples of the stimulation frequency known as harmonics.
 SSVEP response harmonics can be isolated in the frequency domain and depending on the specific design, different harmonics will express different aspects of the brain response.
-\citep{RN1540}.} Here we extend on the previous work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups.
+\citep{RN1540}.
+Here we extend on the previous work by collecting SSVEPs and psychophysical data from human participants viewing the full set of wallpaper groups.
 We measure responses in visual cortex to 16 out of the 17 wallpaper groups, with the 17th serving as a control stimulus.
@@
 transformations that preserve distance \citep{RN1425}.
-The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation.
+The wallpaper groups differ in the number and kind of these transformations and we can uniquely refer to different groups using crystallographic notation.
 In brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotation symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains reflection (m) or glide reflection (g).
@@
 Two of the groups start with a $C$ rather than a $P$, ($CM$ and $CMM$) which indicates that the symmetries are specified relative to a cell that itself contains repetition.
-Full details of the naming convention can be found on \href{https://en.wikipedia.org/wiki/Wallpaper_group}{wikipedia} and examples of the wallpaper groups are shown in Figures \ref{fig:subgroups} and \ref{fig:summary}.}
+Full details of the naming convention can be found on \href{https://en.wikipedia.org/wiki/Wallpaper_group}{wikipedia} and examples of the wallpaper groups are shown in Figures \ref{fig:subgroups} and \ref{fig:summary}.
 
@@
 the number of times the subgroup is found in the supergroup \citep{RN1425}.
-As an example, let us consider groups \textit{P2} and \textit{P6} \textcolor{ForestGreen}{(see Figure \ref{fig:subgroups}B)}.
+As an example, let us consider groups \textit{P2} and \textit{P6} (see Figure \ref{fig:subgroups}B).
 If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained.
@@
 Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \citep{RN1425}.
-\textcolor{ForestGreen}{Similarly, \textit{PMM} contains two reflections and rotations \{$0^{\circ}$, $180^{\circ}$\}.
-\textit{PMM} can be generated by adding an additional reflection to both \textit{P2} (\{$0^{\circ}$, $180^{\circ}$\}) and \textit{PM} (one reflection), so \textit{P2} and \textit{PM} are both subgroups of \textit{PMM} with index 2 (see Figure \ref{fig:subgroups}C).} The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
+Similarly, \textit{PMM} contains two reflections and rotations \{$0^{\circ}$, $180^{\circ}$\}.
+\textit{PMM} can be generated by adding an additional reflection to both \textit{P2} (\{$0^{\circ}$, $180^{\circ}$\}) and \textit{PM} (one reflection), so \textit{P2} and \textit{PM} are both subgroups of \textit{PMM} with index 2 (see Figure \ref{fig:subgroups}C).
+The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.
 
@@
 \includegraphics[width=0.75\linewidth]{../figures/figure1.pdf}
-\caption{\textcolor{ForestGreen}{Subgroup relationships with indices 2 (solid lines) and 3 (dashed line) are shown in (A).
+\caption{Subgroup relationships with indices 2 (solid lines) and 3 (dashed line) are shown in (A).
 All other relationships can be inferred by identifying the shortest path through the hierarchy, and multiplying the subgroup indices.
 For example, \textit{P1} is related to \textit{P6} through \textit{P6}\textrightarrow\textit{P3} (index 2) and \textit{P3}\textrightarrow\textit{P1} (index 3) so \textit{P1} is also a subgroup of \textit{P6} with index $3 \times 2 = 6$.
-We also show enlarged versions of some of the subgroup relationships involving \textit{P6} (B, shown in red) and \textit{PMM} (C, shown in blue) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
+We also show enlarged versions of some of the subgroup relationships involving \textit{P6} (B, shown in red) and \textit{PMM} (C, shown in blue) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}
 \label{fig:subgroups}
@@
 
-Here we show that beyond merely responding to the elementary symmetry operations of reflection \citep{Sasaki_2005, Tyler_2005} and rotation \citep{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every \textcolor{ForestGreen}{combination} of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
+Here we show that beyond merely responding to the elementary symmetry operations of reflection \citep{Sasaki_2005, Tyler_2005} and rotation \citep{RN1725}, the visual system represents the hierarchical structure of the 17 wallpaper groups, and thus every combination of the four fundamental symmetries (rotation, reflection, translation, glide reflection) which comprise the set of regular textures.
 Both SSVEP amplitudes and symmetry detection thresholds preserve the hierarchy of complexity among the wallpaper groups that is captured by the subgroup relationships \citep{RN1711}.
@@
 
-\textcolor{ForestGreen}{The current study uses 16 distinct wallpaper groups, while previous neuroimaging studies focused on a subset of 4 \citep{RN1725,RN1959}.
+The current study uses 16 distinct wallpaper groups, while previous neuroimaging studies focused on a subset of 4 \citep{RN1725,RN1959}.
 This represents a significant conceptual advance, because it makes it possible to investigate the complete subgroup hierarchy among the 17 groups and ask to what extent the hierarchy is reflected in brain activity.
-Our data} provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane.
-\textcolor{ForestGreen}{We do not} independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure \ref{fig:summary}), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
+Our data provide a description of the visual system’s response to the complete set of symmetries in the two-dimensional plane.
+We do not independently measure the response to \textit{P1}, but because each of the 16 other groups produce non-zero odd harmonic amplitudes (see Figure \ref{fig:summary}), we can conclude that the relationships between P1 and all other groups, where \textit{P1} is the subgroup, are also preserved by the visual system.
 The subgroup relationships are in many cases not obvious perceptually, and most participants had no knowledge of group theory.
 Thus, the visual system’s ability to preserve the subgroup hierarchy does not depend on explicit knowledge of the relationships.
-\textcolor{ForestGreen}{Previous brain-imaging studies have found evidence of parametric responses with the number of reflection symmetry folds \cite{keefe_2018, Sasaki_2005, makin_2016} and with the order of rotation symmetry \cite{RN1725}.
-Our study is the first demonstration that the brain encodes symmetry in this parametric fashion across every possible combination of different \textit{symmetry types}, and that this parametric encoding is also reflected in behavior.} Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \citep{RN1253}, they tend to sort exemplars into fewer \textcolor{ForestGreen}{(4-12) sets than the number of wallpaper groups, often placing exemplars from different wallpaper groups in the same set \citep{RN172}.
-The two-interval forced choice approach we use in the current psychophysical experiment makes it possible to directly compare symmetry detection thresholds to the subgroup hierarchy, and reveals that not only can the 17 wallpaper groups be distinguished based on behavioral data, behavior largely follows the subgroup hierarchy.}
+Previous brain-imaging studies have found evidence of parametric responses with the number of reflection symmetry folds \cite{keefe_2018, Sasaki_2005, makin_2016} and with the order of rotation symmetry \cite{RN1725}.
+Our study is the first demonstration that the brain encodes symmetry in this parametric fashion across every possible combination of different \textit{symmetry types}, and that this parametric encoding is also reflected in behavior.
+Previous behavioral experiments have shown that although naïve observers can distinguish many of the wallpaper groups \citep{RN1253}, they tend to sort exemplars into fewer (4-12) sets than the number of wallpaper groups, often placing exemplars from different wallpaper groups in the same set \citep{RN172}.
+The two-interval forced choice approach we use in the current psychophysical experiment makes it possible to directly compare symmetry detection thresholds to the subgroup hierarchy, and reveals that not only can the 17 wallpaper groups be distinguished based on behavioral data, behavior largely follows the subgroup hierarchy.
 
-\textcolor{ForestGreen}{A large literature exists on the \textit{Sustained Posterior Negativity} (SPN), a characteristic negative-going waveform that is known to reflect responses to symmetry and other forms of regularity and structure \citep{makin_2016}.
+A large literature exists on the \textit{Sustained Posterior Negativity} (SPN), a characteristic negative-going waveform that is known to reflect responses to symmetry and other forms of regularity and structure \citep{makin_2016}.
 The SPN scales with the proportion of reflection symmetry in displays that contain a mixture of symmetry and noise \cite{makin_formation_2020, palumbo_scaling_2015}, and both reflection, rotation and translation can produce a measurable SPN \cite{makin_2013}.
@@
 Future work should more firmly establish the connection and determine if the SPN can capture similarly precise symmetry responses as the SSVEPs presented here.
-It would also be worthwhile to ask if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane.}
+It would also be worthwhile to ask if and how \textit{W} can computed for our random-noise based wallpaper textures where combinations of symmetries tile the plane.
 
@@
 This suggests that the two measurements reflect the same underlying symmetry representations in visual cortex.
-\textcolor{ForestGreen}{It should be noted that the correlation is relatively modest ($R^2 = 0.44$).
+It should be noted that the correlation is relatively modest ($R^2 = 0.44$).
 This may be partly due to the fact that different individuals participated in the two experiments.
@@
 Task-driven processing may be important for detecting symmetries that have been subject to perspective distortion, as suggested by SPN measurements \citep{makin_conditions_2015} and somewhat less clearly in a subsequent functional MRI study \citep{keefe_2018}.
-Future work in which behavioral and brain imaging data are collected from the same participants, and task is manipulated in the SSVEP experiment, will help further establish the connection between the two measurements, and elucidate the potential contribution of task-related top-down processing to the current results.}
+Future work in which behavioral and brain imaging data are collected from the same participants, and task is manipulated in the SSVEP experiment, will help further establish the connection between the two measurements, and elucidate the potential contribution of task-related top-down processing to the current results.
 
@@
 
-The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective \textcolor{ForestGreen}{(see above)} and occlusion.
-Near-regular textures, however, abound in the visual world, \textcolor{ForestGreen}{and can be modeled} as deformed versions of the wallpaper groups \citep{RN1519}.
+The 17 wallpaper groups are completely regular, and relatively rare in the visual environment, especially when considering distortions due to perspective (see above) and occlusion.
+Near-regular textures, however, abound in the visual world, and can be modeled as deformed versions of the wallpaper groups \citep{RN1519}.
 The correspondence between visual cortex responses and group theory demonstrated here may indicate that the visual system represents visual textures using a similar scheme, with the wallpaper groups serving as anchor points in representational space.
@@
 Field biologists have demonstrated that animals respond more strongly to exaggerated versions of a learned stimulus, referred to as “supernormal” stimuli \citep{RN1775}.
-In the norm-based encoding framework, wallpaper groups can be considered \textit{supertextures}, exaggerated examples of the near-regular textures \textcolor{ForestGreen}{common in the natural world.
+In the norm-based encoding framework, wallpaper groups can be considered \textit{supertextures}, exaggerated examples of the near-regular textures common in the natural world.
 If non-human animals employ a similar encoding strategy, they would be expected to be sensitive to symmetries in wallpaper groups.
-Recent functional MRI work in macaque monkeys offer some support for that: Macaque visual cortex responds parametrically to reflection and rotation symmetries in wallpaper groups, and the set of brain areas involved largely overlap those observed to be sensitive to symmetry in humans \citep{audurier_symmetry_2021}.} In human societies, visual artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \citep{RN1764}.
+Recent functional MRI work in macaque monkeys offer some support for that: Macaque visual cortex responds parametrically to reflection and rotation symmetries in wallpaper groups, and the set of brain areas involved largely overlap those observed to be sensitive to symmetry in humans \citep{audurier_symmetry_2021}.
+In human societies, visual artists may consciously or unconsciously create supernormal stimuli, to capture the essence of the subject and evoke strong responses in the audience \citep{RN1764}.
 Wallpaper groups are visually compelling, and symmetries have been widely used in human artistic expression going back to the Neolithic age \citep{RN1949}.
@@
 \includegraphics[width=0.5\linewidth]{../analysis/plots/model_roc_style.pdf}
-\caption{This plot shows the proportion of subgroup relationships that satisfy \textcolor{ForestGreen}{$p(\Delta >0 | data) > x$ for the SSVEP data and $p(\Delta <0 | data) > x$ for the psychophysical data}.
-We can see that if we take $x = 0.95$ as our threshold, the subgroup relationships are preserved in 56/63 = 89\% and \textcolor{ForestGreen}{48/63 = 76\%} of the comparisons for the primary SSVEP and threshold duration datasets, receptively.
-This compares to the \textcolor{ForestGreen}{32/63= 51\%} and \textcolor{ForestGreen}{22/63 = 35\%} for the SSVEP control datasets.}
+\caption{This plot shows the proportion of subgroup relationships that satisfy $p(\Delta >0 | data) > x$ for the SSVEP data and $p(\Delta <0 | data) > x$ for the psychophysical data.
+We can see that if we take $x = 0.95$ as our threshold, the subgroup relationships are preserved in 56/63 = 89\% and 48/63 = 76\% of the comparisons for the primary SSVEP and threshold duration datasets, receptively.
+This compares to the 32/63= 51\% and 22/63 = 35\% for the SSVEP control datasets.}
 \label{fig:roc_style}
@@
 Their informed consent was obtained before the experiment under a protocol that was approved by the University of Essex's Ethics Committee.
-\textcolor{ForestGreen}{There was no overlap in participants between the EEG and psychophysics experiments.}
+There was no overlap in participants between the EEG and psychophysics experiments.
 
@@
 To control vigilance, a contrast dimming task was employed.
-Two times per trial, an image pair \textcolor{ForestGreen}{(control \textit{P1} plus exemplar)} was shown at reduced contrast.
-\textcolor{ForestGreen}{P}articipants were instructed to press a button on a response pad \textcolor{ForestGreen}{whenever they noticed a contrast change.
-Reaction times were not taken into account and participants were told to respond at their own pace while being as accurate as possible}.
-We adjusted the \textcolor{ForestGreen}{reduction in contrast} such that average accuracy for each participant was kept at ~85\% correct, in order to keep the difficulty of the vigilance \textcolor{ForestGreen}{task} at a constant level.
+Two times per trial, an image pair (control \textit{P1} plus exemplar) was shown at reduced contrast.
+Participants were instructed to press a button on a response pad whenever they noticed a contrast change.
+Reaction times were not taken into account and participants were told to respond at their own pace while being as accurate as possible.
+We adjusted the reduction in contrast such that average accuracy for each participant was kept at ~85\% correct, in order to keep the difficulty of the vigilance task at a constant level.
 
@@
 These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | \text{data})$.
-For further technical details, please see \href{https://osf.io/f3ex8/}{the Supplementary Materials} where the full R code, model specification, prior and posterior predictive checks, and model diagnostics, can be found.
+For further technical details, please see the Supplementary Materials where the full R code, model specification, prior and posterior predictive checks, and model diagnostics, can be found.
+
+\subsection*{Acknowledgments}
+The authors would like to thank two anonymous reviewers whose comments helped improve and clarify the manuscript, and also express their gratitude to Professor Anthony M.
+Norcia for his invaluable mentorship and contribution to our thinking about the role of symmetry in vision, and about vision and the brain more generally, through years of collaboration and discussions.
+This work was supported by the Vision Science to Applications (VISTA) program funded by the Canada First Research Excellence Fund (CFREF, 2016–2023) and by a Discovery Grant from the Natural Sciences and Engineering Research Council of Canada awarded to PJK.
+The work was also partially supported by a National Science Foundation INSPIRE grant 1248076 awarded to Yanxi Liu, Anthony M.
+Norcia and Rick O.
+Gilmore.
+A preliminary version of the EEG portion of the manuscript was previously deposited on \href{https://doi.org/10.1101/223743}{bioRxiv}.
+
+\subsection*{Data Accessibility}
+Data from the EEG and Psychophysics experiments have been made available with the Supplementary Material on \href{https://osf.io/f3ex8/}{OSF}.
 

### 2021-06-23 Writeup/kohler_clarke_2021.tex
credit David Wade

@@
 For example, \textit{P1} is related to \textit{P6} through \textit{P6}\textrightarrow\textit{P3} (index 2) and \textit{P3}\textrightarrow\textit{P1} (index 3) so \textit{P1} is also a subgroup of \textit{P6} with index $3 \times 2 = 6$.
-We also show enlarged versions of some of the subgroup relationships involving \textit{P6} (B, shown in red) and \textit{PMM} (C, shown in blue) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}
+We also show enlarged versions of some of the subgroup relationships involving \textit{P6} (B, shown in red) and \textit{PMM} (C, shown in blue) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.
+Illustration adapted from \citet{wade_crystal_1993}.}
 \label{fig:subgroups}

### 2021-06-24 analysis/analysis.Rmd
Updated header for supp mat to include title and DOI

@@
 ---
-title: "Supplementary Materials"
+title: "Supplementary Materials: The human visual system preserves the hierarchy of 2-dimensional pattern regularity"
 author: "Peter J. Kohler and Alasdair D. F. Clarke"
@@
 ---
+
+These are the supplementary materials for:
+
+Kohler, P.
+J., & Clarke, A.
+D.
+(2021).
+The human visual system preserves the hierarchy of 2-dimensional pattern regularity.
+Proceedings of the Royal Society B: Biological Sciences.
+10.1098/rspb.2021.1142
 

### 2021-06-28 Writeup/final version for upload/kohler_clarke_2021.tex
version for upload created

--- 2021-06-23 Writeup/kohler_clarke_2021.tex
@@
 \documentclass[11pt, twoside]{article}
-\usepackage{boisik}
-\usepackage[OT1]{fontenc}
+\usepackage{lmodern}
+\usepackage[T1]{fontenc}
 \usepackage[affil-it]{authblk}
@@
 \centering
-\includegraphics[width=0.75\linewidth]{../figures/figure1.pdf}
+\includegraphics[width=0.75\linewidth]{figure1.pdf}
 \caption{Subgroup relationships with indices 2 (solid lines) and 3 (dashed line) are shown in (A).
@@
 \centering
-\includegraphics[width=0.75\linewidth]{../figures/figure2.png}
+\includegraphics[width=0.75\linewidth]{figure2.pdf}
 \caption{Examples of each of the 16 wallpaper groups are shown in the left- and right-most column of the figures, next to the corresponding SSVEP (center-left) and psychological (center-right) data from each group.
@@
 
-The two datasets we present here make it possible to assess the extent to which both behavior and brain responses follow the hierarchy of complexity expressed by the subgroup relationships.
+The two datasets we present here (data and analysis code has been made available on \href{https://osf.io/f3ex8/}{OSF}) make it possible to assess the extent to which both behavior and brain responses follow the hierarchy of complexity expressed by the subgroup relationships.
 Based on previous brain imaging work showing that patterns with more axes of symmetry produce greater activity in visual cortex \citep{Sasaki_2005, Tyler_2005,RN1959, RN1725, keefe_2018}, we hypothesized that more complex groups would produce larger SSVEPs.
@@
 \centering
-\includegraphics[width=0.75\linewidth]{../analysis/plots/subgroup_comp_eeg_rms.png}
+\includegraphics[width=0.75\linewidth]{subgroup_comp_eeg_rms.png}
 \caption{Posterior distributions for the difference in mean SSVEP RMS amplitude.
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is greater than zero.
-We can see that 55/63 subgroup relationships have $p(\Delta <0 | data)>0.99$.}
+We can see that 55/63 subgroup relationships have $p(\Delta >0 | data)>0.99$.}
 \label{fig:eeg_comp}
@@
 \centering
-\includegraphics[width=0.75\linewidth]{../analysis/plots/subgroup_comp_psychophysical.png}
+\includegraphics[width=0.75\linewidth]{subgroup_comp_psychophysical.png}
 \caption{Posterior distributions for the difference in mean symmetry detection threshold durations.
 Colour coding of the text indicates the index of the subgroup, while the colour of the filled distribution relates to the conditional probability that the difference in means is smaller than zero.
-We can see that 43/63 subgroup relationships have $p(\Delta >0 | data)>0.99$.}
+We can see that 43/63 subgroup relationships have $p(\Delta <0 | data)>0.99$.}
 \label{fig:durations_comp}
@@
 \centering
-\includegraphics[width=0.5\linewidth]{../analysis/plots/model_roc_style.pdf}
+\includegraphics[width=0.5\linewidth]{model_roc_style.pdf}
 \caption{This plot shows the proportion of subgroup relationships that satisfy $p(\Delta >0 | data) > x$ for the SSVEP data and $p(\Delta <0 | data) > x$ for the psychophysical data.
@@
 These samples were then recombined to calculate the distribution of differences for each of the 63 pairs of subgroup and supergroup.
-These distributions were then summarised by computing the conditional probability of obtaining a positive (negative) difference, $p(\Delta | \text{data})$.
+These distributions were then summarised by computing the conditional probability of obtaining a positive (or for the psychophysical data, negative) difference, $p(\Delta | \text{data})$.
 For further technical details, please see the Supplementary Materials where the full R code, model specification, prior and posterior predictive checks, and model diagnostics, can be found.
@@
 \bibliographystyle{apalike} 
-\bibliography{literature}
+\bibliography{kohler_clarke_2021.bbl}
 \end{multicols}

### 2021-06-29 analysis/analysis.Rmd
png to pdf for some figures

@@
 both_plts
-ggsave("plots/subgroup_comp_eeg_rms.png", width = 5.5, height = 8)
+ggsave("plots/subgroup_comp_eeg_rms.pdf", width = 5.5, height = 8)
 ```
@@
 both_plts
-ggsave("plots/subgroup_comp_psychophysical.png",  width = 5.5, height = 8)
+ggsave("plots/subgroup_comp_psychophysical.pdf",  width = 5.5, height = 8)
 ```

### 2021-07-12 analysis/scripts/get_subgroup_comparisons.R
nearly done all changes

@@ -49,25 +49,23 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
         ggthemes::scale_colour_ptol(drop = FALSE) +
         theme(
             legend.position = 'bottom',
-            axis.text.y = element_text(
-            face = "italic", #if_else(l$normal==0, "italic", "plain"),
-            colour = l$lab_cols),
+            axis.text.y = element_text(face = "italic"),
             axis.title.y = element_blank(),
             plot.margin=unit(c(0,0.25,0,0.25),"cm"))  
        
      if (is_thresholds) {
         plt <- plt + coord_cartesian((xlim = c(-2.2, 0.7))) + 
-            scale_fill_viridis_d("p(difference < 0 | data)", option = "plasma", drop = FALSE)  
+            scale_fill_grey("p(difference < 0 | data)", start = 1, end = 0.1, drop = FALSE)  
     } else {
          plt <- plt + coord_cartesian((xlim = c(-0.5, 1.5))) + 
-            scale_fill_viridis_d("p(difference > 0 | data)", option = "plasma", drop = FALSE) 
+            scale_fill_grey("p(difference > 0 | data)", start = 1, end = 0.1, drop = FALSE) 
     }
 
-    if (fig_n == 1) {
+    #if (fig_n == 1) {
         plt <- plt + guides(colour = FALSE, fill = guide_legend(title = element_blank(), nrow = 2))
-    } else {
-        plt <- plt + guides(fill = FALSE, colour = guide_legend(title.postion = "top", nrow = 2))
-    }
+   # } else {
+    #    plt <- plt + guides(fill = FALSE, colour = guide_legend(title.postion = "top", nrow = 2))
+   # }
 
     
     return(plt)

### 2021-07-12 analysis/scripts/import_subgroup_comparisons.R

@@ -69,4 +69,4 @@ subgroups %>% filter(
     is.finite(index),
     group != subgroup,
     !(paste(group, subgroup, sep="-") %in% to_remove)) %>%
-    mutate(label = paste(group, "->", subgroup)) -> subgroups
+    mutate(label = paste(subgroup, " ", sprintf('\u2192') , " ", group, " (", index, ")", sep = "")) -> subgroups

### 2021-07-12 analysis/analysis.Rmd
best i can do

@@
 
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
+both_plts <- p1 + p2 +  plot_layout(guides = 'collect')  & theme(legend.position = 'bottom')
 both_plts

### 2021-07-12 analysis/scripts/get_subgroup_comparisons.R

@@ -46,12 +46,7 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
         scale_x_continuous('pdf for subgroup difference', expand = c(0,0), breaks = seq(-2, 1, 0.5)) +
         scale_y_discrete(labels = levels(d$key)) + #labels = lapply(levels(subgroup_comp$key), TeX)
         theme_minimal() +        
-        ggthemes::scale_colour_ptol(drop = FALSE) +
-        theme(
-            legend.position = 'bottom',
-            axis.text.y = element_text(face = "italic"),
-            axis.title.y = element_blank(),
-            plot.margin=unit(c(0,0.25,0,0.25),"cm"))  
+        ggthemes::scale_colour_ptol(drop = FALSE) 
        
      if (is_thresholds) {
         plt <- plt + coord_cartesian((xlim = c(-2.2, 0.7))) + 
@@ -62,7 +57,12 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
     }
 
     #if (fig_n == 1) {
-        plt <- plt + guides(colour = FALSE, fill = guide_legend(title = element_blank(), nrow = 2))
+        plt <- plt + guides(colour = FALSE, fill = guide_legend(title = element_blank(), nrow = 2)) +
+          theme(
+            legend.position = 'bottom',
+            axis.text.y = element_text(face = "italic"),
+            axis.title.y = element_blank(),
+            plot.margin=unit(c(0,0.25,0,0.25),"cm"))  
    # } else {
     #    plt <- plt + guides(fill = FALSE, colour = guide_legend(title.postion = "top", nrow = 2))
    # }

### 2021-07-12 analysis/scripts/get_subgroup_comparisons.R
done (give or take)

@@ -57,7 +57,7 @@ plot_comparisons <- function(d, l, fig_n, is_thresholds = FALSE)
     }
 
     #if (fig_n == 1) {
-        plt <- plt + guides(colour = FALSE, fill = guide_legend(title = element_blank(), nrow = 2)) +
+        plt <- plt + guides(colour = FALSE, fill = guide_legend(title = element_blank(), nrow = 1)) +
           theme(
             legend.position = 'bottom',
             axis.text.y = element_text(face = "italic"),

### 2021-07-12 analysis/analysis.Rmd
Does this work?

--- 2021-06-29 analysis/analysis.Rmd
@@
 
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
+both_plts <- p1 + p2 +  plot_layout(guides = 'collect')  & theme(legend.position = 'bottom')
 both_plts
-ggsave("plots/subgroup_comp_eeg_rms.pdf", width = 5.5, height = 8)
+ggsave("plots/subgroup_comp_eeg_rms.pdf", width = 5.5, height = 8, device =  cairo_pdf)
 ```
@@
 
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
+both_plts <- p1 + p2 +  plot_layout(guides = 'collect')  & theme(legend.position = 'bottom')
 both_plts
-ggsave("plots/subgroup_comp_psychophysical.pdf",  width = 5.5, height = 8)
+ggsave("plots/subgroup_comp_psychophysical.pdf",  width = 5.5, height = 8, device =  cairo_pdf)
 ```

### 2021-07-13 analysis/analysis.Rmd
Fixed sup mat legends

@@
 
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
+both_plts <- p1 + p2 + plot_layout(guides = 'collect')  & theme(legend.position = 'bottom')
 both_plts + plot_annotation(title = 'Odd harmonic data from parietal electrodes')
@@
 
-both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
+both_plts <- p1 + p2 + plot_layout(guides = 'collect')  & theme(legend.position = 'bottom')
 both_plts + plot_annotation(title = 'Even harmonic data from occipital electrodes')
